---
title: "Introduction to BayesianQDM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BayesianQDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(BayesianQDM)
```

## Overview

The BayesianQDM package provides comprehensive methods for Bayesian quantitative decision-making in clinical trials. This vignette demonstrates the basic functionality for both binary and continuous endpoints.

## Decision Framework

The package implements a three-zone decision framework:

- **Go**: Sufficient evidence to proceed (high posterior probability)
- **NoGo**: Insufficient evidence (low posterior probability)
- **Gray**: Inconclusive evidence (intermediate probability)

## Binary Endpoints

### Basic Example

```{r binary_example}
# Calculate decision probabilities for binary endpoint
result_binary <- BayesDecisionProbBinary(
  prob = 'posterior', 
  design = 'controlled', 
  theta.TV = 0.3, theta.MAV = 0.1,
  gamma1 = 0.8, gamma2 = 0.2,
  pi1 = c(0.3, 0.5, 0.7), 
  pi2 = rep(0.2, 3), 
  n1 = 15, n2 = 15,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)

print(result_binary)
```

### Posterior Probability

```{r binary_posterior}
# Calculate posterior probability
posterior_prob <- BayesPostPredBinary(
  prob = 'posterior', 
  design = 'controlled', 
  theta0 = 0.15,
  n1 = 12, n2 = 15, y1 = 7, y2 = 5, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)

cat("Posterior probability:", round(posterior_prob, 4))
```

## Continuous Endpoints

### Basic Example

```{r continuous_example}
# Calculate decision probabilities for continuous endpoint  
result_continuous <- BayesDecisionProbContinuous(
  nsim = 50,  # Small nsim for vignette speed
  prob = 'posterior', 
  design = 'controlled', 
  prior = 'vague', 
  CalcMethod = 'WS',  # Fast method for vignette
  theta.TV = 1.5, theta.MAV = 0.5, theta.NULL = NULL,  # All required thresholds
  nMC = NULL, nMCMCsample = NULL,  # Not needed for WS method
  gamma1 = 0.8, gamma2 = 0.3,
  n1 = 12, n2 = 12, m1 = NULL, m2 = NULL,
  mu1 = 4.5, mu2 = 2.0, sigma1 = 1.5, sigma2 = 1.3,
  seed = 123
)

print(result_continuous)
```

## Study Designs

The package supports three study designs:

1. **Controlled Design**: Standard two-arm randomized trial
2. **Uncontrolled Design**: Single-arm with historical control
3. **External Control Design**: Incorporating historical data via power priors

## Prior Distributions

### Binary Endpoints
- Beta priors for response probabilities
- Default: Beta(0.5, 0.5) for Jeffreys prior

### Continuous Endpoints
- **Vague priors**: Non-informative approach
- **Normal-Inverse-Chi-squared**: Conjugate priors for incorporating prior knowledge

## Calculation Methods

For continuous endpoints, four methods are available:

1. **NI (Numerical Integration)**: Exact calculation using convolution
2. **WS (Welch-Satterthwaite)**: Fast approximation for unequal variances
3. **MC (Monte Carlo)**: Simulation-based flexible approach
4. **MCMC (Markov Chain Monte Carlo)**: For external data incorporation with power priors

## Summary

The BayesianQDM package provides:

1. **Flexible decision frameworks** with customizable thresholds
2. **Multiple calculation methods** balancing accuracy and speed
3. **Support for various designs** including external data incorporation
4. **Comprehensive probability calculations** for evidence-based decisions

For more detailed examples, see the binary-endpoints and continuous-endpoints vignettes.
