---
title: "Bayesian Decision Making for Binary Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Decision Making for Binary Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.alt = "Visualization of Go/NoGo/Gray decision probabilities for binary endpoints"
)
```

```{r setup}
library(BayesianQDM)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Introduction

The BayesianQDM package provides comprehensive methods for Bayesian quantitative decision-making in clinical trials with binary endpoints. This vignette demonstrates how to use the package for calculating posterior probabilities, posterior predictive probabilities, and Go/NoGo/Gray decision probabilities.

## Basic Concepts

### Decision Framework

The Bayesian decision-making framework categorizes trial outcomes into three zones:

- **Go**: Evidence suggests the treatment is effective (proceed to next phase)
- **NoGo**: Evidence suggests the treatment is not effective (stop development)
- **Gray**: Evidence is inconclusive (may need additional data)

### Probability Types

Two types of probabilities can be calculated:

1. **Posterior Probability**: $P(\theta > \theta_0 | data)$ where $\theta$ is the treatment effect
2. **Posterior Predictive Probability**: Probability of future trial success

## Posterior and Posterior Predictive Probabilities

### Basic Usage

Let's start with calculating posterior probabilities for a controlled design:

```{r posterior_basic}
# Calculate posterior probability
posterior_prob <- BayesPostPredBinary(
  prob = 'posterior', 
  design = 'controlled', 
  theta0 = 0.15,
  n1 = 12, n2 = 15, y1 = 7, y2 = 9, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, # Not needed for posterior probability
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL # No external data
)

cat("Posterior probability that treatment effect > 0.15:", round(posterior_prob, 4))
```

Now let's calculate posterior predictive probability:

```{r predictive_basic}
# Calculate posterior predictive probability
predictive_prob <- BayesPostPredBinary(
  prob = 'predictive', 
  design = 'controlled', 
  theta0 = 0.1,
  n1 = 12, n2 = 15, y1 = 7, y2 = 7, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = 30, m2 = 30,    # Required for predictive probability
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL # No external data
)

cat("Posterior predictive probability that future trial effect > 0.1:", round(predictive_prob, 4))
```

### External Control Design

The package supports incorporating external (historical) control data using power priors:

```{r external_design}
# Posterior probability with external control data
external_prob <- BayesPostPredBinary(
  prob = 'posterior', 
  design = 'external', 
  theta0 = 0.15,
  n1 = 12, n2 = 15, y1 = 7, y2 = 9, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, # Not needed for posterior probability
  ne1 = 12, ne2 = 12, ye1 = 6, ye2 = 6, 
  ae1 = 0.5, ae2 = 0.5
)

cat("Posterior probability with external control:", round(external_prob, 4))
```

## Go/NoGo/Gray Decision Probabilities

### Controlled Design with Posterior Probability

```{r decision_posterior}
# Calculate Go/NoGo/Gray probabilities using posterior probability
decision_posterior <- BayesDecisionProbBinary(
  prob = 'posterior', 
  design = 'controlled', 
  theta.TV = 0.3,    # Go threshold
  theta.MAV = 0.1,   # NoGo threshold
  theta.NULL = NULL, # Not needed for posterior probability
  gamma1 = 0.8,      # Minimum probability for Go
  gamma2 = 0.2,      # Maximum probability for NoGo
  pi1 = c(0.2, 0.4, 0.6, 0.8), 
  pi2 = rep(0.2, 4), 
  n1 = 12, n2 = 12,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL,          # Not needed for controlled design
  m1 = NULL, m2 = NULL, # Not needed for posterior probability
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL # No external data
)

print(decision_posterior)
```

### Controlled Design with Posterior Predictive Probability

```{r decision_predictive}
# Calculate Go/NoGo/Gray probabilities using posterior predictive probability
decision_predictive <- BayesDecisionProbBinary(
  prob = 'predictive', 
  design = 'controlled', 
  theta.TV = NULL, theta.MAV = NULL, # Not needed for predictive probability
  theta.NULL = 0.15,  # Threshold for future trial
  gamma1 = 0.9,       # Minimum probability for Go
  gamma2 = 0.3,       # Maximum probability for NoGo
  pi1 = c(0.2, 0.4, 0.6, 0.8), 
  pi2 = rep(0.2, 4), 
  n1 = 12, n2 = 12,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL,           # Not needed for controlled design
  m1 = 30, m2 = 30,   # Required for predictive probability
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL # No external data
)

print(decision_predictive)
```

### Visualization of Decision Probabilities

```{r visualization, fig.width=10, fig.height=6, fig.alt="Comparison of posterior vs posterior predictive decision probabilities showing Go, Gray, and NoGo zones across different treatment effect values"}
# Create comprehensive decision probability data
# Split into separate scenarios to avoid vectorization issues
posterior_results <- BayesDecisionProbBinary(
  prob = 'posterior', design = 'controlled', 
  theta.TV = 0.3, theta.MAV = 0.1, theta.NULL = NULL, 
  gamma1 = 0.8, gamma2 = 0.2,
  pi1 = seq(0.1, 0.9, by = 0.05), pi2 = rep(0.2, length(seq(0.1, 0.9, by = 0.05))), 
  n1 = 12, n2 = 15, a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
) %>%
  mutate(prob = 'posterior', theta = pi1 - pi2)

predictive_results <- BayesDecisionProbBinary(
  prob = 'predictive', design = 'controlled', 
  theta.TV = NULL, theta.MAV = NULL, theta.NULL = 0.15, 
  gamma1 = 0.9, gamma2 = 0.3,
  pi1 = seq(0.1, 0.9, by = 0.05), pi2 = rep(0.2, length(seq(0.1, 0.9, by = 0.05))), 
  n1 = 12, n2 = 15, a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL, m1 = 30, m2 = 30,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
) %>%
  mutate(prob = 'predictive', theta = pi1 - pi2)

# Combine results
results <- bind_rows(posterior_results, predictive_results) %>%
  mutate(Gray = pmax(Gray, 0))  # Ensure non-negative Gray probabilities

# Create visualization
figure <- results %>%
  pivot_longer(
    cols = c(Go, NoGo, Gray), names_to = 'Decision', values_to = 'Prob'
  ) %>%
  mutate(
    prob = factor(prob, levels = c('posterior', 'predictive')),
    Decision = factor(Decision, levels = c('Go', 'Gray', 'NoGo'))
  ) %>%
  ggplot(aes(x = theta, y = Prob)) +
  facet_wrap(~ prob, labeller = labeller(prob = c(
    'posterior' = 'Posterior Probability',
    'predictive' = 'Posterior Predictive Probability'
  ))) +
  geom_line(aes(colour = Decision, linetype = Decision), linewidth = 1.2) +
  theme_bw() +
  scale_color_manual(
    values = c('Go' = '#2E8B57', 'Gray' = '#808080', 'NoGo' = '#DC143C'),
    labels = c('Go', 'Gray', 'NoGo')
  ) +
  scale_linetype_manual(
    values = c('Go' = 'solid', 'Gray' = 'dashed', 'NoGo' = 'dotted'),
    labels = c('Go', 'Gray', 'NoGo')
  ) +
  scale_x_continuous(
    name = expression(paste("Treatment Effect (", pi[1] - pi[2], ")")),
    breaks = seq(-0.1, 0.7, by = 0.2)
  ) +
  scale_y_continuous(
    name = "Probability",
    breaks = seq(0, 1, by = 0.2),
    limits = c(0, 1)
  ) +
  labs(
    title = "Go/Gray/NoGo Decision Probabilities for Binary Endpoints",
    subtitle = "Comparison of posterior vs. posterior predictive approaches"
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    strip.text = element_text(size = 11, face = "bold"),
    axis.title = element_text(size = 11),
    plot.title = element_text(size = 13, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )

print(figure)
```

## Advanced Examples

### Uncontrolled Design

For single-arm trials where there's no concurrent control group:

```{r uncontrolled}
# Uncont
