---
title: "Bayesian Decision Making for Binary Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Decision Making for Binary Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.alt = "Visualization of Go/NoGo/Gray decision probabilities for binary endpoints"
)
```

```{r setup}
library(BayesianQDM)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Introduction

The BayesianQDM package provides comprehensive methods for Bayesian quantitative decision-making in clinical trials with binary endpoints. This vignette demonstrates how to use the package for calculating posterior probabilities, posterior predictive probabilities, and Go/NoGo/Gray decision probabilities.

## Basic Concepts

### Decision Framework

The Bayesian decision-making framework categorizes trial outcomes into three zones:

- **Go**: Evidence suggests the treatment is effective (proceed to next phase)
- **NoGo**: Evidence suggests the treatment is not effective (stop development)
- **Gray**: Evidence is inconclusive (may need additional data)

### Probability Types

Two types of probabilities can be calculated:

1. **Posterior Probability**: $P(\theta > \theta_0 | data)$ where $\theta$ is the treatment effect
2. **Posterior Predictive Probability**: $P(future\ trial\ success | current\ data)$

## Basic Usage

### Posterior Probability

```{r posterior_basic}
# Calculate posterior probability for a single trial outcome
posterior_prob <- pPPsinglebinary(
  prob = 'posterior',
  design = 'controlled',
  theta0 = 0.15,  # Minimum clinically meaningful difference
  n1 = 20, n2 = 20,  # Sample sizes
  y1 = 12, y2 = 6,   # Observed responders
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,  # Beta prior parameters
  m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)

cat("Posterior probability that treatment effect > 0.15:", round(posterior_prob, 4))
```

### Posterior Predictive Probability

```{r predictive_basic}
# Calculate posterior predictive probability for future trial
predictive_prob <- pPPsinglebinary(
  prob = 'predictive',
  design = 'controlled',
  theta0 = 0.1,
  n1 = 20, n2 = 20,     # Current trial sample sizes
  y1 = 12, y2 = 6,      # Current trial outcomes
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = 100, m2 = 100,   # Future trial sample sizes
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)

cat("Probability of future trial success:", round(predictive_prob, 4))
```

## Go/NoGo/Gray Decision Making

### Operating Characteristics

Evaluate decision probabilities across different true response rates:

```{r operating_characteristics}
# Define true response rates to evaluate
true_rates <- expand.grid(
  pi1 = c(0.2, 0.3, 0.4, 0.5, 0.6),
  pi2 = 0.2
)

# Calculate Go/NoGo/Gray probabilities
oc_results <- pGNGsinglebinary(
  prob = 'posterior',
  design = 'controlled',
  theta.TV = 0.25,   # Target value for Go
  theta.MAV = 0.05,  # Minimum acceptable value for NoGo
  theta.NULL = NULL,
  gamma1 = 0.8,      # Go threshold
  gamma2 = 0.2,      # NoGo threshold
  pi1 = true_rates$pi1,
  pi2 = true_rates$pi2,
  n1 = 30, n2 = 30,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)

print(oc_results)
```

### Visualizing Decision Probabilities

```{r decision_plot, fig.width=8, fig.height=5}
# Reshape data for plotting
plot_data <- oc_results %>%
  select(pi1, pi2, Go, NoGo, Gray) %>%
  pivot_longer(cols = c(Go, NoGo, Gray), 
               names_to = "Decision", 
               values_to = "Probability")

# Create visualization
ggplot(plot_data, aes(x = pi1, y = Probability, color = Decision, group = Decision)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c('Go' = '#2E8B57', 'Gray' = '#808080', 'NoGo' = '#DC143C'),
    labels = c('Go', 'Gray', 'NoGo')
  ) +
  labs(
    title = "Operating Characteristics for Binary Endpoint Trial",
    subtitle = "Control response rate = 20%",
    x = "True Treatment Response Rate",
    y = "Decision Probability",
    color = "Decision"
  ) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```

## External Control Design

### Incorporating Historical Data

```{r external_design}
# Calculate probability with external control data
external_result <- pPPsinglebinary(
  prob = 'posterior',
  design = 'external',
  theta0 = 0.15,
  n1 = 20, n2 = 20,     # Current trial
  y1 = 12, y2 = 6,      # Current outcomes
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL,
  ne1 = 30, ne2 = 30,   # Historical trial sizes
  ye1 = 15, ye2 = 8,    # Historical outcomes
  ae1 = 0.5, ae2 = 0.5  # Power prior parameters (50% borrowing)
)

cat("Posterior probability with external data:", round(external_result, 4))
```

### Sensitivity to Historical Data Borrowing

```{r borrowing_sensitivity}
# Test different levels of borrowing
alpha_values <- seq(0, 1, by = 0.2)

borrowing_results <- data.frame(
  alpha = alpha_values,
  probability = sapply(alpha_values, function(a) {
    pPPsinglebinary(
      prob = 'posterior', design = 'external', theta0 = 0.15,
      n1 = 20, n2 = 20, y1 = 12, y2 = 6,
      a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
      m1 = NULL, m2 = NULL,
      ne1 = 30, ne2 = 30, ye1 = 15, ye2 = 8,
      ae1 = a, ae2 = a
    )
  })
)

print(borrowing_results)
```

```{r borrowing_plot, fig.width=7, fig.height=4}
ggplot(borrowing_results, aes(x = alpha, y = probability)) +
  geom_line(linewidth = 1.2, color = '#2E8B57') +
  geom_point(size = 3, color = '#2E8B57') +
  labs(
    title = "Sensitivity to Historical Data Borrowing",
    x = "Power Prior Parameter (α)",
    y = "Posterior Probability",
    subtitle = "α = 0: no borrowing, α = 1: full borrowing"
  ) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```

## Sample Size Considerations

### Power Analysis

```{r power_analysis}
# Evaluate Go probability across different sample sizes
sample_sizes <- seq(10, 50, by = 10)

power_results <- data.frame(
  n = sample_sizes,
  go_prob = sapply(sample_sizes, function(n) {
    result <- pGNGsinglebinary(
      prob = 'posterior', design = 'controlled',
      theta.TV = 0.2, theta.MAV = 0.05, theta.NULL = NULL,
      gamma1 = 0.8, gamma2 = 0.2,
      pi1 = 0.5, pi2 = 0.3,  # Assume true effect of 0.2
      n1 = n, n2 = n,
      a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
      z = NULL, m1 = NULL, m2 = NULL,
      ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
    )
    result$Go[1]
  })
)

print(power_results)
```

```{r power_plot, fig.width=7, fig.height=4}
ggplot(power_results, aes(x = n, y = go_prob)) +
  geom_line(linewidth = 1.2, color = '#2E8B57') +
  geom_point(size = 3, color = '#2E8B57') +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "red") +
  annotate("text", x = 45, y = 0.85, label = "Target: 80%", color = "red") +
  labs(
    title = "Power Analysis: Sample Size vs Go Probability",
    x = "Sample Size (per arm)",
    y = "Go Probability",
    subtitle = "True treatment effect = 20%"
  ) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```

## Practical Guidelines

### Threshold Selection

When designing a trial, consider:

- **θ_TV (Target Value)**: Set based on clinically meaningful difference
- **θ_MAV (Minimum Acceptable Value)**: Set based on smallest worthwhile effect
- **γ₁ (Go threshold)**: Typically 0.8-0.9 for high confidence
- **γ₂ (NoGo threshold)**: Typically 0.2-0.3 for early stopping

### Prior Selection

- **Weakly informative**: Beta(0.5, 0.5) - Jeffreys prior
- **Uniform**: Beta(1, 1) - uniform prior
- **Informative**: Beta(a, b) based on historical data or expert opinion

### Design Selection

1. **Controlled**: Standard two-arm trial (most common)
2. **Uncontrolled**: Single-arm when historical control is well-established
3. **External**: When incorporating historical data via power priors

## Summary

This vignette demonstrated:

1. **Basic probability calculations** for binary endpoints
2. **Go/NoGo/Gray decision framework** with customizable thresholds
3. **Operating characteristics** evaluation across scenarios
4. **External control design** with power priors
5. **Sample size considerations** for trial planning

The BayesianQDM package provides flexible tools for evidence-based decision making in clinical trials with binary endpoints.

## References

For more details on the statistical methods, see:

- The main BayesianQDM vignette for framework overview
- Function documentation for parameter specifications
- The continuous endpoints vignette for analysis of continuous outcomes
