---
title: "Bayesian Decision Making for Binary Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Decision Making for Binary Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(BayesianQDM)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Introduction

The BayesianQDM package provides comprehensive methods for Bayesian quantitative decision-making in clinical trials with binary endpoints. This vignette demonstrates how to use the package for calculating posterior probabilities, posterior predictive probabilities, and Go/NoGo/Gray decision probabilities.

## Basic Concepts

### Decision Framework

The Bayesian decision-making framework categorizes trial outcomes into three zones:

- **Go**: Evidence suggests the treatment is effective (proceed to next phase)
- **NoGo**: Evidence suggests the treatment is not effective (stop development)
- **Gray**: Evidence is inconclusive (may need additional data)

### Probability Types

Two types of probabilities can be calculated:

1. **Posterior Probability**: $P(\theta > \theta_0 | data)$ where $\theta$ is the treatment effect
2. **Posterior Predictive Probability**: Probability of future trial success

## Posterior and Posterior Predictive Probabilities

### Basic Usage

Let's start with calculating posterior probabilities for a controlled design:

```{r posterior_basic}
# Calculate posterior probability
posterior_prob <- BayesPostPredBinary(
  prob = 'posterior', 
  design = 'controlled', 
  theta0 = 0.15,
  n1 = 12, n2 = 15, y1 = 7, y2 = 9, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5
)

cat("Posterior probability that treatment effect > 0.15:", round(posterior_prob, 4))
```

Now let's calculate posterior predictive probability:

```{r predictive_basic}
# Calculate posterior predictive probability
predictive_prob <- BayesPostPredBinary(
  prob = 'predictive', 
  design = 'controlled', 
  theta0 = 0.1,
  n1 = 12, n2 = 15, y1 = 7, y2 = 7, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = 30, m2 = 30
)

cat("Posterior predictive probability that future trial effect > 0.1:", round(predictive_prob, 4))
```

### External Control Design

The package supports incorporating external (historical) control data using power priors:

```{r external_design}
# Posterior probability with external control data
external_prob <- BayesPostPredBinary(
  prob = 'posterior', 
  design = 'external', 
  theta0 = 0.15,
  n1 = 12, n2 = 15, y1 = 7, y2 = 9, 
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  ne1 = 12, ne2 = 12, ye1 = 6, ye2 = 6, 
  ae1 = 0.5, ae2 = 0.5
)

cat("Posterior probability with external control:", round(external_prob, 4))
```

## Go/NoGo/Gray Decision Probabilities

### Controlled Design with Posterior Probability

```{r decision_posterior}
# Calculate Go/NoGo/Gray probabilities using posterior probability
decision_posterior <- BayesDecisionProbBinary(
  prob = 'posterior', 
  design = 'controlled', 
  theta.TV = 0.3,    # Go threshold
  theta.MAV = 0.1,   # NoGo threshold
  gamma1 = 0.8,      # Minimum probability for Go
  gamma2 = 0.2,      # Maximum probability for NoGo
  pi1 = c(0.2, 0.4, 0.6, 0.8), 
  pi2 = rep(0.2, 4), 
  n1 = 12, n2 = 12,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5
)

print(decision_posterior)
```

### Controlled Design with Posterior Predictive Probability

```{r decision_predictive}
# Calculate Go/NoGo/Gray probabilities using posterior predictive probability
decision_predictive <- BayesDecisionProbBinary(
  prob = 'predictive', 
  design = 'controlled', 
  theta.NULL = 0.15,  # Threshold for future trial
  gamma1 = 0.9,       # Minimum probability for Go
  gamma2 = 0.3,       # Maximum probability for NoGo
  pi1 = c(0.2, 0.4, 0.6, 0.8), 
  pi2 = rep(0.2, 4), 
  n1 = 12, n2 = 12,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  m1 = 30, m2 = 30
)

print(decision_predictive)
```

### Visualization of Decision Probabilities

```{r visualization, fig.width=10, fig.height=6}
# Create comprehensive decision probability data
results <- tibble(
  design = rep(c('controlled', 'controlled'), each = 2),
  prob = rep(c('posterior', 'predictive'), 2),
  theta.TV = c(0.3, NA, 0.3, NA),
  theta.MAV = c(0.1, NA, 0.1, NA),
  theta.NULL = c(NA, 0.15, NA, 0.15),
  gamma1 = c(0.8, 0.9, 0.8, 0.9),
  gamma2 = c(0.2, 0.3, 0.2, 0.3)
) %>%
  group_by_all() %>%
  reframe(
    GoNoGoGray = map(list(1), ~ {
      BayesDecisionProbBinary(
        prob = prob, design = design, 
        theta.TV = theta.TV, theta.MAV = theta.MAV, theta.NULL = theta.NULL, 
        gamma1 = gamma1, gamma2 = gamma2,
        pi1 = seq(0.1, 0.9, by = 0.05), pi2 = rep(0.2, length(seq(0.1, 0.9, by = 0.05))), 
        n1 = 12, n2 = 15, a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
        m1 = if(prob == 'predictive') 30 else NULL, 
        m2 = if(prob == 'predictive') 30 else NULL
      )
    })
  ) %>%
  unnest(GoNoGoGray) %>%
  mutate(
    Gray = pmax(Gray, 0),  # Ensure non-negative Gray probabilities
    theta = pi1 - pi2
  )

# Create visualization
figure <- results %>%
  pivot_longer(
    cols = c(Go, NoGo, Gray), names_to = 'Decision', values_to = 'Prob'
  ) %>%
  mutate(
    prob = factor(prob, levels = c('posterior', 'predictive')),
    Decision = factor(Decision, levels = c('Go', 'Gray', 'NoGo'))
  ) %>%
  ggplot(aes(x = theta, y = Prob)) +
  facet_wrap(~ prob, labeller = labeller(prob = c(
    'posterior' = 'Posterior Probability',
    'predictive' = 'Posterior Predictive Probability'
  ))) +
  geom_line(aes(colour = Decision, linetype = Decision), linewidth = 1.2) +
  theme_bw() +
  scale_color_manual(
    values = c('Go' = '#2E8B57', 'Gray' = '#808080', 'NoGo' = '#DC143C'),
    labels = c('Go', 'Gray', 'NoGo')
  ) +
  scale_linetype_manual(
    values = c('Go' = 'solid', 'Gray' = 'dashed', 'NoGo' = 'dotted'),
    labels = c('Go', 'Gray', 'NoGo')
  ) +
  scale_x_continuous(
    name = expression(paste("Treatment Effect (", pi[1] - pi[2], ")")),
    breaks = seq(-0.1, 0.7, by = 0.2)
  ) +
  scale_y_continuous(
    name = "Probability",
    breaks = seq(0, 1, by = 0.2),
    limits = c(0, 1)
  ) +
  labs(
    title = "Go/Gray/NoGo Decision Probabilities for Binary Endpoints",
    subtitle = "Comparison of posterior vs. posterior predictive approaches"
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    strip.text = element_text(size = 11, face = "bold"),
    axis.title = element_text(size = 11),
    plot.title = element_text(size = 13, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )

print(figure)
```

## Advanced Examples

### Uncontrolled Design

For single-arm trials where there's no concurrent control group:

```{r uncontrolled}
# Uncontrolled design example
uncontrolled_result <- BayesDecisionProbBinary(
  prob = 'posterior', 
  design = 'uncontrolled', 
  theta.TV = 0.3, 
  theta.MAV = 0.1,
  gamma1 = 0.8, 
  gamma2 = 0.2,
  pi1 = c(0.3, 0.5, 0.7), 
  pi2 = rep(0.2, 3),  # Historical control rate
  n1 = 20, 
  n2 = NULL,  # No concurrent control
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = 3  # Fixed number of historical control responders
)

print(uncontrolled_result)
```

### External Control with Power Prior

Incorporating historical data with different power prior weights:

```{r external_comparison}
# Compare different power prior weights
power_weights <- c(0.25, 0.5, 0.75, 1.0)
external_results <- map_dfr(power_weights, ~ {
  result <- BayesDecisionProbBinary(
    prob = 'posterior', 
    design = 'external', 
    theta.TV = 0.25, 
    theta.MAV = 0.05,
    gamma1 = 0.8, 
    gamma2 = 0.2,
    pi1 = c(0.3, 0.5), 
    pi2 = rep(0.2, 2), 
    n1 = 15, n2 = 15,
    a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
    ne1 = 20, ne2 = 20, ye1 = 8, ye2 = 4, 
    ae1 = .x, ae2 = .x
  )
  result$power_weight <- .x
  result
})

print(external_results)
```

## Interpretation Guidelines

### Go Decision
- High posterior/predictive probability that treatment effect exceeds efficacy threshold
- Proceed to next phase of development
- Typical threshold: P(θ > θ_TV | data) ≥ 0.8

### NoGo Decision  
- Low posterior/predictive probability that treatment effect exceeds minimal acceptable value
- Stop development or major protocol modification needed
- Typical threshold: P(θ > θ_MAV | data) ≤ 0.2

### Gray Zone
- Inconclusive evidence
- Consider additional patients, protocol modifications, or biomarker analysis
- May warrant conditional advancement with risk mitigation

## Parameter Selection Guidelines

### Prior Selection
- **Vague priors** (a = b = 0.5): Minimal prior influence, data-driven
- **Jeffreys priors** (a = b = 0.5): Non-informative
- **Informative priors**: Incorporate historical knowledge

### Threshold Selection
- **θ_TV (Target Value)**: Clinically meaningful effect size
- **θ_MAV (Minimal Acceptable Value)**: Smallest worthwhile effect
- **γ1**: Confidence level for Go decision (typically 0.8-0.9)
- **γ2**: Maximum acceptable probability for NoGo (typically 0.2-0.3)

### Sample Size Considerations
- Larger sample sizes provide more precise estimates
- Balance statistical power with resource constraints
- Consider interim analyses for adaptive designs

## Summary

The BayesianQDM package provides flexible tools for:

1. **Posterior probability calculations** for treatment effects
2. **Posterior predictive probability** for future trial success  
3. **Go/NoGo/Gray decision frameworks** with customizable thresholds
4. **Multiple design types** including controlled, uncontrolled, and external control
5. **Power prior integration** for historical data incorporation

This Bayesian approach enables evidence-based decision making that accounts for uncertainty and incorporates prior knowledge, making it particularly valuable for early-phase clinical trial decision making.
