---
title: "Overview of BayesianQDM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of BayesianQDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(BayesianQDM)
```

## Introduction

The **BayesianQDM** package provides a comprehensive framework for Bayesian
quantitative decision-making (QDM) in clinical trials.  It implements the
three-zone Go/NoGo/Gray decision rule under two complementary probability
metrics:

* **Posterior probability** — assesses whether the observed PoC data provide
  sufficient evidence that the treatment effect exceeds a clinically meaningful
  threshold.
* **Posterior predictive probability** — assesses the probability that a
  future Phase III trial would be successful given the PoC results.

Both metrics are available for single and two-endpoint analyses, covering
binary and continuous outcomes.

## The Three-Zone Decision Framework

For a single pre-specified threshold $\theta_0$, let
$p = P(\theta > \theta_0 \mid \text{data})$.  The decision rule is:

| Decision | Criterion |
|----------|-----------|
| **Go**   | $p \ge \gamma_1$ |
| **NoGo** | $p \le \gamma_2$ |
| **Gray** | $\gamma_2 < p < \gamma_1$ |

where $\gamma_1 > \gamma_2$ are pre-specified probability thresholds.

For the **posterior** metric, two thresholds are used simultaneously:

* **Go**: $P(\theta > \text{TV} \mid \text{data}) \ge \gamma_1$
* **NoGo**: $P(\theta > \text{MAV} \mid \text{data}) \le \gamma_2$

where TV (Target Value) and MAV (Minimum Acceptable Value) satisfy
TV $>$ MAV.

## Supported Designs and Endpoint Types

```{r design-table, echo = FALSE}
tab <- data.frame(
  Feature    = c("Endpoint type", "Number of endpoints",
                 "PoC design", "External data"),
  Options    = c("Binary, Continuous",
                 "Single, Two",
                 "Controlled, Uncontrolled, External control",
                 "Power prior (conjugate representation)")
)
knitr::kable(tab, col.names = c("Feature", "Available options"))
```

For **single continuous endpoints** three calculation methods are available:

* `'NI'` — numerical integration (exact)
* `'MC'` — Monte Carlo simulation
* `'MM'` — Moment-Matching approximation (fast; recommended for operating
  characteristics evaluation)

For **two continuous endpoints** `'MC'` and `'MM'` are available.

## Motivating Example: Rheumatoid Arthritis PoC Trial

Throughout the vignettes we use a hypothetical Phase IIa proof-of-concept
(PoC) trial in moderate-to-severe rheumatoid arthritis (RA).  The trial
compares an investigational drug (treatment) with placebo (control).

```{r scenario, echo = FALSE}
s <- data.frame(
  Item   = c("Indication", "Phase", "Design",
             "PoC sample size", "Primary endpoint (binary)",
             "Primary endpoint (continuous)",
             "Co-primary endpoints (two-endpoint)"),
  Detail = c("Rheumatoid arthritis (RA)", "Phase IIa PoC",
             "Randomised, double-blind, parallel-group",
             "n1 = n2 = 10 per arm",
             "ACR20 response rate",
             "Change from baseline in DAS28 score",
             "ACR20 + ACR50 response rates; DAS28 + HAQ-DI change")
)
knitr::kable(s, col.names = c("Item", "Detail"))
```

## Quick-Start: Single Binary Endpoint

```{r binary-quick}
# Posterior probability: P(pi_treat - pi_ctrl > 0.10 | data)
# Observed: 7/10 responders in treatment, 3/10 in control
pbayespostpred1bin(
  prob   = 'posterior',
  design = 'controlled',
  theta0 = 0.10,           # MAV threshold (treatment difference)
  n1 = 10, n2 = 10,
  y1 =  7, y2 =  3,
  a1 = 0.5, b1 = 0.5,     # Uniform Beta prior
  a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)
```

## Quick-Start: Single Continuous Endpoint

```{r cont-quick}
# Posterior probability: P(mu_treat - mu_ctrl > 1.0 | data)
# Observed means: 3.5 (treatment), 1.2 (control); SD ~2.0
pbayespostpred1cont(
  prob       = 'posterior',
  design     = 'controlled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta0     = 1.0,
  nMC        = NULL,
  n1 = 10, n2 = 10,
  bar.y1 = 3.5, s1 = 2.0,
  bar.y2 = 1.2, s2 = 2.0,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL
)
```

## Quick-Start: Go/NoGo Decision (Single Binary)

```{r gng-quick}
# Operating characteristics across true response rate scenarios
pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'controlled',
  theta.TV  = 0.20,   theta.MAV  = 0.05, theta.NULL = NULL,
  gamma1    = 0.80,   gamma2     = 0.20,
  pi1       = c(0.20, 0.35, 0.50),   # True treatment rates
  pi2       = rep(0.15, 3),           # True control rate
  n1 = 10,  n2 = 10,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)
```

## Package Structure

The package functions follow a consistent naming convention:

| Prefix | Meaning |
|--------|---------|
| `pbayespostpred1bin`  | Posterior/predictive probability, 1 binary endpoint |
| `pbayespostpred1cont` | Posterior/predictive probability, 1 continuous endpoint |
| `pbayespostpred2bin`  | Posterior/predictive probability, 2 binary endpoints |
| `pbayespostpred2cont` | Posterior/predictive probability, 2 continuous endpoints |
| `pbayesdecisionprob1bin`  | Go/NoGo/Gray probabilities, 1 binary endpoint |
| `pbayesdecisionprob1cont` | Go/NoGo/Gray probabilities, 1 continuous endpoint |
| `pbayesdecisionprob2bin`  | Go/NoGo/Gray probabilities, 2 binary endpoints |
| `pbayesdecisionprob2cont` | Go/NoGo/Gray probabilities, 2 continuous endpoints |

Helper / utility functions:

| Function | Purpose |
|----------|---------|
| `pbetadiff`      | CDF of the difference of two Beta random variables |
| `pbetabinomdiff` | CDF of the difference of two Beta-Binomial proportions |
| `ptdiff_NI`, `ptdiff_MC`, `ptdiff_MM` | CDF of the difference of two non-standardised t variables |
| `rdirichlet`  | Sample from a Dirichlet distribution |
| `rnsbt`       | Sample from a bivariate non-standardised t distribution |
| `getjointbin` | Joint bivariate binary probabilities from marginals and correlation |
| `allmultinom` | Enumerate all multinomial outcome vectors |

## Further Reading

* `vignette("single-endpoint")` — detailed examples for single binary and
  continuous endpoints, covering all three designs and prior types.
* `vignette("two-endpoints")` — detailed examples for two binary and two
  continuous endpoints, including the 9-region decision grid.
