---
title: "Overview of BayesianQDM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of BayesianQDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(BayesianQDM)
```

## Introduction

The **BayesianQDM** package provides a comprehensive framework for Bayesian
Quantitative Decision-Making (QDM) in clinical trials. It implements a
three-zone Go/NoGo/Gray decision rule under two complementary probability
metrics:

- **Posterior probability**: quantifies the evidence from observed PoC data
  that the treatment effect exceeds a clinically meaningful threshold.
- **Posterior predictive probability**: quantifies the probability that a
  future (Phase III) trial would succeed given the observed PoC results.

Both metrics are available for single and two-endpoint analyses, covering
binary and continuous outcomes, across three study designs (controlled,
uncontrolled, external control). The package also provides grid-search
functions (`getgamma` family) to find optimal Go/NoGo thresholds satisfying
user-specified operating characteristic criteria.

---

## 1. Bayesian Model and Posterior Distributions

### 1.1 Binary Endpoint: Beta-Binomial Conjugate Model

Let $\pi_j$ denote the response rate in group $j$ ($j = 1$: treatment,
$j = 2$: control).  We place independent Beta priors on each rate:

$$\pi_j \sim \mathrm{Beta}(a_j,\, b_j).$$

Given $n_j$ patients with $y_j$ responders, the posterior is

$$\pi_j \mid y_j \;\sim\; \mathrm{Beta}\!\left(a_j + y_j,\; b_j + n_j - y_j\right).$$

The treatment effect of interest is the difference $\theta = \pi_1 - \pi_2$.
Because the two groups are independent, the posterior of $\theta$ is the
distribution of the difference of two independent Beta random variables,
which has no closed-form CDF. The package evaluates

$$P(\theta > \theta_0 \mid \mathbf{y}) = P(\pi_1 - \pi_2 > \theta_0 \mid \mathbf{y})$$

by numerical integration via `pbetadiff()`:

$$P(\pi_1 - \pi_2 > \theta_0) = \int_{\max(0,\,\theta_0)}^{1}
  f_{\mathrm{Beta}(a_1^*, b_1^*)}(u)\;
  \bigl[1 - F_{\mathrm{Beta}(a_2^*, b_2^*)}(u - \theta_0)\bigr]\, du,$$

where $a_j^* = a_j + y_j$ and $b_j^* = b_j + n_j - y_j$ are the posterior
shape parameters.

### 1.2 Continuous Endpoint: Normal-Inverse-Chi-Squared Conjugate Model

Let $\mu_j$ and $\sigma_j^2$ denote the mean and variance of the outcome in
group $j$.  Under the Normal-Inverse-Chi-squared (N-Inv-$\chi^2$) conjugate
prior

$$(\mu_j,\, \sigma_j^2) \;\sim\; \mathrm{N\text{-}Inv\text{-}\chi^2}
  (\mu_{0j},\, \kappa_{0j},\, \nu_{0j},\, \sigma_{0j}^2),$$

where $\mu_{0j}$ is the prior mean, $\kappa_{0j}$ the prior precision
(pseudo-sample size), $\nu_{0j}$ the prior degrees of freedom, and
$\sigma_{0j}^2$ the prior scale.

Given sample mean $\bar{y}_j$, sample standard deviation $s_j$, and sample
size $n_j$, the posterior parameters are updated as

$$\kappa_{nj} = \kappa_{0j} + n_j, \qquad
  \nu_{nj} = \nu_{0j} + n_j,$$

$$\mu_{nj} = \frac{\kappa_{0j}\,\mu_{0j} + n_j\,\bar{y}_j}{\kappa_{nj}},$$

$$\sigma_{nj}^2 = \frac{\nu_{0j}\,\sigma_{0j}^2 + (n_j-1)\,s_j^2 +
  \dfrac{n_j\,\kappa_{0j}}{\kappa_{nj}}(\mu_{0j} - \bar{y}_j)^2}{\nu_{nj}}.$$

The marginal posterior of $\mu_j$ is a non-standardised $t$-distribution:

$$\mu_j \mid \mathbf{y} \;\sim\; t_{\nu_{nj}}\!\left(\mu_{nj},\;
  \frac{\sigma_{nj}^2}{\kappa_{nj}}\right),$$

i.e., location $\mu_{nj}$, scale $\sigma_{nj}/\sqrt{\kappa_{nj}}$, and
$\nu_{nj}$ degrees of freedom.

**Vague (Jeffreys) prior** — the limiting case $\kappa_{0j} \to 0$,
$\nu_{0j} \to 0$ yields

$$\mu_j \mid \mathbf{y} \;\sim\; t_{n_j - 1}\!\left(\bar{y}_j,\;
  \frac{s_j}{\sqrt{n_j}}\right).$$

The posterior of the treatment effect $\theta = \mu_1 - \mu_2$ is the
distribution of the difference of two independent non-standardised
$t$-variables. Three methods are provided to evaluate
$P(\theta > \theta_0 \mid \mathbf{y})$; see Section 3 for details.

---

## 2. Posterior Predictive Probability

### 2.1 Binary Endpoint

Let $\tilde{y}_j \mid \pi_j \sim \mathrm{Binomial}(m_j, \pi_j)$ denote
future trial outcomes in group $j$ with future sample size $m_j$.
Integrating out the unknown response rates under their posteriors gives the
posterior predictive (Beta-Binomial) distribution:

$$P(\tilde{y}_j = k) = \binom{m_j}{k}
  \frac{B(a_j^* + k,\; b_j^* + m_j - k)}{B(a_j^*,\, b_j^*)},
  \quad k = 0, 1, \ldots, m_j,$$

where $B(\cdot,\cdot)$ is the Beta function. The posterior predictive
probability of a successful future trial is then

$$P\!\left(\tilde{\theta} > \theta_0 \mid \mathbf{y}\right) =
  \sum_{k_1=0}^{m_1}\sum_{k_2=0}^{m_2}
  \mathbf{1}[k_1/m_1 - k_2/m_2 > \theta_0]\;
  P(\tilde{y}_1 = k_1)\; P(\tilde{y}_2 = k_2),$$

evaluated exactly by `pbetabinomdiff()`.

### 2.2 Continuous Endpoint

The predictive distribution of a single future observation
$\tilde{y}_j$ under the N-Inv-$\chi^2$ posterior is a non-standardised
$t$-distribution. For the future sample mean $\bar{\tilde{y}}_j$ based
on $m_j$ future patients, the scale is reduced by $\sqrt{m_j}$:

$$\bar{\tilde{y}}_j \mid \mathbf{y} \;\sim\;
  t_{\nu_{nj}}\!\left(\mu_{nj},\;
  \sigma_{nj}\sqrt{\frac{1 + \kappa_{nj}}{\kappa_{nj}\, m_j}}\right).$$

Under the vague prior ($\kappa_{0j} = \nu_{0j} = 0$, so $\kappa_{nj} = n_j$,
$\sigma_{nj} = s_j$) this becomes

$$\bar{\tilde{y}}_j \mid \mathbf{y} \;\sim\;
  t_{n_j - 1}\!\left(\bar{y}_j,\;
  s_j\sqrt{\frac{n_j + 1}{n_j\, m_j}}\right).$$

The posterior predictive probability of the future treatment effect
exceeding $\theta_0$ is

$$P\!\left(\bar{\tilde{y}}_1 - \bar{\tilde{y}}_2 > \theta_0
  \mid \mathbf{y}\right),$$

computed by the same three methods as the posterior probability (Section 3),
with the predictive scale replacing the posterior scale.

---

## 3. Computation Methods for Continuous Endpoints

The core computational challenge is evaluating the CDF of the difference
$D = T_1 - T_2$ of two independent non-standardised $t$-variables

$$T_j \sim t_{\nu_j}(\mu_j, \sigma_j), \qquad j = 1, 2,$$

at a threshold $\theta_0$. Three methods are available.

### 3.1 Numerical Integration (NI)

The CDF is expressed as a one-dimensional integral by integrating out $T_1$:

$$P(D \le \theta_0) =
  \int_{-\infty}^{\infty}
  F_{t_{\nu_2}(\mu_2, \sigma_2)}\!\left(t - \theta_0\right)\;
  f_{t_{\nu_1}(\mu_1, \sigma_1)}(t)\; dt,$$

where $F_{t_\nu(\mu,\sigma)}$ denotes the CDF and $f_{t_\nu(\mu,\sigma)}$
the PDF of the non-standardised $t$-distribution. This integral is evaluated
by `stats::integrate()` (adaptive Gauss-Kronrod quadrature) in `ptdiff_NI()`.
This method is exact but evaluates the integral once per call; it is therefore
suitable for single-probability evaluations but slow for large-scale
simulations.

### 3.2 Monte Carlo Simulation (MC)

Independent samples are drawn from each marginal distribution:

$$T_1^{(i)} \sim t_{\nu_1}(\mu_1, \sigma_1), \quad
  T_2^{(i)} \sim t_{\nu_2}(\mu_2, \sigma_2), \quad
  i = 1, \ldots, n_{\mathrm{MC}},$$

and the probability is estimated as the empirical proportion:

$$\widehat{P}(D > \theta_0) =
  \frac{1}{n_{\mathrm{MC}}}
  \sum_{i=1}^{n_{\mathrm{MC}}} \mathbf{1}\!\left[T_1^{(i)} - T_2^{(i)} > \theta_0\right].$$

The package implements this in `ptdiff_MC()` using a vectorised approach:
when called from `pbayesdecisionprob1cont()` with a vector of $n_{\mathrm{sim}}$
scenarios, all draws are generated as a single $n_{\mathrm{MC}} \times n_{\mathrm{sim}}$
matrix to avoid repeated loop overhead.

### 3.3 Moment-Matching Approximation (MM)

The first two even moments (mean and variance) of $D = T_1 - T_2$ are
matched to a single non-standardised $t$-distribution. For a
non-standardised $t$-variable $T \sim t_\nu(\mu, \sigma)$:

$$E[T] = \mu \quad (\nu > 1), \qquad
  \mathrm{Var}(T) = \sigma^2 \cdot \frac{\nu}{\nu - 2} \quad (\nu > 2).$$

The mean and variance of $D$ are therefore

$$E[D] = \mu_1 - \mu_2,$$

$$\mathrm{Var}(D) = \sigma_1^2\,\frac{\nu_1}{\nu_1 - 2} +
  \sigma_2^2\,\frac{\nu_2}{\nu_2 - 2}.$$

A single $t$-distribution $t_{\nu^*}(\mu^*, \sigma^*)$ is fitted by
equating its mean and variance to those of $D$:

$$\mu^* = \mu_1 - \mu_2, \qquad
  \sigma^{*2} \cdot \frac{\nu^*}{\nu^* - 2} = \mathrm{Var}(D).$$

The degrees of freedom $\nu^*$ and scale $\sigma^*$ are determined from
the fourth moment matching or by a closed-form relationship derived from
the Welch-Satterthwaite approximation. This yields a fully closed-form,
vectorised approximation implemented in `ptdiff_MM()`, making it the
recommended method for large-scale simulation studies.

---

## 4. Study Designs

### 4.1 Controlled Design

The standard parallel-group randomised controlled trial (RCT) with
concurrent treatment ($j = 1$) and control ($j = 2$) arms. Both
posterior distributions are updated from the observed PoC data, and the
probability is evaluated for the treatment effect $\theta = \pi_1 - \pi_2$
(binary) or $\theta = \mu_1 - \mu_2$ (continuous).

### 4.2 Uncontrolled Design (Single-Arm with Hypothetical Control)

When no concurrent control arm is available (single-arm PoC trial), the
control distribution is specified from external knowledge rather than
observed data.

**Binary endpoint**: a hypothetical responder count $z$ is specified for the
control group of size $n_2$, yielding posterior shape parameters
$a_2^* = a_2 + z$ and $b_2^* = b_2 + n_2 - z$.

**Continuous endpoint**: the hypothetical control mean $\mu_{02}$ is fixed
from prior knowledge and the control variance is scaled relative to the
treatment posterior variance via a scaling factor $r > 0$:

$$\sigma_{\mathrm{control}} = \sqrt{r} \cdot \sigma_{\mathrm{treatment}},$$

so the scale of the control $t$-distribution is $\sqrt{r}$ times that of
the treatment.

### 4.3 External Control Design (Power Prior)

Historical or external data are incorporated through a **power prior** with
borrowing weight $\alpha_e \in (0, 1]$. Under the Beta-Binomial model for
group $j$, the power prior augments the initial prior by a fraction
$\alpha_{ej}$ of the external data sufficient statistics:

$$(\alpha_j^*,\; \beta_j^*) =
  \left(a_j + \alpha_{ej}\, y_{ej},\;\;
  b_j + \alpha_{ej}(n_{ej} - y_{ej})\right),$$

where $n_{ej}$ and $y_{ej}$ are the external sample size and responder
count for group $j$. These augmented parameters then serve as the prior for
the current PoC data, yielding a closed-form posterior without MCMC.

Under the N-Inv-$\chi^2$ model, the power prior inflates the prior
precision and degrees of freedom by the weighted external sample:

$$\kappa_{0j}^* = \alpha_{ej}\, n_{ej}, \quad
  \nu_{0j}^* = \alpha_{ej}\, n_{ej} - 1,$$

$$\mu_{0j}^* = \bar{y}_{ej}, \quad
  \sigma_{0j}^{*2} = s_{ej}^2,$$

leading to the posterior updates

$$\mu_{nj} = \frac{\alpha_{ej} n_{ej}\, \bar{y}_{ej} + n_j\, \bar{y}_j}
  {\alpha_{ej} n_{ej} + n_j},$$

$$\sigma_{nj}^2 = \frac{\alpha_{ej}(n_{ej}-1) s_{ej}^2 + (n_j-1) s_j^2 +
  \dfrac{\alpha_{ej} n_{ej} n_j (\bar{y}_{ej} - \bar{y}_j)^2}
        {\alpha_{ej} n_{ej} + n_j}}
  {\alpha_{ej} n_{ej} + n_j},$$

with degrees of freedom $\nu_{nj} = \alpha_{ej} n_{ej} + n_j - 1$.
Setting $\alpha_{ej} = 1$ corresponds to full borrowing; $\alpha_{ej} \to 0$
recovers the vague-prior result from the current data alone.

---

## 5. The Three-Zone Go/NoGo/Gray Decision Framework

### 5.1 Single Endpoint

The Go/NoGo/Gray decision rule uses two pre-specified thresholds
$\gamma_1 > \gamma_2$ applied to a pair of probabilities
$(g_{\mathrm{Go}},\, g_{\mathrm{NoGo}})$:

$$\text{Decision} = \begin{cases}
  \textbf{Go}   & \text{if } g_{\mathrm{Go}}   \ge \gamma_1 \text{ AND }
                              g_{\mathrm{NoGo}} <  \gamma_2 \\
  \textbf{NoGo} & \text{if } g_{\mathrm{Go}}   <  \gamma_1 \text{ AND }
                              g_{\mathrm{NoGo}} \ge \gamma_2 \\
  \textbf{Miss} & \text{if } g_{\mathrm{Go}}   \ge \gamma_1 \text{ AND }
                              g_{\mathrm{NoGo}} \ge \gamma_2 \\
  \textbf{Gray} & \text{otherwise (neither Go nor NoGo criterion met).}
\end{cases}$$

A Miss occurs when both Go and NoGo criteria are simultaneously satisfied,
which indicates an inconsistent threshold configuration. By default the
functions stop with an error if Miss probability $> 0$ (`error_if_Miss = TRUE`).

For **posterior probability**, $\theta_{\mathrm{TV}}$ (Target Value) and
$\theta_{\mathrm{MAV}}$ (Minimum Acceptable Value,
$\theta_{\mathrm{TV}} > \theta_{\mathrm{MAV}}$) define two separate
probabilities:

$$g_{\mathrm{Go}} = P(\theta > \theta_{\mathrm{TV}} \mid \mathbf{y}), \qquad
  g_{\mathrm{NoGo}} = P(\theta \le \theta_{\mathrm{MAV}} \mid \mathbf{y}).$$

For **posterior predictive probability**, a single null threshold
$\theta_{\mathrm{NULL}}$ is used and the same probability serves both roles:

$$g_{\mathrm{Go}} = g_{\mathrm{NoGo}} =
  P(\tilde{\theta} > \theta_{\mathrm{NULL}} \mid \mathbf{y}),$$

so the decision reduces to Go if $g \ge \gamma_1$, NoGo if $g < \gamma_2$,
and Gray otherwise (Miss cannot occur in this case).

### 5.2 Operating Characteristics

For a given true parameter value (e.g., true treatment effect), the
**operating characteristics** are the marginal probabilities of each
decision outcome:

$$\Pr(\mathrm{Go}) = \sum_{\mathbf{y}} \mathbf{1}[\mathrm{decision}(\mathbf{y}) = \mathrm{Go}]\;
  P(\mathbf{y} \mid \text{true parameter}),$$

and analogously for $\Pr(\mathrm{NoGo})$ and $\Pr(\mathrm{Gray})$, where
$\mathbf{y}$ denotes the observed data and the sum (or integral for
continuous outcomes) is taken over all possible outcomes under the true
parameter.

For binary endpoints, the exact marginal probabilities over all possible
outcome pairs $(y_1, y_2) \in \{0,\ldots,n_1\} \times \{0,\ldots,n_2\}$ are
computed by `pbayesdecisionprob1bin()` using binomial probability weights:

$$\Pr(\mathrm{Go}) = \sum_{y_1=0}^{n_1} \sum_{y_2=0}^{n_2}
  \mathbf{1}\!\left[g_{\mathrm{Go}}(y_1, y_2) \ge \gamma_1
    \;\text{ and }\; g_{\mathrm{NoGo}}(y_1, y_2) < \gamma_2\right]\;
  \binom{n_1}{y_1} \pi_1^{y_1}(1-\pi_1)^{n_1-y_1}\;
  \binom{n_2}{y_2} \pi_2^{y_2}(1-\pi_2)^{n_2-y_2}.$$

For continuous endpoints, the operating characteristics are estimated by
Monte Carlo simulation over $n_{\mathrm{sim}}$ replicates in
`pbayesdecisionprob1cont()`: for each replicate $i$, a dataset
$(\bar{y}_1^{(i)}, s_1^{(i)}, \bar{y}_2^{(i)}, s_2^{(i)})$ is generated
from the true parameter distribution, the probability $g^{(i)}$ is
computed, and the decision is recorded. Empirical proportions over all
replicates yield the estimated operating characteristics.

### 5.3 Optimal Threshold Search

The `getgamma` family of functions finds the optimal pair
$(\gamma_1, \gamma_2)$ by grid search over a user-supplied candidate grid
$\Gamma = \{\gamma^{(1)}, \ldots, \gamma^{(K)}\} \subset (0, 1)$.
A **two-stage precompute-then-sweep** strategy is used for efficiency:

**Stage 1 (Precomputation):** all posterior/predictive probabilities
$g(y_1, y_2)$ (binary) or $g^{(i)}$ (continuous, over $n_{\mathrm{sim}}$
replicates) are computed once, independently of $\gamma$.

**Stage 2 (Gamma sweep):** for each candidate $\gamma \in \Gamma$,
$\Pr(\mathrm{Go} \mid \gamma)$ and $\Pr(\mathrm{NoGo} \mid \gamma)$ are
obtained by fast indicator summations over the precomputed values without
additional probability evaluations.

The optimal $\gamma_1$ is then selected as the smallest (or largest)
value in $\Gamma$ such that $\Pr(\mathrm{Go} \mid \gamma_1)$ satisfies a
user-specified criterion (e.g., $\Pr(\mathrm{Go}) < 0.05$ under the null
scenario), and $\gamma_2$ is selected analogously for the NoGo criterion.

---

## 6. Two-Endpoint Framework

### 6.1 Nine-Region Grid (Posterior Probability)

When two endpoints are evaluated jointly, the treatment effect vector is
$\boldsymbol{\theta} = (\theta_1, \theta_2)$.  For each endpoint $k = 1, 2$,
three zones are defined by $\theta_{\mathrm{TV}k}$ and $\theta_{\mathrm{MAV}k}$:

- **Go zone**: $\theta_k > \theta_{\mathrm{TV}k}$
- **Gray zone**: $\theta_{\mathrm{MAV}k} < \theta_k \le \theta_{\mathrm{TV}k}$
- **NoGo zone**: $\theta_k \le \theta_{\mathrm{MAV}k}$

Their Cartesian product defines nine mutually exclusive regions
$R_1, \ldots, R_9$:

```{r nine-region, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Nine-region grid for two-endpoint posterior probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="3" style="border:1px solid #aaa; padding:6px;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &gt; &theta;<sub>TV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>TV1</sub> &ge; &theta;<sub>1</sub> &gt; &theta;<sub>MAV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>MAV1</sub> &ge; &theta;<sub>1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="3" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg); font-weight:bold;">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>TV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
    <td style="border:1px solid #aaa; padding:6px;">R7</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>TV2</sub> &ge; &theta;<sub>2</sub> &gt; &theta;<sub>MAV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R5</td>
    <td style="border:1px solid #aaa; padding:6px;">R8</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>MAV2</sub> &ge; &theta;<sub>2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
    <td style="border:1px solid #aaa; padding:6px;">R6</td>
    <td style="border:1px solid #aaa; padding:6px;">R9</td>
  </tr>
</tbody>
</table>
')
```

Region $R_1$ (both endpoints exceed $\theta_{\mathrm{TV}}$) is typically
designated as **Go** and $R_9$ (both endpoints below $\theta_{\mathrm{MAV}}$)
as **NoGo**.

### 6.2 Four-Region Grid (Predictive Probability)

For the predictive probability, each endpoint has a binary success/failure
outcome, yielding four regions:

```{r four-region, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Four-region grid for two-endpoint predictive probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="2" style="border:1px solid #aaa; padding:6px;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &gt; &theta;<sub>NULL1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &le; &theta;<sub>NULL1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="2" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg); font-weight:bold;">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &le; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
  </tr>
</tbody>
</table>
')
```

### 6.3 Joint Posterior for Two Binary Endpoints

For two binary endpoints with response rates $(\pi_1^{(1)}, \pi_2^{(1)})$
(endpoint 1) and $(\pi_1^{(2)}, \pi_2^{(2)})$ (endpoint 2), a
Dirichlet-multinomial model is used. In a treatment arm of size $n_1$,
define the four mutually exclusive categories (both respond, only endpoint 1
responds, only endpoint 2 responds, neither responds) with probabilities
$\mathbf{p}_1 = (p_{11}, p_{10}, p_{01}, p_{00})$.

A Dirichlet prior $\mathrm{Dir}(\mathbf{a}_1)$ and the observed
multinomial counts $\mathbf{x}_1$ yield the conjugate posterior

$$\mathbf{p}_1 \mid \mathbf{x}_1 \;\sim\;
  \mathrm{Dir}(\mathbf{a}_1 + \mathbf{x}_1).$$

The joint probabilities for the nine regions are computed by
Monte Carlo draws from the Dirichlet posteriors in
`pbayespostpred2bin()`.

### 6.4 Joint Posterior for Two Continuous Endpoints

For two continuous endpoints, a Normal-Inverse-Wishart (N-Inv-W)
conjugate prior is placed on $(\boldsymbol{\mu}_j, \boldsymbol{\Sigma}_j)$:

$$(\boldsymbol{\mu}_j, \boldsymbol{\Sigma}_j) \;\sim\;
  \mathrm{N\text{-}Inv\text{-}W}(\mathbf{m}_{0j},\, \kappa_{0j},\,
  \nu_{0j},\, \mathbf{V}_{0j}).$$

Given $n_j$ bivariate observations, the posterior parameters are updated
analogously to the univariate case, yielding a matrix-$t$ predictive
distribution. The joint probability over rectangular regions is evaluated
using the bivariate $t$-distribution via `mvtnorm::pmvt()` for the MC
method, or a Moment-Matching approximation for the MM method (Theorem 3
of Homma and Yamaguchi, 2025).

---

## 7. Package Structure and Function Overview

```{r function-table, echo = FALSE}
knitr::kable(data.frame(
  Function = c(
    "pbayespostpred1bin()", "pbayespostpred1cont()",
    "pbayespostpred2bin()", "pbayespostpred2cont()",
    "pbayesdecisionprob1bin()", "pbayesdecisionprob1cont()",
    "pbayesdecisionprob2bin()", "pbayesdecisionprob2cont()",
    "getgamma1bin()", "getgamma1cont()",
    "getgamma2bin()", "getgamma2cont()",
    "pbetadiff()", "pbetabinomdiff()",
    "ptdiff_NI()", "ptdiff_MC()", "ptdiff_MM()",
    "rdirichlet()",
    "getjointbin()", "allmultinom()"
  ),
  Description = c(
    "Posterior/predictive probability, single binary endpoint",
    "Posterior/predictive probability, single continuous endpoint",
    "Region probabilities, two binary endpoints",
    "Region probabilities, two continuous endpoints",
    "Go/NoGo/Gray probabilities, single binary endpoint",
    "Go/NoGo/Gray probabilities, single continuous endpoint",
    "Go/NoGo/Gray probabilities, two binary endpoints",
    "Go/NoGo/Gray probabilities, two continuous endpoints",
    "Optimal Go/NoGo thresholds, single binary endpoint",
    "Optimal Go/NoGo thresholds, single continuous endpoint",
    "Optimal Go/NoGo thresholds, two binary endpoints",
    "Optimal Go/NoGo thresholds, two continuous endpoints",
    "CDF of difference of two Beta distributions",
    "Beta-binomial posterior predictive probability",
    "CDF of difference of two t-distributions (numerical integration)",
    "CDF of difference of two t-distributions (Monte Carlo)",
    "CDF of difference of two t-distributions (Moment-Matching)",
    "Random sampler for the Dirichlet distribution",
    "Joint binary probability from marginals and correlation",
    "Enumerate all multinomial outcome combinations"
  )
), col.names = c("Function", "Description"))
```

---

## 8. Quick-Start Examples

### 8.1 Posterior Probability: Single Binary Endpoint

```{r qs-bin-post}
# P(pi_treat - pi_ctrl > 0.05 | data)
# Observed: 7/10 responders (treatment), 3/10 (control)
pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.05,
  n1 = 10, n2 = 10, y1 = 7, y2 = 3,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)
```

### 8.2 Posterior Probability: Single Continuous Endpoint

```{r qs-cont-post}
# P(mu_treat - mu_ctrl > 1.0 | data) using MM method
# Observed: mean 3.5 (treatment), 1.2 (control); SD ~ 2.0
pbayespostpred1cont(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  CalcMethod = 'MM', theta0 = 1.0, nMC = NULL,
  n1 = 10, n2 = 10,
  bar.y1 = 3.5, s1 = 2.0,
  bar.y2 = 1.2, s2 = 2.0,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL
)
```

### 8.3 Go/NoGo Decision: Operating Characteristics

```{r qs-oc}
# Operating characteristics for single binary endpoint
pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'controlled',
  theta.TV  = 0.20,  theta.MAV = 0.05,  theta.NULL = NULL,
  gamma1    = 0.80,  gamma2    = 0.20,
  pi1       = c(0.15, 0.30, 0.50),
  pi2       = rep(0.15, 3),
  n1 = 10,  n2 = 10,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)
```

### 8.4 Optimal Threshold Search

```{r qs-gamma}
# Find gamma1 (smallest gamma s.t. Pr(Go) < 0.05 under null)
# and gamma2 (largest gamma s.t. Pr(NoGo) < 0.20 under null)
getgamma1bin(
  prob = 'posterior', design = 'controlled',
  theta.TV = 0.20, theta.MAV = 0.05, theta.NULL = NULL,
  pi1 = 0.15, pi2 = 0.15,
  target_go = 0.05, target_nogo = 0.20,
  crit_go = '<', crit_nogo = '<',
  sel_go = 'smallest', sel_nogo = 'largest',
  n1 = 12L, n2 = 12L,
  a1 = 0.5, a2 = 0.5, b1 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)
```

---

## 9. Further Reading

- `vignette("single-continuous")` — detailed examples for a single
  continuous endpoint, covering all three designs and calculation methods.
- `vignette("single-binary")` — detailed examples for a single binary
  endpoint, covering all three designs and prior specifications.
- `vignette("two-continuous")` — detailed examples for two continuous
  co-primary endpoints.
- `vignette("two-binary")` — detailed examples for two binary co-primary
  endpoints.
