---
title: "Single Binary Endpoint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single Binary Endpoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  fig.path = "figures/single-bin-"
)
library(BayesianQDM)
```

## Motivating Scenario

We use a hypothetical Phase IIa proof-of-concept (PoC) trial in
moderate-to-severe rheumatoid arthritis (RA). The investigational drug
is compared with placebo in a 1:1 randomised controlled trial with
$n_t = n_c = 12$ patients per group.

Endpoint: ACR20 response rate (proportion of patients achieving
$\ge 20\%$ improvement in ACR criteria).

Clinically meaningful thresholds (posterior probability):

- Target Value (TV) = 0.20 (treatment–control difference)
- Minimum Acceptable Value (MAV) = 0.05

Null hypothesis threshold (predictive probability):
$\theta_{\mathrm{NULL}} = 0.10$.

Decision thresholds: $\gamma_{\mathrm{go}} = 0.80$ (Go), $\gamma_{\mathrm{nogo}} = 0.20$ (NoGo).

Observed data (used in Sections 2–4):
$y_t = 8$ responders out of $n_t = 12$ (treatment);
$y_c = 3$ responders out of $n_c = 12$ (control).

---

## 1. Bayesian Model: Beta-Binomial Conjugate

### 1.1 Prior Distribution

Let $\pi_j$ denote the true response rate in group $j$
($j = t$: treatment, $j = c$: control). We model the number of
responders as

$$y_j \mid \pi_j \;\sim\; \mathrm{Binomial}(n_j,\, \pi_j),$$

and place independent Beta priors on each rate:

$$\pi_j \;\sim\; \mathrm{Beta}(a_j,\, b_j),$$

where $a_j > 0$ and $b_j > 0$ are the shape parameters.

The Jeffreys prior corresponds to $a_j = b_j = 0.5$, giving a
weakly informative, symmetric prior on $(0, 1)$.
The uniform prior corresponds to $a_j = b_j = 1$.

### 1.2 Posterior Distribution

By conjugacy of the Beta-Binomial model, the posterior after observing
$y_j$ responders among $n_j$ patients is

$$\pi_j \mid y_j \;\sim\; \mathrm{Beta}(a_j^*,\, b_j^*),$$

where the posterior shape parameters are

$$a_j^* = a_j + y_j, \qquad b_j^* = b_j + n_j - y_j.$$

The posterior mean and variance are

$$E[\pi_j \mid y_j] = \frac{a_j^*}{a_j^* + b_j^*}, \qquad
  \mathrm{Var}(\pi_j \mid y_j) =
  \frac{a_j^*\, b_j^*}{(a_j^* + b_j^*)^2\,(a_j^* + b_j^* + 1)}.$$

### 1.3 Posterior of the Treatment Effect

The treatment effect is $\theta = \pi_t - \pi_c$.  Since the two
groups are independent,
$\theta \mid y_t, y_c$ follows the distribution of the difference
of two independent Beta random variables:

$$\pi_t \mid y_t \;\sim\; \mathrm{Beta}(a_t^*,\, b_t^*), \qquad
  \pi_c \mid y_c \;\sim\; \mathrm{Beta}(a_c^*,\, b_c^*).$$

There is no closed-form CDF for $\theta = \pi_t - \pi_c$, so the
posterior probability

$$P(\theta > \theta_0 \mid y_t, y_c) = P(\pi_t - \pi_c > \theta_0
  \mid y_t, y_c)$$

is evaluated by the convolution integral implemented in `pbetadiff()`:

$$P(\pi_t - \pi_c > \theta_0) =
  \int_0^1
  F_{\mathrm{Beta}(a_c^*, b_c^*)}\!\left(x - \theta_0\right)\,
  f_{\mathrm{Beta}(a_t^*, b_t^*)}\!(x)\; dx,$$

where $F_{\mathrm{Beta}(\alpha,\beta)}$ is the Beta CDF and
$f_{\mathrm{Beta}(\alpha,\beta)}$ is the Beta PDF.
This one-dimensional integral is evaluated by adaptive Gauss-Kronrod
quadrature via `stats::integrate()`.

---

## 2. Posterior Predictive Probability

### 2.1 Beta-Binomial Predictive Distribution

Let $\tilde{y}_j \mid \pi_j \sim \mathrm{Binomial}(m_j, \pi_j)$ be
future trial counts with $m_j$ future patients. Integrating out $\pi_j$
over its Beta posterior gives the Beta-Binomial predictive distribution:

$$P(\tilde{y}_j = k \mid y_j) =
  \binom{m_j}{k}
  \frac{B(a_j^* + k,\; b_j^* + m_j - k)}{B(a_j^*,\; b_j^*)},
  \quad k = 0, 1, \ldots, m_j,$$

where $B(\cdot, \cdot)$ is the Beta function.

### 2.2 Posterior Predictive Probability of Future Success

The posterior predictive probability that the future treatment difference
exceeds $\theta_{\mathrm{NULL}}$ is

$$P\!\left(\frac{\tilde{y}_t}{m_t} - \frac{\tilde{y}_c}{m_c}
  > \theta_{\mathrm{NULL}} \;\Big|\; y_t, y_c\right)
  = \sum_{k_t=0}^{m_t} \sum_{k_c=0}^{m_c}
  \mathbf{1}\!\left[\frac{k_t}{m_t} - \frac{k_c}{m_c} >
  \theta_{\mathrm{NULL}}\right]\,
  P(\tilde{y}_t = k_t \mid y_t)\,
  P(\tilde{y}_c = k_c \mid y_c).$$

This double sum over all $(m_t + 1)(m_c + 1)$ outcome combinations is
evaluated exactly by `pbetabinomdiff()`.

---

## 3. Study Designs

### 3.1 Controlled Design

Standard parallel-group RCT with observed data from both groups.  The
posterior shape parameters for group $j$ are

$$a_j^* = a_j + y_j, \qquad b_j^* = b_j + n_j - y_j,$$

and the posterior probability is $P(\pi_t - \pi_c > \theta_0 \mid y_t, y_c)$
computed via `pbetadiff()`.

Posterior probability at TV and MAV:

```{r ctrl-post}
# P(pi_t - pi_c > TV  | data)
p_tv <- pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.20,
  n_t = 12, n_c = 12, y_t = 8, y_c = 3,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  m_t = NULL, m_c = NULL, z = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  lower.tail = FALSE
)

# P(pi_t - pi_c > MAV | data)
p_mav <- pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.05,
  n_t = 12, n_c = 12, y_t = 8, y_c = 3,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  m_t = NULL, m_c = NULL, z = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  lower.tail = FALSE
)

cat(sprintf("P(theta > TV  | data) = %.4f  -> Go  criterion (>= %.2f): %s\n",
            p_tv,  0.80, ifelse(p_tv  >= 0.80, "YES", "NO")))
cat(sprintf("P(theta <= MAV | data) = %.4f  -> NoGo criterion (>= %.2f): %s\n",
            1 - p_mav, 0.20, ifelse((1 - p_mav) >= 0.20, "YES", "NO")))
cat(sprintf("Decision: %s\n",
            ifelse(p_tv >= 0.80 & (1 - p_mav) < 0.20, "Go",
                   ifelse(p_tv < 0.80 & (1 - p_mav) >= 0.20, "NoGo",
                          ifelse(p_tv >= 0.80 & (1 - p_mav) >= 0.20, "Miss", "Gray")))))
```

Posterior predictive probability (future Phase III: $m_t = m_c = 40$):

```{r ctrl-pred}
p_pred <- pbayespostpred1bin(
  prob = 'predictive', design = 'controlled', theta0 = 0.10,
  n_t = 12, n_c = 12, y_t = 8, y_c = 3,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  m_t = 40, m_c = 40, z = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  lower.tail = FALSE
)
cat(sprintf("Predictive probability (m_t = m_c = 40) = %.4f\n", p_pred))
```

Vectorised computation over all possible outcomes:

The function accepts vectors for `y_t` and `y_c`, enabling efficient
computation of the posterior probability for every possible outcome pair
$(y_t, y_c) \in \{0,\ldots,n_t\} \times \{0,\ldots,n_c\}$.

```{r ctrl-vector}
grid <- expand.grid(y_t = 0:12, y_c = 0:12)
p_all <- pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.20,
  n_t = 12, n_c = 12, y_t = grid$y_t, y_c = grid$y_c,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  m_t = NULL, m_c = NULL, z = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  lower.tail = FALSE
)
# Show results for a selection of outcome pairs
sel <- data.frame(y_t = grid$y_t, y_c = grid$y_c, P_gt_TV = round(p_all, 4))
head(sel[order(-sel$P_gt_TV), ], 10)
```

### 3.2 Uncontrolled Design (Single-Arm)

When no concurrent control group is enrolled, a hypothetical control
responder count $z$ is specified from historical knowledge. The
control posterior is then

$$\pi_c \mid z \;\sim\; \mathrm{Beta}(a_c + z,\; b_c + n_c - z),$$

where $n_c$ is the hypothetical control sample size and $z$ is the
assumed number of hypothetical responders ($0 \le z \le n_c$). Only the
treatment group data $y_t$ are observed; $y_c$ is set to `NULL`.

```{r unctrl-post}
# Hypothetical control: z = 2 responders out of n_c = 12
p_unctrl <- pbayespostpred1bin(
  prob = 'posterior', design = 'uncontrolled', theta0 = 0.20,
  n_t = 12, n_c = 12, y_t = 8, y_c = NULL,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  m_t = NULL, m_c = NULL, z = 2L,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  lower.tail = FALSE
)
cat(sprintf("P(theta > TV | data, uncontrolled, z=2) = %.4f\n", p_unctrl))
```

Effect of the hypothetical control count: varying $z$ shows the
sensitivity to the assumed control rate.

```{r unctrl-z-sens}
z_seq <- 0:12
p_z   <- sapply(z_seq, function(z) {
  pbayespostpred1bin(
    prob = 'posterior', design = 'uncontrolled', theta0 = 0.20,
    n_t = 12, n_c = 12, y_t = 8, y_c = NULL,
    a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
    m_t = NULL, m_c = NULL, z = z,
    ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
    alpha0e_t = NULL, alpha0e_c = NULL,
    lower.tail = FALSE
  )
})
data.frame(z = z_seq, P_gt_TV = round(p_z, 4))
```

### 3.3 External Design (Power Prior)

Historical or external data are incorporated via a power prior with
borrowing weight $\alpha_{0ej} \in (0, 1]$. The augmented prior shape
parameters are

$$a_j^* = a_j + \alpha_{0ej}\, y_{ej}, \qquad
  b_j^* = b_j + \alpha_{0ej}\,(n_{ej} - y_{ej}),$$

where $n_{ej}$ and $y_{ej}$ are the external sample size and responder
count for group $j$
(corresponding to `ne_t`/`ne_c`, `ye_t`/`ye_c`, and `alpha0e_t`/`alpha0e_c`).
These serve as the prior for the current PoC data,
yielding a closed-form posterior:

$$\pi_j \mid y_j \;\sim\;
  \mathrm{Beta}\!\left(a_j^* + y_j,\; b_j^* + n_j - y_j\right).$$

Setting $\alpha_{0ej} = 1$ corresponds to full borrowing (the external
data are treated as if they came from the current trial); $\alpha_{0ej} \to 0$
recovers the result from the current data alone with the original prior.

Here, external data for both groups (external design):
$n_{et} = 15$, $y_{et} = 5$,
$n_{ec} = 15$, $y_{ec} = 4$, borrowing weights $\alpha_{0et} = \alpha_{0ec} = 0.5$:

```{r ext-post}
p_ext <- pbayespostpred1bin(
  prob = 'posterior', design = 'external', theta0 = 0.20,
  n_t = 12, n_c = 12, y_t = 8, y_c = 3,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  m_t = NULL, m_c = NULL, z = NULL,
  ne_t = 15L, ne_c = 15L, ye_t = 5L, ye_c = 4L,
  alpha0e_t = 0.5, alpha0e_c = 0.5,
  lower.tail = FALSE
)
cat(sprintf("P(theta > TV | data, external, alpha0e=0.5) = %.4f\n", p_ext))
```

Effect of the borrowing weight: varying $\alpha_{0ec}$ (control group
only) with fixed $\alpha_{0et} = 0.5$:

```{r ext-borrowing}
# alpha0e must be in (0, 1]; avoid alpha0e = 0
ae_seq <- c(0.01, seq(0.1, 1.0, by = 0.1))
p_ae   <- sapply(ae_seq, function(ae) {
  pbayespostpred1bin(
    prob = 'posterior', design = 'external', theta0 = 0.20,
    n_t = 12, n_c = 12, y_t = 8, y_c = 3,
    a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
    m_t = NULL, m_c = NULL, z = NULL,
    ne_t = 15L, ne_c = 15L, ye_t = 5L, ye_c = 4L,
    alpha0e_t = 0.5, alpha0e_c = ae,
    lower.tail = FALSE
  )
})
data.frame(alpha0e_c = ae_seq, P_gt_TV = round(p_ae, 4))
```

---

## 4. Operating Characteristics

### 4.1 Definition

For given true response rates $(\pi_t, \pi_c)$, the operating
characteristics are computed by exact enumeration over all
$(n_t + 1)(n_c + 1)$ possible outcome pairs
$(y_t, y_c) \in \{0,\ldots,n_t\} \times \{0,\ldots,n_c\}$:

$$\Pr(\mathrm{Go}) =
  \sum_{y_t=0}^{n_t} \sum_{y_c=0}^{n_c}
  \mathbf{1}\!\left[g_{\mathrm{Go}}(y_t, y_c) \ge \gamma_{\mathrm{go}}
    \;\text{ and }\; g_{\mathrm{NoGo}}(y_t, y_c) < \gamma_{\mathrm{nogo}}\right]\,
  \binom{n_t}{y_t} \pi_t^{y_t}(1-\pi_t)^{n_t-y_t}\,
  \binom{n_c}{y_c} \pi_c^{y_c}(1-\pi_c)^{n_c-y_c},$$

where

$$g_{\mathrm{Go}}(y_t, y_c) = P(\pi_t - \pi_c > \theta_{\mathrm{TV}} \mid y_t, y_c),$$
$$g_{\mathrm{NoGo}}(y_t, y_c) = P(\pi_t - \pi_c \le \theta_{\mathrm{MAV}} \mid y_t, y_c),$$

and the decision outcomes are classified as:

- Go: $g_{\mathrm{Go}} \ge \gamma_{\mathrm{go}}$ AND $g_{\mathrm{NoGo}} < \gamma_{\mathrm{nogo}}$
- NoGo: $g_{\mathrm{Go}} < \gamma_{\mathrm{go}}$ AND $g_{\mathrm{NoGo}} \ge \gamma_{\mathrm{nogo}}$
- Miss: $g_{\mathrm{Go}} \ge \gamma_{\mathrm{go}}$ AND $g_{\mathrm{NoGo}} \ge \gamma_{\mathrm{nogo}}$
- Gray: neither Go nor NoGo criterion met

### 4.2 Controlled Design — Multiple Scenarios

```{r oc-controlled}
oc_ctrl <- pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'controlled',
  theta_TV  = 0.20,  theta_MAV = 0.05,  theta_NULL = NULL,
  gamma_go  = 0.80,  gamma_nogo = 0.20,
  pi_t      = c(0.10, 0.20, 0.30, 0.40, 0.55, 0.70),
  pi_c      = rep(0.10, 6),
  n_t = 12,  n_c = 12,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = NULL, m_c = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_ctrl)
```

### 4.3 Figure: Go/NoGo/Gray Probabilities vs. True Treatment Effect

```{r oc-plot, fig.cap = "Operating characteristics (Go/NoGo/Gray) as a function of the true treatment effect (pi_t - pi_c). Controlled design, Jeffreys prior, exact enumeration. True pi_c = 0.10, n_t = n_c = 12, gamma_go = 0.80, gamma_nogo = 0.20."}
pi_t_seq <- seq(0.05, 0.75, by = 0.05)

oc_grid <- pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'controlled',
  theta_TV  = 0.20,  theta_MAV = 0.05,  theta_NULL = NULL,
  gamma_go  = 0.80,  gamma_nogo = 0.20,
  pi_t      = pi_t_seq,
  pi_c      = rep(0.10, length(pi_t_seq)),
  n_t = 12,  n_c = 12,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = NULL, m_c = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)

delta <- oc_grid$pi_t - 0.10   # true treatment effect

old_par <- par(mar = c(4.5, 4.5, 3, 1))

plot(delta, oc_grid$Go,
     type = "b", pch = 16, lwd = 2, col = "#2166AC",
     ylim = c(0, 1),
     xlab = expression(pi[t] - pi[c]),
     ylab = "Probability",
     main = "Operating Characteristics: Controlled Design (Posterior Prob.)",
     las  = 1)
lines(delta, oc_grid$NoGo,
      type = "b", pch = 17, lwd = 2, col = "#D6604D")
lines(delta, oc_grid$Gray,
      type = "b", pch = 15, lwd = 2, col = "#878787")

abline(h = c(0.20, 0.80), lty = 2, col = "gray50")
abline(v = c(0.05, 0.20), lty = 3, col = "gray70")

text(0.05 + 0.01, 0.04, "MAV", col = "gray40", cex = 0.8, adj = 0)
text(0.20 + 0.01, 0.04, "TV",  col = "gray40", cex = 0.8, adj = 0)

legend("right",
       legend = c("Pr(Go)", "Pr(NoGo)", "Pr(Gray)"),
       col    = c("#2166AC", "#D6604D", "#878787"),
       pch    = c(16, 17, 15), lwd = 2, bty = "n")

par(old_par)
```

### 4.4 Uncontrolled Design — Operating Characteristics

For the uncontrolled design, `pbayesdecisionprob1bin()` fixes the
hypothetical control count at $z$ and enumerates only $y_t \in \{0,\ldots,n_t\}$,
weighting by $\mathrm{Binomial}(n_t, \pi_t)$.

```{r oc-uncontrolled}
oc_unctrl <- pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'uncontrolled',
  theta_TV  = 0.20,  theta_MAV = 0.05,  theta_NULL = NULL,
  gamma_go  = 0.80,  gamma_nogo = 0.20,
  pi_t      = c(0.20, 0.30, 0.40, 0.55, 0.70),
  pi_c      = rep(0.10, 5),   # not used, but required
  n_t = 12,  n_c = 12,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = 2L,
  m_t = NULL, m_c = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_unctrl)
```

### 4.5 External Design — Operating Characteristics

```{r oc-external}
oc_ext <- pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'external',
  theta_TV  = 0.20,  theta_MAV = 0.05,  theta_NULL = NULL,
  gamma_go  = 0.80,  gamma_nogo = 0.20,
  pi_t      = c(0.20, 0.30, 0.40, 0.55, 0.70),
  pi_c      = rep(0.10, 5),
  n_t = 12,  n_c = 12,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = NULL, m_c = NULL,
  ne_t = 15L, ne_c = 15L, ye_t = 5L, ye_c = 4L,
  alpha0e_t = 0.5, alpha0e_c = 0.5,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_ext)
```

### 4.6 Predictive Probability — Operating Characteristics

```{r oc-predictive}
oc_pred <- pbayesdecisionprob1bin(
  prob      = 'predictive',
  design    = 'controlled',
  theta_TV  = NULL, theta_MAV = NULL, theta_NULL = 0.10,
  gamma_go  = 0.80, gamma_nogo = 0.20,
  pi_t      = c(0.10, 0.20, 0.30, 0.40, 0.55, 0.70),
  pi_c      = rep(0.10, 6),
  n_t = 12,  n_c = 12,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = 40L, m_c = 40L,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_pred)
```

---

## 5. Optimal Threshold Search

### 5.1 Objective and Algorithm

The `getgamma1bin()` function finds the optimal pair
$(\gamma_{\mathrm{go}}^*, \gamma_{\mathrm{nogo}}^*)$ by grid search over
$\Gamma = \{\gamma^{(1)}, \ldots, \gamma^{(K)}\} \subset (0, 1)$ at a
null scenario $(\pi_t^{\mathrm{null}}, \pi_c^{\mathrm{null}})$.

A two-stage precompute-then-sweep strategy is used:

Stage 1 (Precompute): all posterior/predictive probabilities
$g_{\mathrm{Go}}(y_t, y_c)$ and $g_{\mathrm{NoGo}}(y_t, y_c)$ are
computed once over all $(n_t + 1)(n_c + 1)$ outcome pairs — independently
of $\gamma$.

Stage 2 (Sweep): for each candidate $\gamma \in \Gamma$,
$\Pr(\mathrm{Go} \mid \gamma)$ and $\Pr(\mathrm{NoGo} \mid \gamma)$
are obtained as weighted sums of indicator functions:

$$\Pr(\mathrm{Go} \mid \gamma) = \sum_{y_t, y_c}
  \mathbf{1}\!\left[g_{\mathrm{Go}}(y_t, y_c) \ge \gamma\right]\,
  w(y_t, y_c),$$

$$\Pr(\mathrm{NoGo} \mid \gamma) = \sum_{y_t, y_c}
  \mathbf{1}\!\left[g_{\mathrm{NoGo}}(y_t, y_c) \ge \gamma\right]\,
  w(y_t, y_c),$$

where $w(y_t, y_c) = \mathrm{Binom}(y_t; n_t, \pi_t)\,\mathrm{Binom}(y_c; n_c, \pi_c)$
are the binomial weights under the null scenario.  No further calls to
`pbayespostpred1bin()` are required at this stage.

The sweep is vectorised using `outer()`, making the full grid search
computationally negligible once Stage 1 is complete.

The optimal $\gamma_{\mathrm{go}}^*$ is the smallest (or largest, depending on
`sel_go`) value in $\Gamma$ such that
$\Pr(\mathrm{Go} \mid \gamma_{\mathrm{go}}^*)$ satisfies the criterion against a
target (e.g., $\Pr(\mathrm{Go}) < 0.05$). Analogously for $\gamma_{\mathrm{nogo}}^*$.

### 5.2 Example: Controlled Design, Posterior Probability

```{r getgamma-ctrl}
res_gamma <- getgamma1bin(
  prob      = 'posterior',
  design    = 'controlled',
  theta_TV  = 0.20, theta_MAV = 0.05, theta_NULL = NULL,
  pi_t = 0.10, pi_c = 0.10,          # null scenario: no treatment effect
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n_t = 12L, n_c = 12L,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = NULL, m_c = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  gamma_grid = seq(0.01, 0.99, by = 0.01)
)
print(res_gamma)
```

### 5.3 Figure: Pr(Go) and Pr(NoGo) Over the Gamma Grid

```{r gamma-plot, fig.cap = "Pr(Go) and Pr(NoGo) as a function of the candidate threshold gamma under the null scenario (pi_t = pi_c = 0.10). The optimal gamma_go (blue dashed line) is the smallest gamma satisfying Pr(Go) < 0.05. The optimal gamma_nogo (red dashed line) is the largest gamma satisfying Pr(NoGo) < 0.20."}
grid_df <- data.frame(
  gamma  = res_gamma$gamma_grid,
  PrGo   = res_gamma$PrGo_grid,
  PrNoGo = res_gamma$PrNoGo_grid
)

old_par2 <- par(mar = c(4.5, 4.5, 3, 1))

plot(grid_df$gamma, grid_df$PrGo,
     type = "l", lwd = 2, col = "#2166AC",
     ylim = c(0, 1),
     xlab = expression(gamma),
     ylab = "Probability",
     main = "Optimal Threshold Search: Pr(Go) and Pr(NoGo) vs. Gamma",
     las  = 1)
lines(grid_df$gamma, grid_df$PrNoGo,
      lwd = 2, col = "#D6604D")

abline(h = 0.05, lty = 2, col = "#2166AC", lwd = 1.2)
abline(h = 0.20, lty = 2, col = "#D6604D", lwd = 1.2)

if (!is.na(res_gamma$gamma_go)) {
  abline(v = res_gamma$gamma_go, lty = 2, col = "#2166AC", lwd = 2)
  text(res_gamma$gamma_go + 0.02, 0.88,
       paste0("gamma_go* = ", res_gamma$gamma_go),
       col = "#2166AC", cex = 0.85, adj = 0)
}
if (!is.na(res_gamma$gamma_nogo)) {
  abline(v = res_gamma$gamma_nogo, lty = 2, col = "#D6604D", lwd = 2)
  text(res_gamma$gamma_nogo - 0.02, 0.73,
       paste0("gamma_nogo* = ", res_gamma$gamma_nogo),
       col = "#D6604D", cex = 0.85, adj = 1)
}

legend("topright",
       legend = c("Pr(Go | gamma)", "Pr(NoGo | gamma)",
                  "Target Go = 0.05", "Target NoGo = 0.20"),
       col    = c("#2166AC", "#D6604D", "#2166AC", "#D6604D"),
       lty    = c(1, 1, 2, 2),
       lwd    = c(2, 2, 1.2, 1.2),
       bty    = "n", cex = 0.85)

par(old_par2)
```

### 5.4 Example: Uncontrolled Design, Posterior Probability

```{r getgamma-unctrl}
res_gamma_unctrl <- getgamma1bin(
  prob      = 'posterior',
  design    = 'uncontrolled',
  theta_TV  = 0.20, theta_MAV = 0.05, theta_NULL = NULL,
  pi_t = 0.10, pi_c = NULL,
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n_t = 12L, n_c = 12L,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = 2L, m_t = NULL, m_c = NULL,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  gamma_grid = seq(0.01, 0.99, by = 0.01)
)
print(res_gamma_unctrl)
```

### 5.5 Example: External Design, Posterior Probability

```{r getgamma-ext}
res_gamma_ext <- getgamma1bin(
  prob      = 'posterior',
  design    = 'external',
  theta_TV  = 0.20, theta_MAV = 0.05, theta_NULL = NULL,
  pi_t = 0.10, pi_c = 0.10,
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n_t = 12L, n_c = 12L,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = NULL, m_c = NULL,
  ne_t = 15L, ne_c = 15L, ye_t = 5L, ye_c = 4L,
  alpha0e_t = 0.5, alpha0e_c = 0.5,
  gamma_grid = seq(0.01, 0.99, by = 0.01)
)
print(res_gamma_ext)
```

### 5.6 Example: Controlled Design, Predictive Probability

```{r getgamma-pred}
res_gamma_pred <- getgamma1bin(
  prob      = 'predictive',
  design    = 'controlled',
  theta_TV  = NULL, theta_MAV = NULL, theta_NULL = 0.10,
  pi_t = 0.10, pi_c = 0.10,
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n_t = 12L, n_c = 12L,
  a_t = 0.5, b_t = 0.5, a_c = 0.5, b_c = 0.5,
  z = NULL, m_t = 40L, m_c = 40L,
  ne_t = NULL, ne_c = NULL, ye_t = NULL, ye_c = NULL,
  alpha0e_t = NULL, alpha0e_c = NULL,
  gamma_grid = seq(0.01, 0.99, by = 0.01)
)
print(res_gamma_pred)
```

---

## 6. Summary

This vignette covered single binary endpoint analysis in BayesianQDM:

1. Bayesian model: Beta-Binomial conjugate posterior with explicit
   update formulae for $a_j^* = a_j + y_j$ and $b_j^* = b_j + n_j - y_j$;
   Jeffreys prior ($a_j = b_j = 0.5$) as the recommended weakly
   informative default.
2. Posterior probability: evaluated via the convolution integral
   in `pbetadiff()` — no closed-form CDF exists for the Beta difference.
3. Posterior predictive probability: evaluated by exact enumeration
   of the Beta-Binomial double sum in `pbetabinomdiff()`.
4. Three study designs: controlled (both groups observed), uncontrolled
   (hypothetical control count $z$ fixed from prior knowledge), external
   design (power prior with borrowing weight $\alpha_{0ej}$, augmenting
   the prior by $\alpha_{0ej}$ times the external sufficient statistics).
5. Operating characteristics: computed by exact enumeration over all
   $(n_t+1)(n_c+1)$ outcome pairs with binomial probability weights —
   no simulation required for the binary endpoint.
6. Optimal threshold search: two-stage precompute-then-sweep strategy
   in `getgamma1bin()` using `outer()` for a fully vectorised gamma sweep.

See `vignette("single-continuous")` for the analogous continuous endpoint
analysis.
