---
title: "Single-Endpoint Bayesian QDM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single-Endpoint Bayesian QDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(BayesianQDM)
```

## Motivating Scenario

We use a hypothetical Phase IIa proof-of-concept (PoC) trial in
moderate-to-severe rheumatoid arthritis (RA).  The investigational drug is
compared with placebo in a 1:1 randomised controlled trial with 10 patients per
arm.

* **Binary endpoint** — ACR20 response rate (proportion of patients with
  $\ge$ 20% improvement).
* **Continuous endpoint** — Change from baseline in DAS28 score (larger
  decrease = better).

Clinically meaningful thresholds:

```{r thresholds, echo = FALSE}
knitr::kable(data.frame(
  Endpoint  = c("ACR20 rate difference", "DAS28 change difference"),
  TV        = c("0.20", "1.50"),
  MAV       = c("0.05", "0.50"),
  theta_NULL = c("0.10", "1.00")
), col.names = c("Endpoint", "TV (Target Value)", "MAV (Min. Acceptable)",
                 "theta_NULL (Predictive)"))
```

Decision thresholds: $\gamma_1 = 0.80$ (Go), $\gamma_2 = 0.20$ (NoGo).

---

## Part 1: Binary Endpoint (ACR20 Response Rate)

### 1-1. Posterior Probability — Controlled Design

Suppose 7 of 10 treatment patients and 3 of 10 control patients achieved ACR20.

```{r bin-post-ctrl}
# P(pi_treat - pi_ctrl > 0.05 | data)  [MAV threshold]
p_mav <- pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.05,
  n1 = 10, n2 = 10, y1 = 7, y2 = 3,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)

# P(pi_treat - pi_ctrl > 0.20 | data)  [TV threshold]
p_tv <- pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.20,
  n1 = 10, n2 = 10, y1 = 7, y2 = 3,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)

cat(sprintf("P(effect > MAV | data) = %.4f\n", p_mav))
cat(sprintf("P(effect > TV  | data) = %.4f\n", p_tv))
cat(sprintf("Decision: %s\n",
            ifelse(p_tv >= 0.80, "Go",
                   ifelse(p_mav <= 0.20, "NoGo", "Gray"))))
```

### 1-2. Posterior Predictive Probability — Controlled Design

Assuming a future Phase III trial with 50 patients per arm:

```{r bin-pred-ctrl}
p_pred <- pbayespostpred1bin(
  prob = 'predictive', design = 'controlled', theta0 = 0.10,
  n1 = 10, n2 = 10, y1 = 7, y2 = 3,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = 50, m2 = 50, z = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)
cat(sprintf("Predictive probability = %.4f\n", p_pred))
```

### 1-3. Uncontrolled Design (Single-Arm with Hypothetical Control)

When no concurrent control arm is available, a hypothetical control responder
count `z` is specified from historical data.

```{r bin-post-unctrl}
# Hypothetical control: 2/10 responders (z = 2)
p_unctrl <- pbayespostpred1bin(
  prob = 'posterior', design = 'uncontrolled', theta0 = 0.05,
  n1 = 10, n2 = 10, y1 = 7, y2 = NULL,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = 2L,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)
cat(sprintf("P(effect > MAV | data, uncontrolled) = %.4f\n", p_unctrl))
```

### 1-4. External Control Design (Power Prior)

Suppose historical data are available for both arms: 5 responders in 15
treatment patients and 4 responders in 15 control patients, with a borrowing
weight of 0.5.

```{r bin-post-ext}
p_ext <- pbayespostpred1bin(
  prob = 'posterior', design = 'external', theta0 = 0.05,
  n1 = 10, n2 = 10, y1 = 7, y2 = 3,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = NULL,
  ne1 = 15L, ne2 = 15L, ye1 = 5L, ye2 = 4L, ae1 = 0.5, ae2 = 0.5,
  lower.tail = FALSE
)
cat(sprintf("P(effect > MAV | data, external) = %.4f\n", p_ext))
```

### 1-5. Operating Characteristics — Controlled Design

```{r bin-oc}
oc_bin <- pbayesdecisionprob1bin(
  prob      = 'posterior',
  design    = 'controlled',
  theta.TV  = 0.20,  theta.MAV  = 0.05,  theta.NULL = NULL,
  gamma1    = 0.80,  gamma2     = 0.20,
  pi1       = c(0.15, 0.25, 0.35, 0.50, 0.65),
  pi2       = rep(0.15, 5),
  n1 = 10,  n2 = 10,
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  z = NULL, m1 = NULL, m2 = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL
)
print(oc_bin)
```

### 1-6. Vectorised Posterior Probability Over All Outcomes

`pbayespostpred1bin` accepts vectors for `y1` and `y2`, enabling efficient
computation over all possible observed outcomes.

```{r bin-vector}
# All outcomes where y1 = 0..10 and y2 = 3 (fixed control count)
y1_all <- 0:10
p_all  <- pbayespostpred1bin(
  prob = 'posterior', design = 'controlled', theta0 = 0.05,
  n1 = 10, n2 = 10, y1 = y1_all, y2 = rep(3, 11),
  a1 = 0.5, b1 = 0.5, a2 = 0.5, b2 = 0.5,
  m1 = NULL, m2 = NULL, z = NULL,
  ne1 = NULL, ne2 = NULL, ye1 = NULL, ye2 = NULL, ae1 = NULL, ae2 = NULL,
  lower.tail = FALSE
)
data.frame(y1 = y1_all, P_gt_MAV = round(p_all, 4))
```

---

## Part 2: Continuous Endpoint (DAS28 Change from Baseline)

### 2-1. Posterior Probability — Three Calculation Methods

The same posterior probability can be computed via `'NI'`, `'MC'`, and `'MM'`.
For operating characteristics evaluation the `'MM'` method is recommended.

```{r cont-post-methods}
# Observed: mean change 3.5 (treatment), 1.2 (control); SD ~ 2.0
args_cont <- list(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  theta0 = 1.00,
  n1 = 10, n2 = 10,
  bar.y1 = 3.5, s1 = 2.0,
  bar.y2 = 1.2, s2 = 2.0,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL
)

p_ni <- do.call(pbayespostpred1cont,
                c(args_cont, list(CalcMethod = 'NI', nMC = NULL)))
p_mm <- do.call(pbayespostpred1cont,
                c(args_cont, list(CalcMethod = 'MM', nMC = NULL)))

set.seed(1)
p_mc <- do.call(pbayespostpred1cont,
                c(args_cont, list(CalcMethod = 'MC', nMC = 5000L)))

cat(sprintf("NI : %.4f\n", p_ni))
cat(sprintf("MM : %.4f\n", p_mm))
cat(sprintf("MC : %.4f\n", p_mc))
```

### 2-2. Informative Prior (Normal-Inverse-Chi-Squared)

Historical data suggest mean DAS28 change ~1.0 in control patients
($\kappa_{02} = 5$, $\nu_{02} = 5$, $\mu_{02} = 1.0$, $\sigma_{02} = 2.0$).

```{r cont-post-niChisq}
p_ni_chisq <- pbayespostpred1cont(
  prob = 'posterior', design = 'controlled', prior = 'N-Inv-Chisq',
  CalcMethod = 'NI', theta0 = 1.00, nMC = NULL,
  n1 = 10, n2 = 10,
  bar.y1 = 3.5, s1 = 2.0,
  bar.y2 = 1.2, s2 = 2.0,
  m1 = NULL, m2 = NULL,
  kappa01 = 5,   kappa02 = 5,
  nu01    = 5,   nu02    = 5,
  mu01    = 3.0, mu02    = 1.0,
  sigma01 = 2.0, sigma02 = 2.0,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL
)
cat(sprintf("P(effect > 1.0 | data, N-Inv-Chisq prior) = %.4f\n", p_ni_chisq))
```

### 2-3. Posterior Predictive Probability

```{r cont-pred}
p_pred_cont <- pbayespostpred1cont(
  prob = 'predictive', design = 'controlled', prior = 'vague',
  CalcMethod = 'NI', theta0 = 1.00, nMC = NULL,
  n1 = 10, n2 = 10, m1 = 60, m2 = 60,
  bar.y1 = 3.5, s1 = 2.0,
  bar.y2 = 1.2, s2 = 2.0,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL
)
cat(sprintf("Predictive probability (m1 = m2 = 60) = %.4f\n", p_pred_cont))
```

### 2-4. External Control Design (Power Prior)

Historical control data: mean change 0.9, SD 2.0, n = 20 patients,
borrowing weight $\alpha_{02} = 0.5$.

```{r cont-ext}
p_ext_cont <- pbayespostpred1cont(
  prob = 'posterior', design = 'external', prior = 'vague',
  CalcMethod = 'MM', theta0 = 1.00, nMC = NULL,
  n1 = 10, n2 = 10,
  bar.y1 = 3.5, s1 = 2.0,
  bar.y2 = 1.2, s2 = 2.0,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = 20L, alpha01 = NULL, alpha02 = 0.5,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = 0.9, se2 = 2.0
)
cat(sprintf("P(effect > 1.0 | data, external control) = %.4f\n", p_ext_cont))
```

### 2-5. Operating Characteristics — Controlled Design (MM method)

`sigma1` and `sigma2` are scalars (common SD for all scenarios).
`mu1` accepts a vector of true treatment means (one per scenario).

```{r cont-oc}
oc_cont <- pbayesdecisionprob1cont(
  nsim       = 200L,
  prob       = 'posterior',
  design     = 'controlled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.50,  theta.MAV  = 0.50,  theta.NULL = NULL,
  nMC        = NULL,
  gamma1     = 0.80,  gamma2     = 0.20,
  n1 = 10,   n2 = 10,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  mu1    = c(1.5, 2.5, 3.5, 4.5),
  mu2    = 1.0,
  sigma1 = 2.0,
  sigma2 = 2.0,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE,
  seed = 42L
)
print(oc_cont)
```

### 2-6. Uncontrolled Design (Single-Arm)

When only a single arm is enrolled, the hypothetical control mean `mu02` and
variance ratio `r` are specified from historical knowledge.

```{r cont-uctrl}
oc_unctrl <- pbayesdecisionprob1cont(
  nsim       = 200L,
  prob       = 'posterior',
  design     = 'uncontrolled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.50,  theta.MAV  = 0.50,  theta.NULL = NULL,
  nMC        = NULL,
  gamma1     = 0.80,  gamma2     = 0.20,
  n1 = 10,   n2 = 10,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01    = NULL,  mu02   = 1.0,
  sigma01 = NULL,  sigma02 = NULL,
  mu1    = c(2.5, 3.5, 4.5),
  mu2    = 1.0,
  sigma1 = 2.0,
  sigma2 = 2.0,
  r      = 1.0,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE,
  seed = 1L
)
print(oc_unctrl)
```

---

## Summary

This vignette demonstrated:

1. **Posterior and posterior predictive probability** calculation for binary
   and continuous endpoints (`pbayespostpred1bin`, `pbayespostpred1cont`).
2. **Three calculation methods** for continuous endpoints (`'NI'`, `'MC'`,
   `'MM'`).
3. **Three study designs**: controlled, uncontrolled, and external control.
4. **Informative priors**: Beta conjugate (binary) and Normal-Inverse-Chi-squared
   (continuous).
5. **Operating characteristics** evaluation via `pbayesdecisionprob1bin` and
   `pbayesdecisionprob1cont`.

See `vignette("two-endpoints")` for analyses involving two co-primary
endpoints.
