---
title: "Single Continuous Endpoint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single Continuous Endpoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  fig.path = "figures/single-cont-"
)
library(BayesianQDM)
```

## Motivating Scenario

We use a hypothetical Phase IIa proof-of-concept (PoC) trial in
moderate-to-severe rheumatoid arthritis (RA). The investigational drug
is compared with placebo in a 1:1 randomised controlled trial with
$n_1 = n_2 = 15$ patients per arm.

**Endpoint**: change from baseline in DAS28 score (a larger decrease
indicates greater improvement).

**Clinically meaningful thresholds** (posterior probability):

- Target Value ($\theta_{\mathrm{TV}}$) = 1.5 (ambitious target)
- Minimum Acceptable Value ($\theta_{\mathrm{MAV}}$) = 0.5 (lower bar)

**Null hypothesis threshold** (predictive probability):
$\theta_{\mathrm{NULL}} = 1.0$.

**Decision thresholds**: $\gamma_1 = 0.80$ (Go), $\gamma_2 = 0.20$ (NoGo).

**Observed data** (used in Sections 2–4):
$\bar{y}_1 = 3.2$, $s_1 = 2.0$ (treatment);
$\bar{y}_2 = 1.1$, $s_2 = 1.8$ (control).

---

## 1. Bayesian Model: Normal-Inverse-Chi-Squared Conjugate

### 1.1 Prior Distribution

For group $j$ ($j = 1$: treatment, $j = 2$: control), we model the
continuous outcome as

$$y_{ji} \mid \mu_j, \sigma_j^2 \;\overset{iid}{\sim}\;
  \mathcal{N}(\mu_j,\, \sigma_j^2).$$

The conjugate prior is the Normal-Inverse-Chi-squared (N-Inv-$\chi^2$)
distribution:

$$(\mu_j,\, \sigma_j^2) \;\sim\;
  \mathrm{N\text{-}Inv\text{-}\chi^2}
  \!\left(\mu_{0j},\, \kappa_{0j},\, \nu_{0j},\, \sigma_{0j}^2\right),$$

where $\mu_{0j}$ is the prior mean, $\kappa_{0j} > 0$ the prior precision
(pseudo-sample size), $\nu_{0j} > 0$ the prior degrees of freedom, and
$\sigma_{0j}^2 > 0$ the prior scale.

**Vague (Jeffreys) prior** is obtained by letting
$\kappa_{0j} \to 0$ and $\nu_{0j} \to 0$, which places minimal information
on the parameters.

### 1.2 Posterior Distribution

Given $n_j$ observations with sample mean $\bar{y}_j$ and sample standard
deviation $s_j$, the posterior parameters are updated as follows:

$$\kappa_{nj} = \kappa_{0j} + n_j, \qquad
  \nu_{nj} = \nu_{0j} + n_j,$$

$$\mu_{nj} = \frac{\kappa_{0j}\,\mu_{0j} + n_j\,\bar{y}_j}{\kappa_{nj}},$$

$$\sigma_{nj}^2 = \frac{\nu_{0j}\,\sigma_{0j}^2 + (n_j - 1)\,s_j^2 +
  \dfrac{n_j\,\kappa_{0j}}{\kappa_{nj}}\,(\mu_{0j} - \bar{y}_j)^2}
  {\nu_{nj}}.$$

The marginal posterior of the group mean $\mu_j$ is a non-standardised
$t$-distribution:

$$\mu_j \mid \mathbf{y}_j \;\sim\;
  t_{\nu_{nj}}\!\!\left(\mu_{nj},\;
  \frac{\sigma_{nj}}{\sqrt{\kappa_{nj}}}\right).$$

Under the **vague prior** ($\kappa_{0j} = \nu_{0j} = 0$), this simplifies to

$$\mu_j \mid \mathbf{y}_j \;\sim\;
  t_{n_j - 1}\!\!\left(\bar{y}_j,\;
  \frac{s_j}{\sqrt{n_j}}\right).$$

### 1.3 Posterior of the Treatment Effect

The treatment effect is $\theta = \mu_1 - \mu_2$.  Since the two groups
are independent, the posterior of $\theta$ is the distribution of the
difference of two independent non-standardised $t$-variables:

$$T_j \;\sim\; t_{\nu_{nj}}(\mu_{nj},\, \sigma_{nj}^{\dagger}), \qquad
  j = 1, 2,$$

where the posterior scale is
$\sigma_{nj}^{\dagger} = \sigma_{nj} / \sqrt{\kappa_{nj}}$.

The posterior probability that $\theta$ exceeds a threshold $\theta_0$,

$$P(\theta > \theta_0 \mid \mathbf{y}) =
  P(T_1 - T_2 > \theta_0),$$

is computed by one of three methods described in Section 3.

---

## 2. Posterior Predictive Probability

### 2.1 Predictive Distribution

Let $\bar{\tilde{y}}_j$ denote the future sample mean based on $m_j$
future patients.  Under the N-Inv-$\chi^2$ posterior, the predictive
distribution of $\bar{\tilde{y}}_j$ is again a non-standardised
$t$-distribution:

$$\bar{\tilde{y}}_j \mid \mathbf{y}_j \;\sim\;
  t_{\nu_{nj}}\!\!\left(\mu_{nj},\;
  \sigma_{nj}\sqrt{\frac{1 + \kappa_{nj}}{\kappa_{nj}\, m_j}}\right).$$

Under the vague prior ($\kappa_{nj} = n_j$, $\sigma_{nj} = s_j$) this becomes

$$\bar{\tilde{y}}_j \mid \mathbf{y}_j \;\sim\;
  t_{n_j - 1}\!\!\left(\bar{y}_j,\;
  s_j\sqrt{\frac{n_j + 1}{n_j\, m_j}}\right).$$

### 2.2 Posterior Predictive Probability

The posterior predictive probability that the future treatment effect
exceeds the null threshold $\theta_{\mathrm{NULL}}$ is

$$P\!\left(\bar{\tilde{y}}_1 - \bar{\tilde{y}}_2 > \theta_{\mathrm{NULL}}
  \mid \mathbf{y}\right),$$

computed by the same three methods as the posterior probability but with
the predictive scale replacing the posterior scale.

---

## 3. Three Computation Methods

### 3.1 Numerical Integration (NI)

The exact CDF of $D = T_1 - T_2$ is expressed as a one-dimensional integral:

$$P(D \le \theta_0) =
  \int_{-\infty}^{\infty}
  F_{t_{\nu_2}(\mu_2, \sigma_2)}\!\left(t - \theta_0\right)\,
  f_{t_{\nu_1}(\mu_1, \sigma_1)}(t)\; dt,$$

where $F_{t_\nu(\mu,\sigma)}$ and $f_{t_\nu(\mu,\sigma)}$ are the CDF and
PDF of the non-standardised $t$-distribution respectively. This integral is
evaluated by `stats::integrate()` (adaptive Gauss-Kronrod quadrature) in
`ptdiff_NI()`. The method is exact but evaluates the integral once per call
and is therefore slow for large-scale simulation.

### 3.2 Monte Carlo Simulation (MC)

$n_{\mathrm{MC}}$ independent samples are drawn from each marginal:

$$T_1^{(i)} \;\sim\; t_{\nu_1}(\mu_1, \sigma_1), \quad
  T_2^{(i)} \;\sim\; t_{\nu_2}(\mu_2, \sigma_2), \quad
  i = 1, \ldots, n_{\mathrm{MC}},$$

and the probability is estimated as

$$\widehat{P}(D > \theta_0) =
  \frac{1}{n_{\mathrm{MC}}}
  \sum_{i=1}^{n_{\mathrm{MC}}}
  \mathbf{1}\!\left[T_1^{(i)} - T_2^{(i)} > \theta_0\right].$$

When called from `pbayesdecisionprob1cont()` with $n_{\mathrm{sim}}$
simulation replicates, all draws are generated as a single
$n_{\mathrm{MC}} \times n_{\mathrm{sim}}$ matrix to avoid loop overhead.

### 3.3 Moment-Matching Approximation (MM)

The difference $D = T_1 - T_2$ is approximated by a single
non-standardised $t$-distribution
$D \approx t_{\nu^*}(\mu^*, \sigma^*)$ by matching its first two even
moments.

**Step 1**: The mean of $D$ is

$$\mu^* = \mu_1 - \mu_2.$$

**Step 2**: The variance of each $T_j$ is

$$\mathrm{Var}(T_j) = \sigma_j^2 \cdot \frac{\nu_j}{\nu_j - 2}
  \qquad (\nu_j > 2),$$

so the variance of $D$ is

$$V = \mathrm{Var}(D) = \sigma_1^2\,\frac{\nu_1}{\nu_1 - 2} +
  \sigma_2^2\,\frac{\nu_2}{\nu_2 - 2}.$$

**Step 3**: The effective degrees of freedom $\nu^*$ are obtained from
the fourth-moment equation. The fourth central moment of
$T_j \sim t_{\nu_j}(\mu_j, \sigma_j)$ is

$$\mu_4(T_j) = 3\,\sigma_j^4\,\frac{\nu_j^2}{(\nu_j-2)(\nu_j-4)}
  \qquad (\nu_j > 4).$$

Matching the fourth moment of $D$ to that of $t_{\nu^*}(\mu^*, \sigma^*)$
yields

$$\nu^* =
  \frac{2V^2 - 4Q}{V^2 - Q},$$

where

$$Q = \sigma_1^4\,\frac{\nu_1^2}{(\nu_1-2)(\nu_1-4)} +
      \sigma_2^4\,\frac{\nu_2^2}{(\nu_2-2)(\nu_2-4)} +
      2\,\mathrm{Var}(T_1)\,\mathrm{Var}(T_2).$$

**Step 4**: The scale $\sigma^*$ is determined from the second-moment
equation:

$$\sigma^* = \sqrt{V \cdot \frac{\nu^* - 2}{\nu^*}}.$$

**Step 5**: The CDF is evaluated as

$$P(D > \theta_0) = P\!\left(Z > \frac{\theta_0 - \mu^*}{\sigma^*}\right),
  \quad Z \sim t_{\nu^*}(0, 1),$$

via a single call to `stats::pt()`. This method requires $\nu_j > 4$,
is exact in the normal limit ($\nu_j \to \infty$), and is fully vectorised
— making it the recommended method for large-scale simulation.

### 3.4 Comparison of the Three Methods

```{r method-compare}
# Observed data (nMC excluded from base list; passed individually per method)
args_base <- list(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  theta0 = 1.0,
  n1 = 15, n2 = 15,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = 1.1, s2 = 1.8,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL
)

p_ni <- do.call(pbayespostpred1cont,
                c(args_base, list(CalcMethod = 'NI', nMC = NULL)))
p_mm <- do.call(pbayespostpred1cont,
                c(args_base, list(CalcMethod = 'MM', nMC = NULL)))
set.seed(42)
p_mc <- do.call(pbayespostpred1cont,
                c(args_base, list(CalcMethod = 'MC', nMC = 10000L)))

cat(sprintf("NI : %.6f  (exact)\n",      p_ni))
cat(sprintf("MM : %.6f  (approx)\n",     p_mm))
cat(sprintf("MC : %.6f  (n_MC=10000)\n", p_mc))
```

---

## 4. Study Designs

### 4.1 Controlled Design

Standard parallel-group RCT with observed data from both treatment
($n_1$, $\bar{y}_1$, $s_1$) and control ($n_2$, $\bar{y}_2$, $s_2$) arms.
Both posteriors are updated from the PoC data.

**Vague prior — posterior probability**:

```{r ctrl-post}
# P(mu1 - mu2 > theta_TV | data) and P(mu1 - mu2 <= theta_MAV | data)
p_tv <- pbayespostpred1cont(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  CalcMethod = 'NI', theta0 = 1.5, nMC = NULL,
  n1 = 15, n2 = 15,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = 1.1, s2 = 1.8,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  lower.tail = FALSE
)

p_mav <- pbayespostpred1cont(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  CalcMethod = 'NI', theta0 = 0.5, nMC = NULL,
  n1 = 15, n2 = 15,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = 1.1, s2 = 1.8,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  lower.tail = FALSE
)

cat(sprintf("P(theta > theta_TV  | data) = %.4f  -> Go  criterion: >= 0.80? %s\n",
            p_tv, ifelse(p_tv >= 0.80, "YES", "NO")))
cat(sprintf("P(theta <= theta_MAV | data) = %.4f  -> NoGo criterion: >= 0.20? %s\n",
            1 - p_mav, ifelse((1 - p_mav) >= 0.20, "YES", "NO")))
cat(sprintf("Decision: %s\n",
            ifelse(p_tv >= 0.80 & (1 - p_mav) < 0.20, "Go",
                   ifelse(p_tv < 0.80 & (1 - p_mav) >= 0.20, "NoGo",
                          ifelse(p_tv >= 0.80 & (1 - p_mav) >= 0.20, "Miss", "Gray")))))
```

**N-Inv-$\chi^2$ informative prior**:

Historical knowledge suggests treatment mean $\sim 3.0$ and control mean
$\sim 1.0$ with prior pseudo-sample size $\kappa_{0j} = 5$ and
degrees of freedom $\nu_{0j} = 5$.

```{r ctrl-post-niChisq}
p_tv_inf <- pbayespostpred1cont(
  prob = 'posterior', design = 'controlled', prior = 'N-Inv-Chisq',
  CalcMethod = 'NI', theta0 = 1.5, nMC = NULL,
  n1 = 15, n2 = 15,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = 1.1, s2 = 1.8,
  m1 = NULL, m2 = NULL,
  kappa01 = 5, kappa02 = 5,
  nu01    = 5, nu02    = 5,
  mu01    = 3.0, mu02  = 1.0,
  sigma01 = 2.0, sigma02 = 1.8,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  lower.tail = FALSE
)
cat(sprintf("P(theta > theta_TV | data, N-Inv-Chisq prior) = %.4f\n", p_tv_inf))
```

**Posterior predictive probability** (future Phase III: $m_1 = m_2 = 60$):

```{r ctrl-pred}
p_pred <- pbayespostpred1cont(
  prob = 'predictive', design = 'controlled', prior = 'vague',
  CalcMethod = 'NI', theta0 = 1.0, nMC = NULL,
  n1 = 15, n2 = 15, m1 = 60, m2 = 60,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = 1.1, s2 = 1.8,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  lower.tail = FALSE
)
cat(sprintf("Predictive probability (m1 = m2 = 60) = %.4f\n", p_pred))
```

### 4.2 Uncontrolled Design (Single-Arm)

When no concurrent control arm is enrolled, the control distribution is
specified from external knowledge:

- Hypothetical control mean: $\mu_{02}$ (fixed scalar)
- Variance scaling: $\sigma_{\mathrm{control}} = \sqrt{r} \cdot \sigma_{\mathrm{treatment}}$,
  so $r = 1$ assumes equal variances

The posterior of the control mean is not updated from observed data;
only the treatment posterior is updated.

```{r unctrl-post}
p_unctrl <- pbayespostpred1cont(
  prob = 'posterior', design = 'uncontrolled', prior = 'vague',
  CalcMethod = 'MM', theta0 = 1.5, nMC = NULL,
  n1 = 15, n2 = 15,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = NULL, s2 = NULL,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = 1.0,   # hypothetical control mean
  sigma01 = NULL, sigma02 = NULL,
  r = 1.0,                   # equal variance assumption
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  lower.tail = FALSE
)
cat(sprintf("P(theta > theta_TV | data, uncontrolled) = %.4f\n", p_unctrl))
```

### 4.3 External Control Design (Power Prior)

Historical external data from a prior study are incorporated via a power
prior with borrowing weight $\alpha_{e2} \in (0, 1]$. The augmented
posterior parameters for the control group are:

$$\mu_{n2} = \frac{\alpha_{e2} n_{e2}\, \bar{y}_{e2} + n_2\, \bar{y}_2}
  {\alpha_{e2} n_{e2} + n_2},$$

$$\sigma_{n2}^2 = \frac{\alpha_{e2}(n_{e2}-1) s_{e2}^2 + (n_2-1) s_2^2 +
  \dfrac{\alpha_{e2} n_{e2} n_2 (\bar{y}_{e2} - \bar{y}_2)^2}
        {\alpha_{e2} n_{e2} + n_2}}
  {\alpha_{e2} n_{e2} + n_2},$$

with degrees of freedom $\nu_{n2} = \alpha_{e2} n_{e2} + n_2 - 1$.

Here, $n_{e2} = 20$, $\bar{y}_{e2} = 0.9$, $s_{e2} = 1.8$,
$\alpha_{e2} = 0.5$ (50% borrowing from external control):

```{r ext-post}
p_ext <- pbayespostpred1cont(
  prob = 'posterior', design = 'external', prior = 'vague',
  CalcMethod = 'MM', theta0 = 1.5, nMC = NULL,
  n1 = 15, n2 = 15,
  bar.y1 = 3.2, s1 = 2.0,
  bar.y2 = 1.1, s2 = 1.8,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = 20L, alpha01 = NULL, alpha02 = 0.5,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = 0.9, se2 = 1.8,
  lower.tail = FALSE
)
cat(sprintf("P(theta > theta_TV | data, external control, alpha=0.5) = %.4f\n",
            p_ext))
```

**Effect of the borrowing weight**: varying $\alpha_{e2}$ from 0 to 1
shows how external data influence the posterior.

```{r ext-borrowing}
# alpha02 must be in (0, 1]; start from 0.01 to avoid validation error
alpha_seq <- c(0.01, seq(0.1, 1.0, by = 0.1))
p_alpha   <- sapply(alpha_seq, function(a) {
  pbayespostpred1cont(
    prob = 'posterior', design = 'external', prior = 'vague',
    CalcMethod = 'MM', theta0 = 1.5, nMC = NULL,
    n1 = 15, n2 = 15,
    bar.y1 = 3.2, s1 = 2.0,
    bar.y2 = 1.1, s2 = 1.8,
    m1 = NULL, m2 = NULL,
    kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
    mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
    r = NULL,
    ne1 = NULL, ne2 = 20L, alpha01 = NULL, alpha02 = a,
    bar.ye1 = NULL, se1 = NULL, bar.ye2 = 0.9, se2 = 1.8,
    lower.tail = FALSE
  )
})
data.frame(alpha_e2 = alpha_seq, P_gt_TV = round(p_alpha, 4))
```

---

## 5. Operating Characteristics

### 5.1 Definition

For a given true parameter $(\mu_1, \mu_2, \sigma_1, \sigma_2)$, the
operating characteristics are the probabilities of each decision outcome
under the Go/NoGo/Gray/Miss rule with thresholds $(\gamma_1, \gamma_2)$.
These are estimated by Monte Carlo simulation over $n_{\mathrm{sim}}$
replicates:

$$\widehat{\Pr}(\mathrm{Go}) =
  \frac{1}{n_{\mathrm{sim}}} \sum_{i=1}^{n_{\mathrm{sim}}}
  \mathbf{1}\!\left[g_{\mathrm{Go}}^{(i)} \ge \gamma_1
    \;\text{ and }\; g_{\mathrm{NoGo}}^{(i)} < \gamma_2\right],$$

$$\widehat{\Pr}(\mathrm{NoGo}) =
  \frac{1}{n_{\mathrm{sim}}} \sum_{i=1}^{n_{\mathrm{sim}}}
  \mathbf{1}\!\left[g_{\mathrm{Go}}^{(i)} < \gamma_1
    \;\text{ and }\; g_{\mathrm{NoGo}}^{(i)} \ge \gamma_2\right],$$

$$\widehat{\Pr}(\mathrm{Gray}) = 1 -
  \widehat{\Pr}(\mathrm{Go}) - \widehat{\Pr}(\mathrm{NoGo}) -
  \widehat{\Pr}(\mathrm{Miss}),$$

where

$$g_{\mathrm{Go}}^{(i)} = P(\theta > \theta_{\mathrm{TV}} \mid \mathbf{y}^{(i)}),
\qquad
g_{\mathrm{NoGo}}^{(i)} = P(\theta \le \theta_{\mathrm{MAV}} \mid \mathbf{y}^{(i)})$$

are evaluated for the $i$-th simulated dataset $\mathbf{y}^{(i)}$.
A Miss ($g_{\mathrm{Go}} \ge \gamma_1$ AND $g_{\mathrm{NoGo}} \ge \gamma_2$)
indicates an inconsistent threshold configuration and triggers an error by
default (`error_if_Miss = TRUE`).

### 5.2 Controlled Design — Multiple Scenarios

```{r oc-controlled}
oc_ctrl <- pbayesdecisionprob1cont(
  nsim       = 500L,
  prob       = 'posterior',
  design     = 'controlled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.5,  theta.MAV = 0.5,  theta.NULL = NULL,
  nMC        = NULL,
  gamma1     = 0.80, gamma2    = 0.20,
  n1 = 15,   n2 = 15,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  mu1    = c(1.0, 1.5, 2.0, 2.5, 3.0, 3.5),
  mu2    = 1.0,
  sigma1 = 2.0,
  sigma2 = 2.0,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE,
  seed = 42L
)
print(oc_ctrl)
```

### 5.3 Figure: Go/NoGo/Gray Probabilities vs. True Treatment Effect

```{r oc-plot, fig.cap = "Operating characteristics (Go/NoGo/Gray) as a function of the true treatment effect (mu1 - mu2). Controlled design, vague prior, MM method. True mu2 = 1.0, sigma1 = sigma2 = 2.0, n1 = n2 = 15, gamma1 = 0.80, gamma2 = 0.20."}
# Compute OC over a finer grid of true treatment means
mu1_seq <- seq(1.0, 4.5, by = 0.5)

oc_grid <- pbayesdecisionprob1cont(
  nsim       = 500L,
  prob       = 'posterior',
  design     = 'controlled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.5,  theta.MAV = 0.5,  theta.NULL = NULL,
  nMC        = NULL,
  gamma1     = 0.80, gamma2    = 0.20,
  n1 = 15,   n2 = 15,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  mu1    = mu1_seq,
  mu2    = 1.0,
  sigma1 = 2.0,
  sigma2 = 2.0,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE,
  seed = 42L
)

delta   <- oc_grid$mu1 - 1.0  # true treatment effect (mu1 - mu2)

# --- Plot ---
old_par <- par(mar = c(4.5, 4.5, 3, 1))

plot(delta, oc_grid$Go,
     type = "b", pch = 16, lwd = 2,
     col  = "#2166AC",
     ylim = c(0, 1), xlab = expression(mu[1] - mu[2]),
     ylab = "Probability",
     main = "Operating Characteristics: Controlled Design (Posterior Prob.)",
     las  = 1)
lines(delta, oc_grid$NoGo,
      type = "b", pch = 17, lwd = 2, col = "#D6604D")
lines(delta, oc_grid$Gray,
      type = "b", pch = 15, lwd = 2, col = "#878787")

abline(h  = c(0.20, 0.80), lty = 2, col = "gray50")
abline(v  = c(0.5, 1.5),   lty = 3, col = "gray70")

legend("right",
       legend = c("Pr(Go)", "Pr(NoGo)", "Pr(Gray)"),
       col    = c("#2166AC", "#D6604D", "#878787"),
       pch    = c(16, 17, 15), lwd = 2, bty = "n")

text(0.5,  0.05, expression(theta[MAV]), col = "gray40", cex = 0.8)
text(1.5,  0.05, expression(theta[TV]),  col = "gray40", cex = 0.8)

par(old_par)
```

### 5.4 Uncontrolled Design — Operating Characteristics

```{r oc-uncontrolled}
oc_unctrl <- pbayesdecisionprob1cont(
  nsim       = 500L,
  prob       = 'posterior',
  design     = 'uncontrolled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.5,  theta.MAV = 0.5,  theta.NULL = NULL,
  nMC        = NULL,
  gamma1     = 0.80, gamma2    = 0.20,
  n1 = 15,   n2 = 15,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = 1.0, sigma01 = NULL, sigma02 = NULL,
  mu1    = c(1.5, 2.0, 2.5, 3.0, 3.5),
  mu2    = 1.0,
  sigma1 = 2.0,
  sigma2 = 2.0,
  r      = 1.0,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = NULL, se2 = NULL,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE,
  seed = 1L
)
print(oc_unctrl)
```

### 5.5 External Control Design — Operating Characteristics

```{r oc-external}
oc_ext <- pbayesdecisionprob1cont(
  nsim       = 500L,
  prob       = 'posterior',
  design     = 'external',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.5,  theta.MAV = 0.5,  theta.NULL = NULL,
  nMC        = NULL,
  gamma1     = 0.80, gamma2    = 0.20,
  n1 = 15,   n2 = 15,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  mu1    = c(1.5, 2.0, 2.5, 3.0, 3.5),
  mu2    = 1.0,
  sigma1 = 2.0,
  sigma2 = 2.0,
  r = NULL,
  ne1 = NULL, ne2 = 20L, alpha01 = NULL, alpha02 = 0.5,
  bar.ye1 = NULL, se1 = NULL, bar.ye2 = 0.9, se2 = 1.8,
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE,
  seed = 2L
)
print(oc_ext)
```

---

## 6. Optimal Threshold Search

### 6.1 Objective

The `getgamma1cont()` function finds the optimal pair
$(\gamma_1^*, \gamma_2^*)$ by grid search over
$\Gamma = \{\gamma^{(1)}, \ldots, \gamma^{(K)}\} \subset (0, 1)$, where:

- $\gamma_1^*$: selected so that $\Pr(\mathrm{Go} \mid \gamma_1^*)$ satisfies
  a criterion against a target value (e.g., $\Pr(\mathrm{Go}) < 0.05$
  under the null scenario $\mu_1 = \mu_2$).
- $\gamma_2^*$: selected so that $\Pr(\mathrm{NoGo} \mid \gamma_2^*)$
  satisfies a criterion (e.g., $\Pr(\mathrm{NoGo}) < 0.20$ under the null
  scenario).

The search uses a two-stage strategy:

**Stage 1 (Simulate once)**: $n_{\mathrm{sim}}$ datasets are generated
from the true parameter distribution and the probabilities
$g_{\mathrm{Go}}^{(i)}$ and $g_{\mathrm{NoGo}}^{(i)}$ are computed for
each replicate — independently of $\gamma$.

**Stage 2 (Sweep over $\Gamma$)**: for each candidate $\gamma \in \Gamma$,
$\Pr(\mathrm{Go} \mid \gamma)$ and $\Pr(\mathrm{NoGo} \mid \gamma)$ are
computed as empirical proportions over the precomputed indicator values
without further probability evaluations.

### 6.2 Example: Controlled Design, Posterior Probability

```{r getgamma-ctrl}
res_gamma <- getgamma1cont(
  nsim       = 1000L,
  prob       = 'posterior',
  design     = 'controlled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.5, theta.MAV = 0.5, theta.NULL = NULL,
  nMC        = NULL,
  mu1 = 1.0, mu2 = 1.0,   # null scenario: no treatment effect
  sigma1 = 2.0, sigma2 = 2.0,
  target_go   = 0.05, target_nogo   = 0.20,
  crit_go     = '<',  crit_nogo     = '<',
  sel_go      = 'smallest', sel_nogo = 'largest',
  n1 = 15L, n2 = 15L, m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, bar.ye2 = NULL, se1 = NULL, se2 = NULL,
  gamma_grid = seq(0.01, 0.99, by = 0.01),
  seed = 42L
)
print(res_gamma)
```

### 6.3 Figure: Pr(Go) and Pr(NoGo) Over the Gamma Grid

```{r gamma-plot, fig.cap = "Pr(Go) and Pr(NoGo) as a function of the candidate threshold gamma. The optimal gamma1 (vertical blue dashed line) is the smallest gamma satisfying Pr(Go) < 0.05 under the null scenario. The optimal gamma2 (vertical red dashed line) is the largest gamma satisfying Pr(NoGo) < 0.20 under the null scenario."}
grid_df <- data.frame(
  gamma   = res_gamma$gamma_grid,
  PrGo    = res_gamma$PrGo_grid,
  PrNoGo  = res_gamma$PrNoGo_grid
)

old_par2 <- par(mar = c(4.5, 4.5, 3, 1))

# Initialize the plot before calling low-level plotting functions (lines, abline, text)
plot(grid_df$gamma, grid_df$PrGo,
     type = "l", lwd = 2, col = "#2166AC",
     ylim = c(0, 1),
     xlab = expression(gamma),
     ylab = "Probability",
     main = "Optimal Threshold Search: Pr(Go) and Pr(NoGo) vs. Gamma",
     las  = 1)

lines(grid_df$gamma, grid_df$PrNoGo,
      lwd = 2, col = "#D6604D")

abline(h = 0.05, lty = 2, col = "#2166AC", lwd = 1.2)
abline(h = 0.20, lty = 2, col = "#D6604D", lwd = 1.2)

if (!is.na(res_gamma$gamma1)) {
  abline(v = res_gamma$gamma1, lty = 2, col = "#2166AC", lwd = 2)
  text(res_gamma$gamma1 + 0.02, 0.85,
       paste0("gamma1* = ", res_gamma$gamma1),
       col = "#2166AC", cex = 0.85, adj = 0)
}

if (!is.na(res_gamma$gamma2)) {
  abline(v = res_gamma$gamma2, lty = 2, col = "#D6604D", lwd = 2)
  text(res_gamma$gamma2 - 0.02, 0.70,
       paste0("gamma2* = ", res_gamma$gamma2),
       col = "#D6604D", cex = 0.85, adj = 1)
}

legend("topright",
       legend = c("Pr(Go | gamma)", "Pr(NoGo | gamma)",
                  "Target Go = 0.05", "Target NoGo = 0.20"),
       col    = c("#2166AC", "#D6604D", "#2166AC", "#D6604D"),
       lty    = c(1, 1, 2, 2), lwd = c(2, 2, 1.2, 1.2),
       bty    = "n", cex = 0.85)

par(old_par2)
```

### 6.4 Example: Uncontrolled Design, Posterior Probability

```{r getgamma-unctrl}
res_gamma_unctrl <- getgamma1cont(
  nsim       = 1000L,
  prob       = 'posterior',
  design     = 'uncontrolled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = 1.5, theta.MAV = 0.5, theta.NULL = NULL,
  nMC        = NULL,
  mu1 = 1.0, mu2 = NULL,
  sigma1 = 2.0, sigma2 = NULL,
  target_go   = 0.05, target_nogo   = 0.20,
  crit_go     = '<',  crit_nogo     = '<',
  sel_go      = 'smallest', sel_nogo = 'largest',
  n1 = 15L, n2 = 15L, m1 = NULL, m2 = NULL,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = 1.0, sigma01 = NULL, sigma02 = NULL,
  r = 1.0,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, bar.ye2 = NULL, se1 = NULL, se2 = NULL,
  gamma_grid = seq(0.01, 0.99, by = 0.01),
  seed = 1L
)
print(res_gamma_unctrl)
```

### 6.5 Example: Controlled Design, Predictive Probability

```{r getgamma-pred}
res_gamma_pred <- getgamma1cont(
  nsim       = 1000L,
  prob       = 'predictive',
  design     = 'controlled',
  prior      = 'vague',
  CalcMethod = 'MM',
  theta.TV   = NULL, theta.MAV = NULL, theta.NULL = 1.0,
  nMC        = NULL,
  mu1 = 1.0, mu2 = 1.0,
  sigma1 = 2.0, sigma2 = 2.0,
  target_go   = 0.05, target_nogo   = 0.20,
  crit_go     = '<',  crit_nogo     = '<',
  sel_go      = 'smallest', sel_nogo = 'largest',
  n1 = 15L, n2 = 15L, m1 = 60L, m2 = 60L,
  kappa01 = NULL, kappa02 = NULL, nu01 = NULL, nu02 = NULL,
  mu01 = NULL, mu02 = NULL, sigma01 = NULL, sigma02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01 = NULL, alpha02 = NULL,
  bar.ye1 = NULL, bar.ye2 = NULL, se1 = NULL, se2 = NULL,
  gamma_grid = seq(0.01, 0.99, by = 0.01),
  seed = 3L
)
print(res_gamma_pred)
```

---

## 7. Summary

This vignette covered single continuous endpoint analysis in BayesianQDM:

1. **Bayesian model**: N-Inv-$\chi^2$ conjugate posterior (vague and
   informative priors) with explicit update formulae for $\kappa_{nj}$,
   $\nu_{nj}$, $\mu_{nj}$, and $\sigma_{nj}^2$.
2. **Posterior and predictive distributions**: marginal $t$-distributions
   for $\mu_j$ and $\bar{\tilde{y}}_j$, with explicit scale expressions.
3. **Three computation methods**: NI (exact), MC (simulation), MM
   (moment-matching via fourth-moment equation); MM is recommended for
   large-scale simulation ($\nu_j > 4$ required).
4. **Three study designs**: controlled (both arms observed), uncontrolled
   ($\mu_{02}$ and $r$ fixed from prior knowledge), external control
   (power prior with borrowing weight $\alpha_{ej}$).
5. **Operating characteristics**: Monte Carlo estimation of
   $\Pr(\mathrm{Go})$, $\Pr(\mathrm{NoGo})$, $\Pr(\mathrm{Gray})$
   across true parameter scenarios.
6. **Optimal threshold search**: two-stage precompute-then-sweep strategy
   in `getgamma1cont()` with user-specified criteria and selection rules.

See `vignette("single-binary")` for the analogous binary endpoint analysis.
