---
title: "Two-Endpoint Bayesian QDM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two-Endpoint Bayesian QDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(BayesianQDM)
```

## Motivating Scenario

We extend the hypothetical Phase IIa PoC trial in rheumatoid arthritis (RA)
to co-primary endpoints, evaluating both efficacy dimensions simultaneously.

* **Two binary endpoints** — ACR20 and ACR50 response rates.
* **Two continuous endpoints** — Change from baseline in DAS28 score (Endpoint 1)
  and HAQ-DI disability index (Endpoint 2).

The 9-region decision grid (posterior probability) and 4-region grid
(predictive probability) are defined over the Cartesian product of the
Go/Gray/NoGo zones for each endpoint individually.

```{r region-table, echo = FALSE}
knitr::kable(data.frame(
  Region = paste0("R", 1:9),
  EP1    = c(">TV1",  ">TV1",  ">TV1",
             "MAV1 < . <= TV1", "MAV1 < . <= TV1", "MAV1 < . <= TV1",
             "<= MAV1", "<= MAV1", "<= MAV1"),
  EP2    = c(">TV2",  "MAV2 < . <= TV2", "<= MAV2",
             ">TV2",  "MAV2 < . <= TV2", "<= MAV2",
             ">TV2",  "MAV2 < . <= TV2", "<= MAV2")
), col.names = c("Region", "Endpoint 1", "Endpoint 2"),
caption = "9-region grid for posterior probability (prob = 'posterior')")
```

Typically **R1** (both endpoints exceed TV) is designated as the Go region
and **R9** (both endpoints below MAV) as the NoGo region.

---

## Part 1: Two Binary Endpoints (ACR20 + ACR50)

The Dirichlet-multinomial conjugate model is used.  Observed counts are the
four bivariate response patterns: (ACR20=0, ACR50=0), (ACR20=0, ACR50=1),
(ACR20=1, ACR50=0), (ACR20=1, ACR50=1).  Since ACR50 implies ACR20, the
(0,1) cell is impossible in practice; we set `x_01 = 0`.

```{r bin2-scenario, echo = FALSE}
knitr::kable(data.frame(
  Design   = c("Treatment arm", "Control arm"),
  x_00 = c(2, 6),
  x_01 = c(0, 0),
  x_10 = c(3, 3),
  x_11 = c(5, 1)
), col.names = c("Arm", "Neither (0,0)", "ACR50 only (0,1)",
                  "ACR20 only (1,0)", "Both (1,1)"),
caption = "Observed bivariate counts (n1 = n2 = 10)")
```

### 1-1. Posterior Probability — Controlled Design

```{r bin2-post}
prob_bin2 <- pbayespostpred2bin(
  prob   = 'posterior',
  design = 'controlled',
  theta.TV1  = 0.20,  theta.MAV1 = 0.05,
  theta.TV2  = 0.20,  theta.MAV2 = 0.05,
  x1_00 = 2L, x1_01 = 0L, x1_10 = 3L, x1_11 = 5L,
  x2_00 = 6L, x2_01 = 0L, x2_10 = 3L, x2_11 = 1L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 2000L
)
round(prob_bin2, 4)
```

The Go decision rule is $P(R1) \ge \gamma_1 = 0.80$; NoGo is
$P(R9) \ge \gamma_2 = 0.20$.

```{r bin2-decision}
cat(sprintf("P(R1) = %.4f  ->  %s\n", prob_bin2["R1"],
            ifelse(prob_bin2["R1"] >= 0.80, "Go", "not Go")))
cat(sprintf("P(R9) = %.4f  ->  %s\n", prob_bin2["R9"],
            ifelse(prob_bin2["R9"] >= 0.20, "NoGo", "not NoGo")))
```

### 1-2. Posterior Predictive Probability — Controlled Design

For predictive probability, TV and MAV must be equal (single null threshold).
The 4-region grid applies.

```{r bin2-pred}
prob_bin2_pred <- pbayespostpred2bin(
  prob   = 'predictive',
  design = 'controlled',
  theta.TV1  = 0.10,  theta.MAV1 = 0.10,
  theta.TV2  = 0.10,  theta.MAV2 = 0.10,
  x1_00 = 2L, x1_01 = 0L, x1_10 = 3L, x1_11 = 5L,
  x2_00 = 6L, x2_01 = 0L, x2_10 = 3L, x2_11 = 1L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = 50L, m2 = 50L,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 2000L
)
round(prob_bin2_pred, 4)
```

### 1-3. Operating Characteristics — Controlled Design

`error_if_Miss = FALSE` is used here because with small samples the Go and
NoGo regions can both be triggered simultaneously (Miss event).

```{r bin2-oc}
oc_bin2 <- pbayesdecisionprob2bin(
  prob        = 'posterior',
  design      = 'controlled',
  GoRegions   = 1L,
  NoGoRegions = 9L,
  gamma1      = 0.60,
  gamma2      = 0.60,
  pi_t1  = c(0.40, 0.60, 0.80),
  pi_t2  = c(0.25, 0.40, 0.55),
  rho_t  = rep(0.5, 3),
  pi_c1  = rep(0.15, 3),
  pi_c2  = rep(0.05, 3),
  rho_c  = rep(0.3, 3),
  n1 = 10, n2 = 10,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  theta.TV1  = 0.20,  theta.MAV1 = 0.05,
  theta.TV2  = 0.20,  theta.MAV2 = 0.05,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 500L,
  error_if_Miss = FALSE, Gray_inc_Miss = FALSE
)
print(oc_bin2)
```

### 1-4. Uncontrolled Design

When no concurrent control arm is available, hypothetical control pseudo-counts
`z00`, `z01`, `z10`, `z11` are specified.

```{r bin2-unctrl}
prob_bin2_unctrl <- pbayespostpred2bin(
  prob   = 'posterior',
  design = 'uncontrolled',
  theta.TV1  = 0.20,  theta.MAV1 = 0.05,
  theta.TV2  = 0.20,  theta.MAV2 = 0.05,
  x1_00 = 2L, x1_01 = 0L, x1_10 = 3L, x1_11 = 5L,
  x2_00 = NULL, x2_01 = NULL, x2_10 = NULL, x2_11 = NULL,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  z00 = 5L, z01 = 0L, z10 = 3L, z11 = 2L,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 2000L
)
round(prob_bin2_unctrl, 4)
```

---

## Part 2: Two Continuous Endpoints (DAS28 + HAQ-DI)

The bivariate Normal-Inverse-Wishart (NIW) conjugate model is used.
Sufficient statistics supplied are the sample mean vectors and
sum-of-squares matrices.

```{r cont2-data}
# Sample means
ybar1 <- c(3.5, 0.6)   # treatment: DAS28 change, HAQ-DI change
ybar2 <- c(1.2, 0.2)   # control

# Sum-of-squares matrices (S = sum of (y_j - ybar)(y_j - ybar)^T)
S1 <- matrix(c(32, 5, 5, 8), 2, 2)   # treatment (n1 = 10)
S2 <- matrix(c(28, 4, 4, 7), 2, 2)   # control   (n2 = 10)
```

Clinically meaningful thresholds:

| Endpoint | TV   | MAV  |
|----------|------|------|
| DAS28 change difference | 1.50 | 0.50 |
| HAQ-DI change difference | 0.40 | 0.10 |

### 2-1. Posterior Probability — Vague Prior, MM Method

```{r cont2-post-vague}
prob_cont2 <- pbayespostpred2cont(
  prob   = 'posterior',
  design = 'controlled',
  prior  = 'vague',
  theta.TV1  = 1.50,  theta.MAV1 = 0.50,
  theta.TV2  = 0.40,  theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 10L, n2 = 10L,
  ybar1 = ybar1, S1 = S1,
  ybar2 = ybar2, S2 = S2,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, Se1 = NULL, ybar_e2 = NULL, Se2 = NULL,
  nMC = 2000L, method = 'MM'
)
round(prob_cont2, 4)
```

```{r cont2-decision}
cat(sprintf("P(R1) = %.4f  ->  %s\n", prob_cont2["R1"],
            ifelse(prob_cont2["R1"] >= 0.80, "Go", "not Go")))
cat(sprintf("P(R9) = %.4f  ->  %s\n", prob_cont2["R9"],
            ifelse(prob_cont2["R9"] >= 0.20, "NoGo", "not NoGo")))
```

### 2-2. Informative Prior (Normal-Inverse-Wishart)

Prior information from a similar previous trial is encoded in the NIW
hyperparameters.

```{r cont2-post-niw}
Lambda0 <- matrix(c(20, 3, 3, 5), 2, 2)

prob_cont2_niw <- pbayespostpred2cont(
  prob   = 'posterior',
  design = 'controlled',
  prior  = 'N-Inv-Wishart',
  theta.TV1  = 1.50,  theta.MAV1 = 0.50,
  theta.TV2  = 0.40,  theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 10L, n2 = 10L,
  ybar1 = ybar1, S1 = S1,
  ybar2 = ybar2, S2 = S2,
  m1 = NULL, m2 = NULL,
  kappa01 = 3,  nu01 = 5,  mu01 = c(3.0, 0.5), Lambda01 = Lambda0,
  kappa02 = 3,  nu02 = 5,  mu02 = c(1.0, 0.2), Lambda02 = Lambda0,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, Se1 = NULL, ybar_e2 = NULL, Se2 = NULL,
  nMC = 2000L, method = 'MM'
)
round(prob_cont2_niw, 4)
```

### 2-3. Posterior Predictive Probability — Vague Prior

```{r cont2-pred}
set.seed(1)
prob_cont2_pred <- pbayespostpred2cont(
  prob   = 'predictive',
  design = 'controlled',
  prior  = 'vague',
  theta.TV1  = NULL, theta.MAV1 = NULL,
  theta.TV2  = NULL, theta.MAV2 = NULL,
  theta.NULL1 = 1.00, theta.NULL2 = 0.30,
  n1 = 10L, n2 = 10L,
  ybar1 = ybar1, S1 = S1,
  ybar2 = ybar2, S2 = S2,
  m1 = 60L, m2 = 60L,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, Se1 = NULL, ybar_e2 = NULL, Se2 = NULL,
  nMC = 2000L, method = 'MC'
)
round(prob_cont2_pred, 4)
```

### 2-4. Operating Characteristics — Vague Prior, MM Method

`error_if_Miss = FALSE` is used because with small samples the Go and NoGo
regions can both be triggered simultaneously.

```{r cont2-oc}
# True treatment mean matrix: each row = one scenario (EP1, EP2)
mu1_mat <- rbind(c(1.5, 0.3),
                 c(2.5, 0.5),
                 c(3.5, 0.6))
mu2_mat <- rbind(c(1.0, 0.2),
                 c(1.0, 0.2),
                 c(1.0, 0.2))

Sigma_true <- matrix(c(4.0, 0.8, 0.8, 1.0), 2, 2)

oc_cont2 <- pbayesdecisionprob2cont(
  nsim   = 100L,
  prob   = 'posterior',
  design = 'controlled',
  prior  = 'vague',
  GoRegions   = 1L,
  NoGoRegions = 9L,
  gamma1 = 0.60,  gamma2 = 0.60,
  theta.TV1  = 1.50,  theta.MAV1 = 0.50,
  theta.TV2  = 0.40,  theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 10L, n2 = 10L, m1 = NULL, m2 = NULL,
  mu1    = mu1_mat,
  Sigma1 = Sigma_true,
  mu2    = mu2_mat,
  Sigma2 = Sigma_true,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, Se1 = NULL, ybar_e2 = NULL, Se2 = NULL,
  method = 'MM', nMC = 500L,
  error_if_Miss = FALSE, Gray_inc_Miss = FALSE,
  seed = 1L
)
print(oc_cont2)
```

### 2-5. Uncontrolled Design (Single-Arm)

For a single-arm study, the hypothetical control mean vector `mu02` and
variance scaling ratio `r` replace the control arm data.

```{r cont2-unctrl}
oc_cont2_unctrl <- pbayesdecisionprob2cont(
  nsim   = 100L,
  prob   = 'posterior',
  design = 'uncontrolled',
  prior  = 'vague',
  GoRegions   = 1L,
  NoGoRegions = 9L,
  gamma1 = 0.60,  gamma2 = 0.60,
  theta.TV1  = 1.50,  theta.MAV1 = 0.50,
  theta.TV2  = 0.40,  theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 10L, n2 = NULL, m1 = NULL, m2 = NULL,
  mu1    = mu1_mat,
  Sigma1 = Sigma_true,
  mu2    = mu2_mat,
  Sigma2 = Sigma_true,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = c(1.0, 0.2), Lambda02 = NULL,
  r = 1.0,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, Se1 = NULL, ybar_e2 = NULL, Se2 = NULL,
  method = 'MM', nMC = 500L,
  error_if_Miss = FALSE, Gray_inc_Miss = FALSE,
  seed = 2L
)
print(oc_cont2_unctrl)
```

---

## Summary

This vignette demonstrated:

1. **Two binary endpoints** (ACR20 + ACR50) using the Dirichlet-multinomial
   model (`pbayespostpred2bin`, `pbayesdecisionprob2bin`).
2. **Two continuous endpoints** (DAS28 + HAQ-DI) using the bivariate NIW model
   (`pbayespostpred2cont`, `pbayesdecisionprob2cont`).
3. The **9-region** (posterior) and **4-region** (predictive) decision grids.
4. **Vague and informative priors**, controlled and uncontrolled designs.
5. **MM method** for efficient operating characteristics evaluation.

See `vignette("single-endpoint")` for single-endpoint analyses, and
`vignette("overview")` for the overall framework.
