---
title: "Two Continuous Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two Continuous Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 5.5,
  fig.path = "figures/2cont-"
)
library(BayesianQDM)
```

## Motivating Scenario

We extend the single-endpoint RA trial to a bivariate design with two
co-primary continuous endpoints:

- Endpoint 1: change from baseline in DAS28 score
- Endpoint 2: change from baseline in HAQ-DI (Health Assessment
  Questionnaire -- Disability Index)

The trial enrolls $n_t = n_c = 20$ patients per group (treatment vs. placebo)
in a 1:1 randomised controlled design.

Clinically meaningful thresholds (posterior probability):

|                                       | Endpoint 1 | Endpoint 2 |
|---------------------------------------|:----------:|:----------:|
| $\theta_{\mathrm{TV},1}$             | 1.5        | 1.0        |
| $\theta_{\mathrm{MAV},1}$            | 0.5        | 0.3        |

Null hypothesis thresholds (predictive probability):
$\theta_{\mathrm{NULL},1} = 0.5$, $\theta_{\mathrm{NULL},2} = 0.3$.

Decision thresholds: $\gamma_1 = 0.80$ (Go), $\gamma_2 = 0.20$ (NoGo).

Observed data (used in Sections 2--4):

- Treatment: $\bar{\mathbf{y}}_t = (3.5, 2.1)^\top$,
  $S_t = \begin{pmatrix}18.0 & 3.6\\3.6 & 9.0\end{pmatrix}$
- Control: $\bar{\mathbf{y}}_c = (1.8, 1.0)^\top$,
  $S_c = \begin{pmatrix}16.0 & 2.8\\2.8 & 8.5\end{pmatrix}$

where $S_j = \sum_{i=1}^{n_j}(\mathbf{y}_{ij} - \bar{\mathbf{y}}_j)
(\mathbf{y}_{ij} - \bar{\mathbf{y}}_j)^\top$ ($j = t, c$) is the sum-of-squares matrix.

---

## 1. Bayesian Model: Normal-Inverse-Wishart Conjugate

### 1.1 Prior Distribution

For group $j$ ($j = t$: treatment, $j = c$: control), we model the
bivariate outcomes for patient $i$ ($i = 1, \ldots, n_j$) as

$$\mathbf{y}_{ij} \mid \boldsymbol{\mu}_j, \Sigma_j \;\overset{iid}{\sim}\;
  \mathcal{N}_2(\boldsymbol{\mu}_j,\, \Sigma_j).$$

The conjugate prior is the Normal-Inverse-Wishart (NIW) distribution:

$$(\boldsymbol{\mu}_j, \Sigma_j) \;\sim\;
  \mathrm{NIW}(\boldsymbol{\mu}_{0j},\, \kappa_{0j},\, \nu_{0j},\,
  \Lambda_{0j}),$$

where $\boldsymbol{\mu}_{0j} \in \mathbb{R}^2$ is the prior mean
(argument `mu0_t` or `mu0_c`),
$\kappa_{0j} > 0$ is the prior concentration
(`kappa0_t` or `kappa0_c`),
$\nu_{0j} > 3$ is the prior degrees of freedom
(`nu0_t` or `nu0_c`), and
$\Lambda_{0j}$ is a $2 \times 2$ positive-definite prior scale matrix
(`Lambda0_t` or `Lambda0_c`).

The vague (Jeffreys) prior is obtained by letting
$\kappa_{0j} \to 0$ and $\nu_{0j} \to 0$ (`prior = 'vague'`).

### 1.2 Posterior Distribution

Given $n_j$ observations with sample mean $\bar{\mathbf{y}}_j$
(`ybar_t` or `ybar_c`) and
sum-of-squares matrix $S_j$ (`S_t` or `S_c`),
the NIW posterior has updated parameters:

$$\kappa_{nj} = \kappa_{0j} + n_j, \qquad
  \nu_{nj} = \nu_{0j} + n_j,$$

$$\boldsymbol{\mu}_{nj} =
  \frac{\kappa_{0j}\,\boldsymbol{\mu}_{0j} + n_j\,\bar{\mathbf{y}}_j}
       {\kappa_{nj}},$$

$$\Lambda_{nj} = \Lambda_{0j} + S_j +
  \frac{\kappa_{0j}\,n_j}{\kappa_{nj}}
  (\bar{\mathbf{y}}_j - \boldsymbol{\mu}_{0j})
  (\bar{\mathbf{y}}_j - \boldsymbol{\mu}_{0j})^\top.$$

The marginal posterior of the group mean $\boldsymbol{\mu}_j$ is a
bivariate non-standardised $t$-distribution:

$$\boldsymbol{\mu}_j \mid \mathbf{Y}_j \;\sim\;
  t_{\nu_{nj} - 1}\!\!\left(\boldsymbol{\mu}_{nj},\;
  \frac{\Lambda_{nj}}{\kappa_{nj}(\nu_{nj} - 1)}\right).$$

Under the vague prior, this simplifies to

$$\boldsymbol{\mu}_j \mid \mathbf{Y}_j \;\sim\;
  t_{n_j - 2}\!\!\left(\bar{\mathbf{y}}_j,\;
  \frac{S_j}{n_j(n_j - 2)}\right).$$

### 1.3 Posterior of the Bivariate Treatment Effect

The bivariate treatment effect is $\boldsymbol{\theta} =
\boldsymbol{\mu}_t - \boldsymbol{\mu}_c$.  Since the two groups are
independent, the posterior of $\boldsymbol{\theta}$ is the distribution
of the difference of two independent bivariate $t$-vectors:

$$T_j \;\sim\; t_{\nu_{nj}-1}(\boldsymbol{\mu}_{nj},\, V_j),$$

where the posterior scale matrix is
$V_j = \Lambda_{nj} / [\kappa_{nj}(\nu_{nj}-1)]$.

The posterior probability that $\boldsymbol{\theta}$ falls in a
rectangular region $R$ defined by thresholds on each component is

$$P(\boldsymbol{\theta} \in R \mid \mathbf{Y}) =
  P(T_t - T_c \in R),$$

computed by one of two methods (MC or MM) described in Section 3.

### 1.4 Nine-Region Grid (Posterior Probability)

The bivariate threshold grid for $(\theta_1, \theta_2)$ creates a
$3 \times 3$ partition using `theta_TV1`, `theta_MAV1`, `theta_TV2`,
`theta_MAV2`:

```{r nine-region-cont, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Nine-region grid for two-endpoint posterior probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="3" style="border:1px solid #aaa; padding:6px; font-weight:normal;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px; font-weight:normal;">&theta;<sub>1</sub> &gt; &theta;<sub>TV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px; font-weight:normal;">&theta;<sub>TV1</sub> &ge; &theta;<sub>1</sub> &gt; &theta;<sub>MAV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px; font-weight:normal;">&theta;<sub>MAV1</sub> &ge; &theta;<sub>1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="3" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg);">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>TV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
    <td style="border:1px solid #aaa; padding:6px;">R7</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>TV2</sub> &ge; &theta;<sub>2</sub> &gt; &theta;<sub>MAV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R5</td>
    <td style="border:1px solid #aaa; padding:6px;">R8</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>MAV2</sub> &ge; &theta;<sub>2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
    <td style="border:1px solid #aaa; padding:6px;">R6</td>
    <td style="border:1px solid #aaa; padding:6px;">R9</td>
  </tr>
</tbody>
</table>
')
```

Region $R_1$ (both endpoints exceed $\theta_{\mathrm{TV}}$) is typically
designated as Go and $R_9$ (both endpoints below $\theta_{\mathrm{MAV}}$)
as NoGo.

### 1.5 Four-Region Grid (Predictive Probability)

For the predictive probability, a $2 \times 2$ partition is used,
defined by `theta_NULL1` and `theta_NULL2`:

```{r four-region-cont, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Four-region grid for two-endpoint predictive probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="2" style="border:1px solid #aaa; padding:6px; font-weight:normal;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px; font-weight:normal;">&theta;<sub>1</sub> &gt; &theta;<sub>NULL1</sub></th>
    <th style="border:1px solid #aaa; padding:6px; font-weight:normal;">&theta;<sub>1</sub> &le; &theta;<sub>NULL1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="2" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg);">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &le; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
  </tr>
</tbody>
</table>
')
```

Region $R_1$ (both endpoints exceed $\theta_{\mathrm{NULL}}$) is typically
designated as Go and $R_4$ (both endpoints at or below $\theta_{\mathrm{NULL}}$)
as NoGo.

---

## 2. Posterior Predictive Distribution

The predictive distribution of the future sample mean
$\bar{\tilde{\mathbf{y}}}_j$ based on $m_j$ future patients
(`m_t` or `m_c`) is again a bivariate $t$-distribution:

$$\bar{\tilde{\mathbf{y}}}_j \mid \mathbf{Y}_j \;\sim\;
  t_{\nu_{nj}-1}\!\!\left(\boldsymbol{\mu}_{nj},\;
  \frac{\Lambda_{nj}}{\kappa_{nj}(\nu_{nj}-1)}
  \cdot \frac{\kappa_{nj} + 1}{\kappa_{nj}} \cdot \frac{1}{m_j}
  \right)$$

under the NIW prior.  Under the vague prior the scale inflates to
$S_j(n_j + 1) / [n_j^2(n_j - 2) m_j]$.

---

## 3. Two Computation Methods

The computation method is specified via the `CalcMethod` argument.

### 3.1 Monte Carlo Simulation (`CalcMethod = 'MC'`)

For each of $n_\mathrm{MC}$ draws (`nMC`), independent samples are generated:

$$T_t^{(i)} \;\sim\; t_{\nu_{nt}-1}(\boldsymbol{\mu}_{nt}, V_t), \qquad
  T_c^{(i)} \;\sim\; t_{\nu_{nc}-1}(\boldsymbol{\mu}_{nc}, V_c),$$

and the probability of region $R_\ell$ is estimated as the empirical
proportion:

$$\widehat{P}(R_\ell) =
  \frac{1}{n_\mathrm{MC}} \sum_{i=1}^{n_\mathrm{MC}}
  \mathbf{1}\!\left[T_t^{(i)} - T_c^{(i)} \in R_\ell\right].$$

When `pbayespostpred2cont()` is called in vectorised mode over
$n_\mathrm{sim}$ replicates, all $n_\mathrm{MC}$ standard normal and
chi-squared draws are pre-generated once and reused; only the
Cholesky factor of the replicate-specific scale matrix is recomputed
per observation.

### 3.2 Moment-Matching Approximation (`CalcMethod = 'MM'`)

The difference $\boldsymbol{D} = T_t - T_c$ is approximated by a
bivariate non-standardised $t$-distribution
(Yamaguchi et al., 2025, Theorem 3).
Let $V_j$ be the scale matrix of $T_j$ ($j = t, c$), and define

$$A = \left(\frac{\nu_t}{\nu_t - 2} V_t +
            \frac{\nu_c}{\nu_c - 2} V_c\right)^{\!-1},$$

$$Q_m = \frac{1}{p(p+2)}\left[
  \frac{\nu_t^2}{(\nu_t-2)(\nu_t-4)}
  \!\left\{(\mathrm{tr}(AV_t))^2 + 2\,\mathrm{tr}(AV_t AV_t)\right\}
  + \frac{\nu_c^2}{(\nu_c-2)(\nu_c-4)}
  \!\left\{(\mathrm{tr}(AV_c))^2 + 2\,\mathrm{tr}(AV_c AV_c)\right\}
  + \frac{2\nu_t\nu_c}{(\nu_t-2)(\nu_c-2)}
  \!\left\{\mathrm{tr}(AV_t)\,\mathrm{tr}(AV_c) +
            2\,\mathrm{tr}(AV_t AV_c)\right\}
\right]$$

with $p = 2$. Then $\boldsymbol{D} \approx Z^* \sim
t_{\nu^*}(\boldsymbol{\mu}^*, \Sigma^*)$, where

$$\boldsymbol{\mu}^* = \boldsymbol{\mu}_t - \boldsymbol{\mu}_c,$$

$$\nu^* = \frac{2 - 4Q_m}{1 - Q_m},$$

$$\Sigma^* = \left(\frac{\nu_t}{\nu_t - 2} V_t +
  \frac{\nu_c}{\nu_c - 2} V_c\right) \frac{\nu^* - 2}{\nu^*}.$$

The joint probability over each rectangular region is evaluated by a
single call to `mvtnorm::pmvt()`, which avoids simulation entirely and
is exact in the normal limit ($\nu_{nj} \to \infty$).

The MM method requires $\nu_{nj} - 1 > 4$ for both groups ($\nu_j > 4$
in the notation above); if this condition is not met, the function falls
back to MC automatically.

---

## 4. Study Designs

### 4.1 Controlled Design

Both groups are observed in the PoC trial.  All NIW posterior parameters
are updated from the current data.

Posterior probability -- vague prior:

```{r ctrl-post-vague}
S_t <- matrix(c(18.0, 3.6, 3.6, 9.0), 2, 2)
S_c <- matrix(c(16.0, 2.8, 2.8, 8.5), 2, 2)

set.seed(42)
p_post_vague <- pbayespostpred2cont(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  n_t = 20L, n_c = 20L,
  ybar_t = c(3.5, 2.1), S_t = S_t,
  ybar_c = c(1.8, 1.0), S_c = S_c,
  m_t = NULL, m_c = NULL,
  kappa0_t = NULL, nu0_t = NULL, mu0_t = NULL, Lambda0_t = NULL,
  kappa0_c = NULL, nu0_c = NULL, mu0_c = NULL, Lambda0_c = NULL,
  r = NULL,
  ne_t = NULL, ne_c = NULL, alpha0e_t = NULL, alpha0e_c = NULL,
  bar_ye_t = NULL, bar_ye_c = NULL, se_t = NULL, se_c = NULL,
  nMC = 5000L
)
print(round(p_post_vague, 4))
cat(sprintf(
  "g_Go  = P(R1 | data) = %.4f\n", p_post_vague["R1"]
))
cat(sprintf(
  "g_NoGo = P(R9 | data) = %.4f\n\n", p_post_vague["R9"]
))
cat(sprintf(
  "Go  criterion: g_Go  >= gamma1 (0.80)? %s\n",
  ifelse(p_post_vague["R1"] >= 0.80, "YES", "NO")
))
cat(sprintf(
  "NoGo criterion: g_NoGo >= gamma2 (0.20)? %s\n",
  ifelse(p_post_vague["R9"] >= 0.20, "YES", "NO")
))
cat(sprintf("Decision: %s\n",
  ifelse(p_post_vague["R1"] >= 0.80 & p_post_vague["R9"] <  0.20, "Go",
  ifelse(p_post_vague["R1"] <  0.80 & p_post_vague["R9"] >= 0.20, "NoGo",
  ifelse(p_post_vague["R1"] >= 0.80 & p_post_vague["R9"] >= 0.20, "Miss",
                                                                    "Gray")))
))
```

Posterior probability -- NIW informative prior:

```{r ctrl-post-niw}
L0 <- matrix(c(8.0, 0.0, 0.0, 2.0), 2, 2)

set.seed(42)
p_post_niw <- pbayespostpred2cont(
  prob = 'posterior', design = 'controlled', prior = 'N-Inv-Wishart',
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  n_t = 20L, n_c = 20L,
  ybar_t = c(3.5, 2.1), S_t = S_t,
  ybar_c = c(1.8, 1.0), S_c = S_c,
  m_t = NULL, m_c = NULL,
  kappa0_t = 2.0, nu0_t = 5.0, mu0_t = c(2.0, 1.0), Lambda0_t = L0,
  kappa0_c = 2.0, nu0_c = 5.0, mu0_c = c(0.0, 0.0), Lambda0_c = L0,
  r = NULL,
  ne_t = NULL, ne_c = NULL, alpha0e_t = NULL, alpha0e_c = NULL,
  bar_ye_t = NULL, bar_ye_c = NULL, se_t = NULL, se_c = NULL,
  nMC = 5000L
)
print(round(p_post_niw, 4))
```

Posterior predictive probability (future Phase III: $m_t = m_c = 60$):

```{r ctrl-pred}
set.seed(42)
p_pred <- pbayespostpred2cont(
  prob = 'predictive', design = 'controlled', prior = 'vague',
  theta_TV1 = NULL, theta_MAV1 = NULL,
  theta_TV2 = NULL, theta_MAV2 = NULL,
  theta_NULL1 = 0.5, theta_NULL2 = 0.3,
  n_t = 20L, n_c = 20L,
  ybar_t = c(3.5, 2.1), S_t = S_t,
  ybar_c = c(1.8, 1.0), S_c = S_c,
  m_t = 60L, m_c = 60L,
  kappa0_t = NULL, nu0_t = NULL, mu0_t = NULL, Lambda0_t = NULL,
  kappa0_c = NULL, nu0_c = NULL, mu0_c = NULL, Lambda0_c = NULL,
  r = NULL,
  ne_t = NULL, ne_c = NULL, alpha0e_t = NULL, alpha0e_c = NULL,
  bar_ye_t = NULL, bar_ye_c = NULL, se_t = NULL, se_c = NULL,
  nMC = 5000L
)
print(round(p_pred, 4))
cat(sprintf(
  "\nGo  region (R1): P = %.4f  >= gamma1 (0.80)? %s\n",
  p_pred["R1"], ifelse(p_pred["R1"] >= 0.80, "YES", "NO")
))
cat(sprintf(
  "NoGo region (R4): P = %.4f  >= gamma2 (0.20)? %s\n",
  p_pred["R4"], ifelse(p_pred["R4"] >= 0.20, "YES", "NO")
))
cat(sprintf("Decision: %s\n",
  ifelse(p_pred["R1"] >= 0.80 & p_pred["R4"] <  0.20, "Go",
  ifelse(p_pred["R1"] <  0.80 & p_pred["R4"] >= 0.20, "NoGo",
  ifelse(p_pred["R1"] >= 0.80 & p_pred["R4"] >= 0.20, "Miss",
                                                        "Gray")))
))
```

### 4.2 Uncontrolled Design

No concurrent control group is enrolled.  Under the NIW prior, the
hypothetical control distribution is

$$\boldsymbol{\mu}_c \;\sim\;
  t_{\nu_{nt}-1}\!\!\left(\boldsymbol{\mu}_{0c},\;
  r \cdot \frac{\Lambda_{nt}}{\kappa_{nt}(\nu_{nt}-1)}\right),$$

where $\boldsymbol{\mu}_{0c}$ (`mu0_c`) is the assumed control mean,
$r > 0$ (`r`) scales
the covariance relative to the treatment group, and $\Lambda_{nt}$ is the
posterior scale matrix of the treatment group.

```{r unctrl-post}
set.seed(1)
p_unctrl <- pbayespostpred2cont(
  prob = 'posterior', design = 'uncontrolled', prior = 'N-Inv-Wishart',
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  n_t = 20L, n_c = NULL,
  ybar_t = c(3.5, 2.1), S_t = S_t,
  ybar_c = NULL,        S_c = NULL,
  m_t = NULL, m_c = NULL,
  kappa0_t = 2.0, nu0_t = 5.0, mu0_t = c(2.0, 1.0), Lambda0_t = L0,
  kappa0_c = NULL, nu0_c = NULL, mu0_c = c(0.0, 0.0), Lambda0_c = NULL,
  r = 1.0,
  ne_t = NULL, ne_c = NULL, alpha0e_t = NULL, alpha0e_c = NULL,
  bar_ye_t = NULL, bar_ye_c = NULL, se_t = NULL, se_c = NULL,
  nMC = 5000L
)
print(round(p_unctrl, 4))
```

### 4.3 External Design (Power Prior)

External data for group $j$ (sample size $n_{e,j}$, sample mean
$\bar{\mathbf{y}}_{e,j}$, sum-of-squares matrix $S_{e,j}$) are incorporated via
a power prior with weight $\alpha_{0e,j} \in (0,1]$.
Both `prior = 'vague'` and `prior = 'N-Inv-Wishart'` are supported.

#### Vague prior (`prior = 'vague'`)

The posterior parameters after combining the vague power prior with the
PoC data are given by Corollary 2 of Huang et al. (2025):

$$\boldsymbol{\mu}_{nj}^* =
  \frac{\alpha_{0e,j}\,n_{e,j}\,\bar{\mathbf{y}}_{e,j} + n_j\,\bar{\mathbf{y}}_j}
       {\kappa_{nj}^*},$$

$$\kappa_{nj}^* = \alpha_{0e,j}\,n_{e,j} + n_j, \qquad
  \nu_{nj}^* = \alpha_{0e,j}\,n_{e,j} + n_j - 1,$$

$$\Lambda_{nj}^* = \alpha_{0e,j}\,S_{e,j} + S_j +
  \frac{\alpha_{0e,j}\,n_{e,j}\,n_j}{\kappa_{nj}^*}
  (\bar{\mathbf{y}}_j - \bar{\mathbf{y}}_{e,j})
  (\bar{\mathbf{y}}_j - \bar{\mathbf{y}}_{e,j})^\top.$$

The marginal posterior of $\boldsymbol{\mu}_j$ is then

$$\boldsymbol{\mu}_j \mid \mathbf{Y}_j \;\sim\;
  t_{\nu_{nj}^* - 1}\!\!\left(\boldsymbol{\mu}_{nj}^*,\;
  \frac{\Lambda_{nj}^*}{\kappa_{nj}^*(\nu_{nj}^* - 1)}\right).$$

```{r ext-post-vague}
Se2_ext <- matrix(c(15.0, 2.5, 2.5, 7.5), 2, 2)

set.seed(2)
p_ext_vague <- pbayespostpred2cont(
  prob = 'posterior', design = 'external', prior = 'vague',
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  n_t = 20L, n_c = 20L,
  ybar_t = c(3.5, 2.1), S_t = S_t,
  ybar_c = c(1.8, 1.0), S_c = S_c,
  m_t = NULL, m_c = NULL,
  kappa0_t = NULL, nu0_t = NULL, mu0_t = NULL, Lambda0_t = NULL,
  kappa0_c = NULL, nu0_c = NULL, mu0_c = NULL, Lambda0_c = NULL,
  r = NULL,
  ne_t = NULL, ne_c = 10L, alpha0e_t = NULL, alpha0e_c = 0.5,
  bar_ye_t = NULL, bar_ye_c = c(1.5, 0.8), se_t = NULL, se_c = Se2_ext,
  nMC = 5000L
)
print(round(p_ext_vague, 4))
```

#### N-Inv-Wishart prior (`prior = 'N-Inv-Wishart'`)

The power prior first combines the initial NIW prior with the discounted
external data (Theorem 5 of Huang et al., 2025):

$$\boldsymbol{\mu}_{e,j} =
  \frac{\alpha_{0e,j}\,n_{e,j}\,\bar{\mathbf{y}}_{e,j} +
        \kappa_{0j}\,\boldsymbol{\mu}_{0j}}
       {\kappa_{e,j}},$$

$$\kappa_{e,j} = \alpha_{0e,j}\,n_{e,j} + \kappa_{0j}, \qquad
  \nu_{e,j} = \alpha_{0e,j}\,n_{e,j} + \nu_{0j},$$

$$\Lambda_{e,j} = \alpha_{0e,j}\,S_{e,j} + \Lambda_{0j} +
  \frac{\kappa_{0j}\,\alpha_{0e,j}\,n_{e,j}}{\kappa_{e,j}}
  (\boldsymbol{\mu}_{0j} - \bar{\mathbf{y}}_{e,j})
  (\boldsymbol{\mu}_{0j} - \bar{\mathbf{y}}_{e,j})^\top.$$

This intermediate NIW result is then updated with the current PoC data
by standard conjugate updating (Theorem 6 of Huang et al., 2025):

$$\boldsymbol{\mu}_{nj}^* =
  \frac{\kappa_{e,j}\,\boldsymbol{\mu}_{e,j} + n_j\,\bar{\mathbf{y}}_j}
       {\kappa_{nj}^*},$$

$$\kappa_{nj}^* = \kappa_{e,j} + n_j, \qquad
  \nu_{nj}^* = \nu_{e,j} + n_j,$$

$$\Lambda_{nj}^* = \Lambda_{e,j} + S_j +
  \frac{\kappa_{e,j}\,n_j}{\kappa_{nj}^*}
  (\bar{\mathbf{y}}_j - \boldsymbol{\mu}_{e,j})
  (\bar{\mathbf{y}}_j - \boldsymbol{\mu}_{e,j})^\top.$$

The marginal posterior of $\boldsymbol{\mu}_j$ is then

$$\boldsymbol{\mu}_j \mid \mathbf{Y}_j \;\sim\;
  t_{\nu_{nj}^* - 1}\!\!\left(\boldsymbol{\mu}_{nj}^*,\;
  \frac{\Lambda_{nj}^*}{\kappa_{nj}^*(\nu_{nj}^* - 1)}\right).$$

```{r ext-post}
set.seed(3)
p_ext <- pbayespostpred2cont(
  prob = 'posterior', design = 'external', prior = 'N-Inv-Wishart',
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  n_t = 20L, n_c = 20L,
  ybar_t = c(3.5, 2.1), S_t = S_t,
  ybar_c = c(1.8, 1.0), S_c = S_c,
  m_t = NULL, m_c = NULL,
  kappa0_t = 2.0, nu0_t = 5.0, mu0_t = c(2.0, 1.0), Lambda0_t = L0,
  kappa0_c = 2.0, nu0_c = 5.0, mu0_c = c(0.0, 0.0), Lambda0_c = L0,
  r = NULL,
  ne_t = NULL, ne_c = 10L, alpha0e_t = NULL, alpha0e_c = 0.5,
  bar_ye_t = NULL, bar_ye_c = c(1.5, 0.8), se_t = NULL, se_c = Se2_ext,
  nMC = 5000L
)
print(round(p_ext, 4))
```

---

## 5. Operating Characteristics

### 5.1 Definition

For given true parameters $(\boldsymbol{\mu}_t, \Sigma_t,
\boldsymbol{\mu}_c, \Sigma_c)$ (`mu_t`, `Sigma_t`, `mu_c`, `Sigma_c`),
the operating characteristics are
estimated by Monte Carlo over $n_\mathrm{sim}$ replicates (`nsim`).  Let
$\hat{g}_{\mathrm{Go},i} = P(\text{GoRegions} \mid \mathbf{Y}^{(i)})$
and $\hat{g}_{\mathrm{NoGo},i} = P(\text{NoGoRegions} \mid \mathbf{Y}^{(i)})$
for the $i$-th simulated dataset.  Then:

$$\widehat{\Pr}(\mathrm{Go}) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},i} \ge \gamma_1 \;\text{and}\;
  \hat{g}_{\mathrm{NoGo},i} < \gamma_2\right],$$

$$\widehat{\Pr}(\mathrm{NoGo}) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{NoGo},i} \ge \gamma_2 \;\text{and}\;
  \hat{g}_{\mathrm{Go},i} < \gamma_1\right],$$

$$\widehat{\Pr}(\mathrm{Miss}) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},i} \ge \gamma_1 \;\text{and}\;
  \hat{g}_{\mathrm{NoGo},i} \ge \gamma_2\right],$$

$$\widehat{\Pr}(\mathrm{Gray}) = 1 -
  \widehat{\Pr}(\mathrm{Go}) - \widehat{\Pr}(\mathrm{NoGo}) -
  \widehat{\Pr}(\mathrm{Miss}).$$

Here $\gamma_1$ and $\gamma_2$ correspond to arguments `gamma_go` and
`gamma_nogo`, respectively.
A Miss ($\hat{g}_{\mathrm{Go}} \ge \gamma_1$ and
$\hat{g}_{\mathrm{NoGo}} \ge \gamma_2$ simultaneously) indicates an
inconsistent threshold configuration and triggers an error by default
(`error_if_Miss = TRUE`).

### 5.2 Example: Controlled Design, Posterior Probability

Treatment effect scenarios are defined over a two-dimensional grid of
$(\mu_{t1}, \mu_{t2})$ while fixing
$\boldsymbol{\mu}_c = (0, 0)^\top$.
The full factorial combination of unique $\mu_{t1}$ and $\mu_{t2}$ values
triggers the tile plot in `plot()`, with colour intensity encoding
$\Pr(\mathrm{Go})$:

```{r oc-controlled, fig.width = 8, fig.height = 6}
Sigma <- matrix(c(4.0, 0.8, 0.8, 1.0), 2, 2)

mu_t1_seq <- seq(0.0, 3.5, by = 0.5)
mu_t2_seq <- seq(0.0, 2.1, by = 0.3)
n_scen    <- length(mu_t1_seq) * length(mu_t2_seq)

oc_ctrl <- pbayesdecisionprob2cont(
  nsim = 100L, prob = 'posterior', design = 'controlled',
  prior = 'vague',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma_go = 0.80, gamma_nogo = 0.20,
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  n_t = 20L, n_c = 20L, m_t = NULL, m_c = NULL,
  mu_t    = cbind(rep(mu_t1_seq, times = length(mu_t2_seq)),
                  rep(mu_t2_seq, each  = length(mu_t1_seq))),
  Sigma_t = Sigma,
  mu_c    = matrix(0, nrow = n_scen, ncol = 2),
  Sigma_c = Sigma,
  kappa0_t = NULL, nu0_t = NULL, mu0_t = NULL, Lambda0_t = NULL,
  kappa0_c = NULL, nu0_c = NULL, mu0_c = NULL, Lambda0_c = NULL,
  r = NULL,
  ne_t = NULL, ne_c = NULL, alpha0e_t = NULL, alpha0e_c = NULL,
  bar_ye_t = NULL, bar_ye_c = NULL, se_t = NULL, se_c = NULL,
  nMC = 500L, CalcMethod = 'MC',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE, seed = 42L
)
print(oc_ctrl)
plot(oc_ctrl, base_size = 20)
```

The same function supports `design = 'uncontrolled'` (with hypothetical
control mean `mu0_c` and scale factor `r`) and `design = 'external'`
(with power prior arguments `ne_c`, `bar_ye_c`, `se_c`, `alpha0e_c`).
The argument `prob = 'predictive'` switches to predictive probability
(with future sample sizes `m_t`, `m_c` and null thresholds
`theta_NULL1`, `theta_NULL2`). The function signature and output format
are identical across all combinations.

---

## 6. Optimal Threshold Search

### 6.1 Objective and Algorithm

Because there are two decision thresholds $(\gamma_1, \gamma_2)$, the
optimal search is performed over a two-dimensional grid
$\Gamma_1 \times \Gamma_2$ (arguments `gamma_go_grid` and `gamma_nogo_grid`).
The two thresholds are calibrated independently under separate scenarios:

- $\gamma_1^*$: calibrated under the Go-calibration scenario
  (`mu_t_go`, `Sigma_t_go`, `mu_c_go`, `Sigma_c_go`; typically the
  null scenario $\boldsymbol{\mu}_t = \boldsymbol{\mu}_c$), so that
  $\Pr(\mathrm{Go}) < \texttt{target\_go}$.
- $\gamma_2^*$: calibrated under the NoGo-calibration scenario
  (`mu_t_nogo`, `Sigma_t_nogo`, `mu_c_nogo`, `Sigma_c_nogo`; typically
  the alternative scenario), so that
  $\Pr(\mathrm{NoGo}) < \texttt{target\_nogo}$.

`getgamma2cont()` uses a three-stage strategy:

Stage 1 (Simulate and precompute): $n_\mathrm{sim}$ bivariate datasets
are generated separately from each calibration scenario.  For each
replicate $i$, `pbayespostpred2cont()` is called once in vectorised
mode to return the full region probability vector.  The marginal Go and
NoGo probabilities per replicate are:

$$\hat{g}_{\mathrm{Go},i} = \sum_{\ell \in \text{GoRegions}} \hat{p}_{i\ell}, \qquad
  \hat{g}_{\mathrm{NoGo},i} = \sum_{\ell \in \text{NoGoRegions}} \hat{p}_{i\ell}.$$

Stage 2 (Gamma sweep): For each candidate $\gamma_1 \in \Gamma_1$
(under the Go-calibration scenario) and each $\gamma_2 \in \Gamma_2$
(under the NoGo-calibration scenario):

$$\Pr(\mathrm{Go} \mid \gamma_1) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},i} \ge \gamma_1\right],$$

$$\Pr(\mathrm{NoGo} \mid \gamma_2) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{NoGo},i} \ge \gamma_2\right].$$

The results are stored as vectors `PrGo_grid` and `PrNoGo_grid` indexed
by `gamma_go_grid` and `gamma_nogo_grid`, respectively.

Stage 3 (Optimal selection): $\gamma_1^*$ is the smallest grid value
satisfying $\Pr(\mathrm{Go} \mid \gamma_1) < \texttt{target\_go}$;
$\gamma_2^*$ is the smallest grid value satisfying
$\Pr(\mathrm{NoGo} \mid \gamma_2) < \texttt{target\_nogo}$.

### 6.2 Example: Controlled Design, Posterior Probability

Go-calibration scenario: $\boldsymbol{\mu}_t = (0.5, 0.3)^\top$,
$\boldsymbol{\mu}_c = (0, 0)^\top$ (effect at MAV boundary -- above which
Go is not yet warranted, so $\Pr(\mathrm{Go}) < \texttt{target\_go}$ is
required). NoGo-calibration scenario: $\boldsymbol{\mu}_t = (1.0, 0.6)^\top$,
$\boldsymbol{\mu}_c = (0, 0)^\top$ (effect between MAV and TV -- above
which NoGo should be unlikely).

```{r getgamma-ctrl, fig.width = 8, fig.height = 6}
res_gamma <- getgamma2cont(
  nsim = 500L, prob = 'posterior', design = 'controlled',
  prior = 'vague',
  GoRegions = 1L, NoGoRegions = 9L,
  mu_t_go    = c(0.5, 0.3), Sigma_t_go    = Sigma,
  mu_c_go    = c(0.0, 0.0), Sigma_c_go    = Sigma,
  mu_t_nogo  = c(1.0, 0.6), Sigma_t_nogo  = Sigma,
  mu_c_nogo  = c(0.0, 0.0), Sigma_c_nogo  = Sigma,
  target_go = 0.05, target_nogo = 0.20,
  n_t = 20L, n_c = 20L,
  theta_TV1  = 1.5, theta_MAV1 = 0.5,
  theta_TV2  = 1.0, theta_MAV2 = 0.3,
  theta_NULL1 = NULL, theta_NULL2 = NULL,
  m_t = NULL, m_c = NULL,
  kappa0_t = NULL, nu0_t = NULL, mu0_t = NULL, Lambda0_t = NULL,
  kappa0_c = NULL, nu0_c = NULL, mu0_c = NULL, Lambda0_c = NULL,
  r = NULL,
  ne_t = NULL, ne_c = NULL, alpha0e_t = NULL, alpha0e_c = NULL,
  bar_ye_t = NULL, bar_ye_c = NULL, se_t = NULL, se_c = NULL,
  nMC = 500L, CalcMethod = 'MC',
  gamma_go_grid   = seq(0.05, 0.95, by = 0.05),
  gamma_nogo_grid = seq(0.05, 0.95, by = 0.05),
  seed = 42L
)
plot(res_gamma, base_size = 20)
```

The same function supports `design = 'uncontrolled'` (with `mu0_c` and
`r`) and `design = 'external'` (with power prior arguments `ne_c`,
`bar_ye_c`, `se_c`, `alpha0e_c`). Setting `prob = 'predictive'`
switches to predictive probability calibration (with `theta_NULL1`,
`theta_NULL2`, `m_t`, `m_c`). The function signature is identical
across all combinations.

---

## 7. Summary

This vignette covered two continuous endpoint analysis in BayesianQDM:

1. Bayesian model: NIW conjugate posterior with explicit update
   formulae for $\kappa_{nj}$, $\nu_{nj}$, $\boldsymbol{\mu}_{nj}$,
   and $\Lambda_{nj}$; bivariate non-standardised $t$ marginal posterior.
2. Nine-region grid for posterior probability and four-region grid
   for predictive probability; typical choice Go = R1, NoGo = R9 (posterior)
   or R4 (predictive).
3. Two computation methods: MC (vectorised pre-generated draws with
   Cholesky reuse) and MM (Mahalanobis-distance fourth-moment matching
   via Yamaguchi et al. (2025), Theorem 3, + `mvtnorm::pmvt` for joint
   probability; requires $\nu_j > 4$).
4. Three study designs: controlled, uncontrolled ($\boldsymbol{\mu}_{0c}$
   and $r$ fixed), external design (vague power prior follows Corollary 2
   of Huang et al. (2025); NIW power prior follows Theorems 5--6).
5. Operating characteristics: Monte Carlo estimation with vectorised
   region probability computation.
6. Optimal threshold search: three-stage strategy in `getgamma2cont()`
   calibrating $\gamma_1^*$ and $\gamma_2^*$ independently under separate
   Go- and NoGo-calibration scenarios, with calibration curves visualised
   via `plot()` and optimal thresholds marked at the target probability levels
   ($\Pr(\mathrm{Go}) < 0.05$ under null, $\Pr(\mathrm{NoGo}) < 0.20$
   under alternative).

See `vignette("two-binary")` for the analogous two binary endpoint analysis.
