---
title: "Two Continuous Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two Continuous Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 5.5
)
library(BayesianQDM)
```

## Motivating Scenario

We extend the single-endpoint RA trial to a **bivariate** design with two
co-primary continuous endpoints:

- **Endpoint 1**: change from baseline in DAS28 score
- **Endpoint 2**: change from baseline in HAQ-DI (Health Assessment
  Questionnaire – Disability Index)

The trial enrolls $n_1 = n_2 = 20$ patients per arm (treatment vs. placebo)
in a 1:1 randomised controlled design.

**Clinically meaningful thresholds** (posterior probability):

|                              | Endpoint 1 | Endpoint 2 |
|------------------------------|:----------:|:----------:|
| $\theta_{\mathrm{TV}}$       | 1.5        | 1.0        |
| $\theta_{\mathrm{MAV}}$      | 0.5        | 0.3        |

**Null hypothesis thresholds** (predictive probability):
$\theta_{\mathrm{NULL},1} = 0.5$, $\theta_{\mathrm{NULL},2} = 0.3$.

**Decision thresholds**: $\gamma_1 = 0.80$ (Go), $\gamma_2 = 0.20$ (NoGo).

**Observed data** (used in Sections 2–4):

- Treatment: $\bar{\mathbf{y}}_1 = (3.5, 2.1)^\top$,
  $S_1 = \begin{pmatrix}18.0 & 3.6\\3.6 & 9.0\end{pmatrix}$
- Control: $\bar{\mathbf{y}}_2 = (1.8, 1.0)^\top$,
  $S_2 = \begin{pmatrix}16.0 & 2.8\\2.8 & 8.5\end{pmatrix}$

where $S_k = \sum_{i=1}^{n_k}(\mathbf{y}_{ki} - \bar{\mathbf{y}}_k)
(\mathbf{y}_{ki} - \bar{\mathbf{y}}_k)^\top$ is the sum-of-squares matrix.

---

## 1. Bayesian Model: Normal-Inverse-Wishart Conjugate

### 1.1 Prior Distribution

For group $k$ ($k = t$: treatment, $k = c$: control), we model the
bivariate outcomes as

$$\mathbf{y}_{ki} \mid \boldsymbol{\mu}_k, \Sigma_k \;\overset{iid}{\sim}\;
  \mathcal{N}_2(\boldsymbol{\mu}_k,\, \Sigma_k).$$

The conjugate prior is the Normal-Inverse-Wishart (NIW) distribution:

$$(\boldsymbol{\mu}_k, \Sigma_k) \;\sim\;
  \mathrm{NIW}(\boldsymbol{\mu}_{0k},\, \kappa_{0k},\, \nu_{0k},\,
  \Lambda_{0k}),$$

where $\boldsymbol{\mu}_{0k} \in \mathbb{R}^2$ is the prior mean,
$\kappa_{0k} > 0$ the prior precision (pseudo-sample size),
$\nu_{0k} > 3$ the prior degrees of freedom, and
$\Lambda_{0k}$ a $2 \times 2$ positive-definite prior scale matrix.

The **vague (Jeffreys) prior** is obtained by letting
$\kappa_{0k} \to 0$ and $\nu_{0k} \to 0$.

### 1.2 Posterior Distribution

Given $n_k$ observations with sample mean $\bar{\mathbf{y}}_k$ and
sum-of-squares matrix $S_k$, the NIW posterior has updated parameters:

$$\kappa_{nk} = \kappa_{0k} + n_k, \qquad
  \nu_{nk} = \nu_{0k} + n_k,$$

$$\boldsymbol{\mu}_{nk} =
  \frac{\kappa_{0k}\,\boldsymbol{\mu}_{0k} + n_k\,\bar{\mathbf{y}}_k}
       {\kappa_{nk}},$$

$$\Lambda_{nk} = \Lambda_{0k} + S_k +
  \frac{\kappa_{0k}\,n_k}{\kappa_{nk}}
  (\bar{\mathbf{y}}_k - \boldsymbol{\mu}_{0k})
  (\bar{\mathbf{y}}_k - \boldsymbol{\mu}_{0k})^\top.$$

The marginal posterior of the group mean $\boldsymbol{\mu}_k$ is a
bivariate non-standardised $t$-distribution:

$$\boldsymbol{\mu}_k \mid \mathbf{Y}_k \;\sim\;
  t_{\nu_{nk} - 1}\!\!\left(\boldsymbol{\mu}_{nk},\;
  \frac{\Lambda_{nk}}{\kappa_{nk}(\nu_{nk} - 1)}\right).$$

Under the **vague prior**, this simplifies to

$$\boldsymbol{\mu}_k \mid \mathbf{Y}_k \;\sim\;
  t_{n_k - 2}\!\!\left(\bar{\mathbf{y}}_k,\;
  \frac{S_k}{n_k(n_k - 2)}\right).$$

### 1.3 Posterior of the Bivariate Treatment Effect

The bivariate treatment effect is $\boldsymbol{\theta} =
\boldsymbol{\mu}_t - \boldsymbol{\mu}_c$.  Since the two groups are
independent, the posterior of $\boldsymbol{\theta}$ is the distribution
of the difference of two independent bivariate $t$-vectors:

$$T_k \;\sim\; t_{\nu_{nk}-1}(\boldsymbol{\mu}_{nk},\, V_k),$$

where the posterior scale matrix is
$V_k = \Lambda_{nk} / [\kappa_{nk}(\nu_{nk}-1)]$.

The posterior probability that $\boldsymbol{\theta}$ falls in a
rectangular region $R$ defined by thresholds on each component is

$$P(\boldsymbol{\theta} \in R \mid \mathbf{Y}) =
  P(T_t - T_c \in R),$$

computed by one of two methods (MC or MM) described in Section 3.

### 1.4 Nine-Region Grid (Posterior Probability)

The bivariate threshold grid for $(\theta_1, \theta_2)$ creates the same
$3 \times 3$ partition as for the two binary endpoint case:

```{r nine-region-cont, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Nine-region grid for two-endpoint posterior probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="3" style="border:1px solid #aaa; padding:6px;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &gt; &theta;<sub>TV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>TV1</sub> &ge; &theta;<sub>1</sub> &gt; &theta;<sub>MAV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>MAV1</sub> &ge; &theta;<sub>1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="3" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg); font-weight:bold;">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>TV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
    <td style="border:1px solid #aaa; padding:6px;">R7</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>TV2</sub> &ge; &theta;<sub>2</sub> &gt; &theta;<sub>MAV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R5</td>
    <td style="border:1px solid #aaa; padding:6px;">R8</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>MAV2</sub> &ge; &theta;<sub>2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
    <td style="border:1px solid #aaa; padding:6px;">R6</td>
    <td style="border:1px solid #aaa; padding:6px;">R9</td>
  </tr>
</tbody>
</table>
')
```

Typical choices: **Go if $P(R1) \ge \gamma_1$**,
**NoGo if $P(R9) \ge \gamma_2$**.

### 1.5 Four-Region Grid (Predictive Probability)

For the predictive probability, a $2 \times 2$ partition is used:

```{r four-region-cont, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Four-region grid for two-endpoint predictive probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="2" style="border:1px solid #aaa; padding:6px;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &gt; &theta;<sub>NULL1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &le; &theta;<sub>NULL1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="2" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg); font-weight:bold;">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &le; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
  </tr>
</tbody>
</table>
')
```

Typical choices: **Go if $P(R1) \ge \gamma_1$**,
**NoGo if $P(R4) \ge \gamma_2$**.

---

## 2. Posterior Predictive Distribution

The predictive distribution of the future sample mean
$\bar{\tilde{\mathbf{y}}}_k$ based on $m_k$ future patients is again a
bivariate $t$-distribution:

$$\bar{\tilde{\mathbf{y}}}_k \mid \mathbf{Y}_k \;\sim\;
  t_{\nu_{nk}-1}\!\!\left(\boldsymbol{\mu}_{nk},\;
  \frac{\Lambda_{nk}}{\kappa_{nk}(\nu_{nk}-1)}
  \cdot \frac{\kappa_{nk} + 1}{\kappa_{nk}} \cdot \frac{1}{m_k}
  \right)$$

under the NIW prior.  Under the vague prior the scale inflates to
$S_k(n_k + 1) / [n_k^2(n_k - 2) m_k]$.

---

## 3. Two Computation Methods

### 3.1 Monte Carlo Simulation (MC)

For each of $n_\mathrm{MC}$ draws, independent samples are generated:

$$T_t^{(i)} \;\sim\; t_{\nu_{nt}-1}(\boldsymbol{\mu}_{nt}, V_t), \qquad
  T_c^{(i)} \;\sim\; t_{\nu_{nc}-1}(\boldsymbol{\mu}_{nc}, V_c),$$

and the probability of region $R_j$ is estimated as the empirical
proportion:

$$\widehat{P}(R_j) =
  \frac{1}{n_\mathrm{MC}} \sum_{i=1}^{n_\mathrm{MC}}
  \mathbf{1}\!\left[T_t^{(i)} - T_c^{(i)} \in R_j\right].$$

When `pbayespostpred2cont()` is called in vectorised mode over
$n_\mathrm{sim}$ replicates, all $n_\mathrm{MC}$ standard normal and
chi-squared draws are pre-generated once and reused; only the
Cholesky factor of the replicate-specific scale matrix is recomputed
per observation.

### 3.2 Moment-Matching Approximation (MM)

For each marginal component $\ell = 1, 2$, the difference
$D_\ell = T_{t,\ell} - T_{c,\ell}$ is approximated by a univariate
$t_{\nu_\ell^*}(\mu_\ell^*, \sigma_\ell^*)$ using the same fourth-moment
equations as for the single-endpoint case (see
`vignette("single-continuous")`).

The joint probability over the rectangular region $R$ is then evaluated
using the bivariate $t$-distribution from `mvtnorm::pmvt()`, treating
the two approximate marginals as jointly $t$-distributed with a
correlation derived from the posterior covariance matrices.  This avoids
simulation entirely and is exact in the normal limit
($\nu_{nk} \to \infty$).

The MM method requires $\nu_{nk} - 1 > 4$ for both arms; if this
condition is not met, the function falls back to MC automatically.

---

## 4. Study Designs

### 4.1 Controlled Design

Both arms are observed in the PoC trial.  All NIW posterior parameters
are updated from the current data.

**Posterior probability — vague prior:**

```{r ctrl-post-vague}
S1 <- matrix(c(18.0, 3.6, 3.6, 9.0), 2, 2)
S2 <- matrix(c(16.0, 2.8, 2.8, 8.5), 2, 2)

set.seed(42)
p_post_vague <- pbayespostpred2cont(
  prob = 'posterior', design = 'controlled', prior = 'vague',
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = 20L,
  ybar1 = c(3.5, 2.1), S1 = S1,
  ybar2 = c(1.8, 1.0), S2 = S2,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 5000L
)
print(round(p_post_vague, 4))
cat(sprintf(
  "g_Go  = P(R1 | data) = %.4f\n", p_post_vague["R1"]
))
cat(sprintf(
  "g_NoGo = P(R9 | data) = %.4f\n\n", p_post_vague["R9"]
))
cat(sprintf(
  "Go  criterion: g_Go  >= gamma1 (0.80)? %s\n",
  ifelse(p_post_vague["R1"] >= 0.80, "YES", "NO")
))
cat(sprintf(
  "NoGo criterion: g_NoGo >= gamma2 (0.20)? %s\n",
  ifelse(p_post_vague["R9"] >= 0.20, "YES", "NO")
))
cat(sprintf("Decision: %s\n",
  ifelse(p_post_vague["R1"] >= 0.80 & p_post_vague["R9"] <  0.20, "Go",
  ifelse(p_post_vague["R1"] <  0.80 & p_post_vague["R9"] >= 0.20, "NoGo",
  ifelse(p_post_vague["R1"] >= 0.80 & p_post_vague["R9"] >= 0.20, "Miss",
                                                                    "Gray")))
))
```

**Posterior probability — NIW informative prior:**

```{r ctrl-post-niw}
L0 <- matrix(c(8.0, 0.0, 0.0, 2.0), 2, 2)

set.seed(42)
p_post_niw <- pbayespostpred2cont(
  prob = 'posterior', design = 'controlled', prior = 'N-Inv-Wishart',
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = 20L,
  ybar1 = c(3.5, 2.1), S1 = S1,
  ybar2 = c(1.8, 1.0), S2 = S2,
  m1 = NULL, m2 = NULL,
  kappa01 = 2.0, nu01 = 5.0, mu01 = c(2.0, 1.0), Lambda01 = L0,
  kappa02 = 2.0, nu02 = 5.0, mu02 = c(0.0, 0.0), Lambda02 = L0,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 5000L
)
print(round(p_post_niw, 4))
```

**Posterior predictive probability** (future Phase III: $m_1 = m_2 = 60$):

```{r ctrl-pred}
set.seed(42)
p_pred <- pbayespostpred2cont(
  prob = 'predictive', design = 'controlled', prior = 'vague',
  theta.TV1 = NULL, theta.MAV1 = NULL,
  theta.TV2 = NULL, theta.MAV2 = NULL,
  theta.NULL1 = 0.5, theta.NULL2 = 0.3,
  n1 = 20L, n2 = 20L,
  ybar1 = c(3.5, 2.1), S1 = S1,
  ybar2 = c(1.8, 1.0), S2 = S2,
  m1 = 60L, m2 = 60L,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 5000L
)
print(round(p_pred, 4))
cat(sprintf(
  "\nGo  region (R1): P = %.4f  >= gamma1 (0.80)? %s\n",
  p_pred["R1"], ifelse(p_pred["R1"] >= 0.80, "YES", "NO")
))
cat(sprintf(
  "NoGo region (R4): P = %.4f  >= gamma2 (0.20)? %s\n",
  p_pred["R4"], ifelse(p_pred["R4"] >= 0.20, "YES", "NO")
))
cat(sprintf("Decision: %s\n",
  ifelse(p_pred["R1"] >= 0.80 & p_pred["R4"] <  0.20, "Go",
  ifelse(p_pred["R1"] <  0.80 & p_pred["R4"] >= 0.20, "NoGo",
  ifelse(p_pred["R1"] >= 0.80 & p_pred["R4"] >= 0.20, "Miss",
                                                        "Gray")))
))
```

### 4.2 Uncontrolled Design

No concurrent control arm is enrolled.  Under the NIW prior, the
hypothetical control distribution is

$$\boldsymbol{\mu}_c \;\sim\;
  t_{\nu_{nt}-1}\!\!\left(\boldsymbol{\mu}_{0c},\;
  r \cdot \frac{\Lambda_{nt}}{\kappa_{nt}(\nu_{nt}-1)}\right),$$

where $\boldsymbol{\mu}_{0c}$ is the assumed control mean, $r > 0$ scales
the covariance relative to the treatment arm, and $\Lambda_{nt}$ is the
posterior scale matrix of the treatment arm.

```{r unctrl-post}
set.seed(1)
p_unctrl <- pbayespostpred2cont(
  prob = 'posterior', design = 'uncontrolled', prior = 'N-Inv-Wishart',
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = NULL,
  ybar1 = c(3.5, 2.1), S1 = S1,
  ybar2 = NULL,        S2 = NULL,
  m1 = NULL, m2 = NULL,
  kappa01 = 2.0, nu01 = 5.0, mu01 = c(2.0, 1.0), Lambda01 = L0,
  kappa02 = NULL, nu02 = NULL, mu02 = c(0.0, 0.0), Lambda02 = NULL,
  r = 1.0,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 5000L
)
print(round(p_unctrl, 4))
```

### 4.3 External Control Design (Power Prior)

External data for the control arm (sample size $n_{e2}$, sample mean
$\bar{\mathbf{y}}_{e2}$, sum-of-squares $S_{e2}$) are incorporated via
a power prior with weight $\alpha_{e2} \in (0,1]$.

Under the NIW conjugate formulation, the augmented prior for the control
arm is first constructed from the external data:

$$\kappa_{0c}^{\mathrm{aug}} = \kappa_{0c} + \alpha_{e2}\,n_{e2}, \qquad
  \nu_{0c}^{\mathrm{aug}} = \nu_{0c} + \alpha_{e2}\,n_{e2},$$

$$\boldsymbol{\mu}_{0c}^{\mathrm{aug}} =
  \frac{\kappa_{0c}\,\boldsymbol{\mu}_{0c} +
        \alpha_{e2}\,n_{e2}\,\bar{\mathbf{y}}_{e2}}
       {\kappa_{0c}^{\mathrm{aug}}},$$

$$\Lambda_{0c}^{\mathrm{aug}} = \Lambda_{0c} + \alpha_{e2}\,S_{e2} +
  \frac{\kappa_{0c}\,\alpha_{e2}\,n_{e2}}
       {\kappa_{0c}^{\mathrm{aug}}}
  (\bar{\mathbf{y}}_{e2} - \boldsymbol{\mu}_{0c})
  (\bar{\mathbf{y}}_{e2} - \boldsymbol{\mu}_{0c})^\top.$$

These augmented parameters replace $(\kappa_{0c}, \nu_{0c},
\boldsymbol{\mu}_{0c}, \Lambda_{0c})$ in the posterior update.

```{r ext-post}
Se2_ext <- matrix(c(15.0, 2.5, 2.5, 7.5), 2, 2)

set.seed(2)
p_ext <- pbayespostpred2cont(
  prob = 'posterior', design = 'external', prior = 'N-Inv-Wishart',
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = 20L,
  ybar1 = c(3.5, 2.1), S1 = S1,
  ybar2 = c(1.8, 1.0), S2 = S2,
  m1 = NULL, m2 = NULL,
  kappa01 = 2.0, nu01 = 5.0, mu01 = c(2.0, 1.0), Lambda01 = L0,
  kappa02 = 2.0, nu02 = 5.0, mu02 = c(0.0, 0.0), Lambda02 = L0,
  r = NULL,
  ne1 = NULL, ne2 = 10L, alpha01e = NULL, alpha02e = 0.5,
  ybar_e1 = NULL, ybar_e2 = c(1.5, 0.8), Se1 = NULL, Se2 = Se2_ext,
  nMC = 5000L
)
print(round(p_ext, 4))
```

---

## 5. Operating Characteristics

### 5.1 Definition

For given true parameters $(\boldsymbol{\mu}_1, \Sigma_1,
\boldsymbol{\mu}_2, \Sigma_2)$, the operating characteristics are
estimated by Monte Carlo over $n_\mathrm{sim}$ replicates.  Let
$\hat{g}_{\mathrm{Go},i} = P(\text{GoRegions} \mid \mathbf{Y}^{(i)})$
and $\hat{g}_{\mathrm{NoGo},i} = P(\text{NoGoRegions} \mid \mathbf{Y}^{(i)})$
for the $i$-th simulated dataset.  Then:

$$\widehat{\Pr}(\mathrm{Go}) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},i} \ge \gamma_1 \;\text{and}\;
  \hat{g}_{\mathrm{NoGo},i} < \gamma_2\right],$$

$$\widehat{\Pr}(\mathrm{NoGo}) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{NoGo},i} \ge \gamma_2 \;\text{and}\;
  \hat{g}_{\mathrm{Go},i} < \gamma_1\right],$$

$$\widehat{\Pr}(\mathrm{Miss}) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},i} \ge \gamma_1 \;\text{and}\;
  \hat{g}_{\mathrm{NoGo},i} \ge \gamma_2\right],$$

$$\widehat{\Pr}(\mathrm{Gray}) = 1 -
  \widehat{\Pr}(\mathrm{Go}) - \widehat{\Pr}(\mathrm{NoGo}) -
  \widehat{\Pr}(\mathrm{Miss}).$$

A **Miss** ($\hat{g}_{\mathrm{Go}} \ge \gamma_1$ and
$\hat{g}_{\mathrm{NoGo}} \ge \gamma_2$ simultaneously) indicates an
inconsistent threshold configuration and triggers an error by default
(`error_if_Miss = TRUE`).

### 5.2 Controlled Design — Multiple Scenarios

Treatment effect scenarios are defined by varying $\boldsymbol{\mu}_1$
while fixing $\boldsymbol{\mu}_2 = (0, 0)^\top$:

```{r oc-controlled}
Sigma <- matrix(c(4.0, 0.8, 0.8, 1.0), 2, 2)
L0oc  <- matrix(c(8.0, 0.0, 0.0, 2.0), 2, 2)

oc_ctrl <- pbayesdecisionprob2cont(
  nsim = 300L, prob = 'posterior', design = 'controlled',
  prior = 'vague',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.20,
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = 20L, m1 = NULL, m2 = NULL,
  mu1 = rbind(c(0.5, 0.3), c(1.5, 1.0), c(2.5, 1.5), c(3.5, 2.0)),
  Sigma1 = Sigma,
  mu2 = rbind(c(0.0, 0.0), c(0.0, 0.0), c(0.0, 0.0), c(0.0, 0.0)),
  Sigma2 = Sigma,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 500L, method = 'MC',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE, seed = 42L
)
print(oc_ctrl)
```

### 5.3 Figure: Go/NoGo/Gray Probabilities vs. True Treatment Effect

Operating characteristics are plotted as a function of the treatment
effect on Endpoint 1 ($\mu_{1,1} - \mu_{2,1}$). Both endpoints increase
together ($\mu_{1,2} = 0.6 \times \mu_{1,1}$) to show the transition
from NoGo through Gray to Go clearly.

```{r oc-plot, fig.cap = "Operating characteristics (Go/NoGo/Gray) as a function of the true treatment effect on Endpoint 1. Both endpoints increase proportionally (mu1_2 = 0.6 * mu1_1). Controlled design, vague prior, MC method. n1 = n2 = 20, gamma1 = 0.80, gamma2 = 0.20."}
delta1_seq <- seq(0.0, 3.5, by = 0.25)

oc_grid <- pbayesdecisionprob2cont(
  nsim = 300L, prob = 'posterior', design = 'controlled',
  prior = 'vague',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.20,
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = 20L, m1 = NULL, m2 = NULL,
  mu1    = cbind(delta1_seq, 0.6 * delta1_seq),
  Sigma1 = Sigma,
  mu2    = matrix(0, nrow = length(delta1_seq), ncol = 2),
  Sigma2 = Sigma,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 500L, method = 'MC',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE, seed = 42L
)

old_par <- par(mar = c(4.5, 4.5, 3, 1))

plot(delta1_seq, oc_grid$Go,
     type = "b", pch = 16, lwd = 2, col = "#2166AC",
     ylim = c(0, 1),
     xlab = expression(mu[paste("1,1")] - mu[paste("2,1")]),
     ylab = "Probability",
     main = "Operating Characteristics: Two Continuous Endpoints",
     las  = 1)
lines(delta1_seq, oc_grid$NoGo,
      type = "b", pch = 17, lwd = 2, col = "#D6604D")
lines(delta1_seq, oc_grid$Gray,
      type = "b", pch = 15, lwd = 2, col = "#878787")

abline(h = c(0.20, 0.80), lty = 2, col = "gray50")
abline(v = c(0.5, 1.5),   lty = 3, col = "gray70")

text(0.5 + 0.05, 0.04, expression(theta[MAV1]), col = "gray40", cex = 0.8, adj = 0)
text(1.5 + 0.05, 0.04, expression(theta[TV1]),  col = "gray40", cex = 0.8, adj = 0)

legend("right",
       legend = c("Pr(Go)", "Pr(NoGo)", "Pr(Gray)"),
       col    = c("#2166AC", "#D6604D", "#878787"),
       pch    = c(16, 17, 15), lwd = 2, bty = "n")

par(old_par)
```

### 5.4 Uncontrolled Design — Operating Characteristics

```{r oc-uncontrolled}
oc_unctrl <- pbayesdecisionprob2cont(
  nsim = 300L, prob = 'posterior', design = 'uncontrolled',
  prior = 'N-Inv-Wishart',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.20,
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = NULL, m1 = NULL, m2 = NULL,
  mu1    = rbind(c(0.5, 0.3), c(1.5, 1.0), c(2.5, 1.5), c(3.5, 2.0)),
  Sigma1 = Sigma,
  mu2    = rbind(c(0.0, 0.0), c(0.0, 0.0), c(0.0, 0.0), c(0.0, 0.0)),
  Sigma2 = Sigma,
  kappa01 = 2.0, nu01 = 5.0, mu01 = c(2.0, 1.0), Lambda01 = L0oc,
  kappa02 = NULL, nu02 = NULL, mu02 = c(0.0, 0.0), Lambda02 = NULL,
  r = 1.0,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 500L, method = 'MC',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE, seed = 1L
)
print(oc_unctrl)
```

### 5.5 External Control Design — Operating Characteristics

```{r oc-external}
Se2_oc <- matrix(c(7.0, 1.2, 1.2, 1.8), 2, 2)

oc_ext <- pbayesdecisionprob2cont(
  nsim = 300L, prob = 'posterior', design = 'external',
  prior = 'N-Inv-Wishart',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.20,
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  n1 = 20L, n2 = 20L, m1 = NULL, m2 = NULL,
  mu1    = rbind(c(0.5, 0.3), c(1.5, 1.0), c(2.5, 1.5), c(3.5, 2.0)),
  Sigma1 = Sigma,
  mu2    = rbind(c(0.0, 0.0), c(0.0, 0.0), c(0.0, 0.0), c(0.0, 0.0)),
  Sigma2 = Sigma,
  kappa01 = 2.0, nu01 = 5.0, mu01 = c(2.0, 1.0), Lambda01 = L0oc,
  kappa02 = 2.0, nu02 = 5.0, mu02 = c(0.0, 0.0), Lambda02 = L0oc,
  r = NULL,
  ne1 = NULL, ne2 = 15L, alpha01e = NULL, alpha02e = 0.5,
  ybar_e1 = NULL, ybar_e2 = c(0.2, 0.1), Se1 = NULL, Se2 = Se2_oc,
  nMC = 500L, method = 'MC',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE, seed = 2L
)
print(oc_ext)
```

---

## 6. Optimal Threshold Search

### 6.1 Objective and Algorithm

Because there are two decision thresholds $(\gamma_1, \gamma_2)$, the
optimal search is performed over a **two-dimensional grid**
$\Gamma_1 \times \Gamma_2$.

`getgamma2cont()` uses a three-stage strategy:

**Stage 1 (Simulate and precompute)**: $n_\mathrm{sim}$ bivariate datasets
are generated from the null scenario $(\boldsymbol{\mu}_1^0,
\boldsymbol{\mu}_2^0, \Sigma_1, \Sigma_2)$.  For each replicate $i$,
`pbayespostpred2cont()` is called once in vectorised mode to return the
full region probability vector.  The Go and NoGo probabilities per
replicate are:

$$\hat{g}_{\mathrm{Go},i} = \sum_{j \in \text{GoRegions}} \hat{p}_{ij}, \qquad
  \hat{g}_{\mathrm{NoGo},i} = \sum_{j \in \text{NoGoRegions}} \hat{p}_{ij}.$$

**Stage 2 (Gamma sweep)**: For every pair
$(\gamma_1, \gamma_2) \in \Gamma_1 \times \Gamma_2$:

$$\Pr(\mathrm{Go} \mid \gamma_1, \gamma_2) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},i} \ge \gamma_1 \text{ and }
  \hat{g}_{\mathrm{NoGo},i} < \gamma_2\right],$$

$$\Pr(\mathrm{NoGo} \mid \gamma_1, \gamma_2) =
  \frac{1}{n_\mathrm{sim}} \sum_{i=1}^{n_\mathrm{sim}}
  \mathbf{1}\!\left[\hat{g}_{\mathrm{NoGo},i} \ge \gamma_2 \text{ and }
  \hat{g}_{\mathrm{Go},i} < \gamma_1\right].$$

The results are stored as $|\Gamma_1| \times |\Gamma_2|$ matrices
`PrGo_grid` and `PrNoGo_grid` — ideal for contour visualisation.

**Stage 3 (Optimal selection)**: For each candidate $\gamma_1$, the
worst-case (maximum) $\Pr(\mathrm{Go})$ over all $\gamma_2$ is compared
against `target_go` using `crit_go`; analogously for $\gamma_2$.

### 6.2 Example: Controlled Design, Posterior Probability

```{r getgamma-ctrl}
res_gamma <- getgamma2cont(
  nsim = 500L, prob = 'posterior', design = 'controlled',
  prior = 'vague',
  GoRegions = 1L, NoGoRegions = 9L,
  mu1 = c(0.5, 0.3), Sigma1 = Sigma,   # null: effect at MAV boundary
  mu2 = c(0.0, 0.0), Sigma2 = Sigma,
  target_go = 0.05, target_nogo = 0.20,
  crit_go = '<', crit_nogo = '<',
  sel_go  = 'smallest', sel_nogo = 'largest',
  n1 = 20L, n2 = 20L,
  theta.TV1  = 1.5, theta.MAV1 = 0.5,
  theta.TV2  = 1.0, theta.MAV2 = 0.3,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  m1 = NULL, m2 = NULL,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 500L, method = 'MC',
  gamma1_grid = seq(0.05, 0.95, by = 0.05),
  gamma2_grid = seq(0.05, 0.95, by = 0.05),
  seed = 42L
)
cat(sprintf("Optimal gamma1* = %.2f  (Pr(Go) = %.4f)\n",
            res_gamma$gamma1, res_gamma$PrGo_at_gamma))
cat(sprintf("Optimal gamma2* = %.2f  (Pr(NoGo) = %.4f)\n",
            res_gamma$gamma2, res_gamma$PrNoGo_at_gamma))
```

### 6.3 Figure: Contour Plot of Pr(Go) over $(\gamma_1, \gamma_2)$ Grid

```{r gamma-contour-go, fig.cap = "Contour plot of Pr(Go) as a function of the candidate threshold pair (gamma1, gamma2) under the null scenario (mu1 = (0.5, 0.3), mu2 = (0, 0)). The dashed lines mark the optimal gamma1* (blue) and gamma2* (red). The 0.05 contour is highlighted in bold."}
g1   <- res_gamma$gamma1_grid
g2   <- res_gamma$gamma2_grid
Z_go <- res_gamma$PrGo_grid   # rows = gamma1, cols = gamma2

# Cap upper limit at 95th percentile to avoid sparse high-value bands
go_upper  <- ceiling(quantile(Z_go, 0.95, na.rm = TRUE) * 20) / 20
go_upper  <- max(go_upper, 0.10)
go_breaks <- seq(0, go_upper, length.out = 11)
go_cols   <- hcl.colors(10, palette = "Blues", rev = TRUE)

old_par2 <- par(mar = c(4.5, 4.5, 3.5, 5))

image(x = g1, y = g2, z = pmin(Z_go, go_upper),
      breaks = go_breaks, col = go_cols,
      xlab = expression(gamma[1]),
      ylab = expression(gamma[2]),
      main = expression("Pr(Go) over (" * gamma[1] * ", " * gamma[2] * ") grid \u2014 Null scenario"),
      las  = 1)
contour(x = g1, y = g2, z = Z_go,
        levels = 0.05, add = TRUE,
        lwd = 2.5, col = "#2166AC", labcex = 0.9)
if (!is.na(res_gamma$gamma1))
  abline(v = res_gamma$gamma1, lty = 2, lwd = 2, col = "#2166AC")
if (!is.na(res_gamma$gamma2))
  abline(h = res_gamma$gamma2, lty = 2, lwd = 2, col = "#D6604D")
legend("topright",
       legend = c(
         paste0("gamma1* = ", res_gamma$gamma1),
         paste0("gamma2* = ", res_gamma$gamma2),
         "Pr(Go) = 0.05 contour"),
       col = c("#2166AC", "#D6604D", "#2166AC"),
       lty = c(2, 2, 1), lwd = c(2, 2, 2.5),
       bty = "n", cex = 0.85)

# Color bar
par(new = TRUE, mar = c(4.5, 0, 3.5, 3))
barplot(rep(1, 10), col = go_cols, border = NA,
        horiz = FALSE, axes = FALSE, space = 0,
        width = 1, xlim = c(0, 10))
axis(4, at = seq(0, 10, length.out = 6),
     labels = round(seq(0, go_upper, length.out = 6), 2),
     las = 1, cex.axis = 0.8)

par(old_par2)
```

### 6.4 Figure: Contour Plot of Pr(NoGo) over $(\gamma_1, \gamma_2)$ Grid

```{r gamma-contour-nogo, fig.cap = "Contour plot of Pr(NoGo) as a function of (gamma1, gamma2) under the null scenario. The 0.20 contour is highlighted in bold."}
Z_nogo <- res_gamma$PrNoGo_grid

nogo_upper  <- ceiling(quantile(Z_nogo, 0.95, na.rm = TRUE) * 20) / 20
nogo_upper  <- max(nogo_upper, 0.30)
nogo_breaks <- seq(0, nogo_upper, length.out = 11)
nogo_cols   <- hcl.colors(10, palette = "Reds", rev = TRUE)

old_par3 <- par(mar = c(4.5, 4.5, 3.5, 5))

image(x = g1, y = g2, z = pmin(Z_nogo, nogo_upper),
      breaks = nogo_breaks, col = nogo_cols,
      xlab = expression(gamma[1]),
      ylab = expression(gamma[2]),
      main = expression("Pr(NoGo) over (" * gamma[1] * ", " * gamma[2] * ") grid \u2014 Null scenario"),
      las  = 1)
contour(x = g1, y = g2, z = Z_nogo,
        levels = 0.20, add = TRUE,
        lwd = 2.5, col = "#D6604D", labcex = 0.9)
if (!is.na(res_gamma$gamma1))
  abline(v = res_gamma$gamma1, lty = 2, lwd = 2, col = "#2166AC")
if (!is.na(res_gamma$gamma2))
  abline(h = res_gamma$gamma2, lty = 2, lwd = 2, col = "#D6604D")
legend("topright",
       legend = c(
         paste0("gamma1* = ", res_gamma$gamma1),
         paste0("gamma2* = ", res_gamma$gamma2),
         "Pr(NoGo) = 0.20 contour"),
       col = c("#2166AC", "#D6604D", "#D6604D"),
       lty = c(2, 2, 1), lwd = c(2, 2, 2.5),
       bty = "n", cex = 0.85)

# Color bar
par(new = TRUE, mar = c(4.5, 0, 3.5, 3))
barplot(rep(1, 10), col = nogo_cols, border = NA,
        horiz = FALSE, axes = FALSE, space = 0,
        width = 1, xlim = c(0, 10))
axis(4, at = seq(0, 10, length.out = 6),
     labels = round(seq(0, nogo_upper, length.out = 6), 2),
     las = 1, cex.axis = 0.8)

par(old_par3)
```

### 6.5 Example: Controlled Design, Predictive Probability

```{r getgamma-pred}
res_gamma_pred <- getgamma2cont(
  nsim = 500L, prob = 'predictive', design = 'controlled',
  prior = 'vague',
  GoRegions = 1L, NoGoRegions = 4L,
  mu1 = c(0.5, 0.3), Sigma1 = Sigma,
  mu2 = c(0.0, 0.0), Sigma2 = Sigma,
  target_go = 0.05, target_nogo = 0.20,
  crit_go = '<', crit_nogo = '<',
  sel_go  = 'smallest', sel_nogo = 'largest',
  n1 = 20L, n2 = 20L,
  theta.TV1 = NULL, theta.MAV1 = NULL,
  theta.TV2 = NULL, theta.MAV2 = NULL,
  theta.NULL1 = 0.5, theta.NULL2 = 0.3,
  m1 = 60L, m2 = 60L,
  kappa01 = NULL, nu01 = NULL, mu01 = NULL, Lambda01 = NULL,
  kappa02 = NULL, nu02 = NULL, mu02 = NULL, Lambda02 = NULL,
  r = NULL,
  ne1 = NULL, ne2 = NULL, alpha01e = NULL, alpha02e = NULL,
  ybar_e1 = NULL, ybar_e2 = NULL, Se1 = NULL, Se2 = NULL,
  nMC = 500L, method = 'MC',
  gamma1_grid = seq(0.05, 0.95, by = 0.05),
  gamma2_grid = seq(0.05, 0.95, by = 0.05),
  seed = 3L
)
cat(sprintf("Optimal gamma1* = %.2f  (Pr(Go) = %.4f)\n",
            res_gamma_pred$gamma1, res_gamma_pred$PrGo_at_gamma))
cat(sprintf("Optimal gamma2* = %.2f  (Pr(NoGo) = %.4f)\n",
            res_gamma_pred$gamma2, res_gamma_pred$PrNoGo_at_gamma))
```

---

## 7. Summary

This vignette covered two continuous endpoint analysis in BayesianQDM:

1. **Bayesian model**: NIW conjugate posterior with explicit update
   formulae for $\kappa_{nk}$, $\nu_{nk}$, $\boldsymbol{\mu}_{nk}$,
   and $\Lambda_{nk}$; bivariate non-standardised $t$ marginal posterior.
2. **Nine-region grid** for posterior probability and **four-region grid**
   for predictive probability; typical choice Go = R1, NoGo = R9 (or R4).
3. **Two computation methods**: MC (vectorised pre-generated draws with
   Cholesky reuse) and MM (fourth-moment matching per component +
   `mvtnorm::pmvt` for the joint probability).
4. **Three study designs**: controlled, uncontrolled ($\boldsymbol{\mu}_{0c}$
   and $r$ fixed), external (NIW power prior with explicit augmentation
   formulae for $\kappa_{0c}^{\mathrm{aug}}$, $\nu_{0c}^{\mathrm{aug}}$,
   $\boldsymbol{\mu}_{0c}^{\mathrm{aug}}$, $\Lambda_{0c}^{\mathrm{aug}}$).
5. **Operating characteristics**: Monte Carlo estimation with vectorised
   region probability computation.
6. **Optimal threshold search**: three-stage strategy in `getgamma2cont()`
   producing $|\Gamma_1| \times |\Gamma_2|$ matrices `PrGo_grid` and
   `PrNoGo_grid` — visualised as contour plots with target level curves
   ($\Pr(\mathrm{Go}) < 0.05$ under null, $\Pr(\mathrm{NoGo}) \ge 0.20$
   under null).

See `vignette("two-binary")` for the analogous two binary endpoint analysis.
