---
title: "Two Binary Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two Binary Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 5.5
)
library(BayesianQDM)
```

## Motivating Scenario

We extend the single-endpoint RA trial to a **bivariate binary** design
with two co-primary binary endpoints:

- **Endpoint 1**: ACR20 response (1 = responder, 0 = non-responder)
- **Endpoint 2**: DAS28 remission (1 = achieved, 0 = not achieved)

The trial enrolls $n_1 = n_2 = 10$ patients per arm in a 1:1 randomised
controlled design.

**Clinically meaningful thresholds** (posterior probability):

|                          | Endpoint 1 | Endpoint 2 |
|--------------------------|-----------|-----------|
| $\theta_{\mathrm{TV}}$   | 0.20      | 0.20      |
| $\theta_{\mathrm{MAV}}$  | 0.10      | 0.10      |

**Null hypothesis thresholds** (predictive probability):
$\theta_{\mathrm{NULL},1} = 0.15$, $\theta_{\mathrm{NULL},2} = 0.15$.

**Decision thresholds**: $\gamma_1 = 0.80$ (Go), $\gamma_2 = 0.80$ (NoGo).

**Observed data** (used in Sections 2--4): treatment arm counts for the
four response patterns $(Y_1, Y_2) \in \{(0,0),(0,1),(1,0),(1,1)\}$:

| Pattern | Treatment ($x_{1,lm}$) | Control ($x_{2,lm}$) |
|---------|:---:|:---:|
| (0, 0)  | 2   | 4   |
| (0, 1)  | 1   | 2   |
| (1, 0)  | 3   | 2   |
| (1, 1)  | 4   | 2   |

Marginal responder counts: treatment $y_{1,1} = 7$, $y_{1,2} = 5$;
control $y_{2,1} = 4$, $y_{2,2} = 4$.

---

## 1. Bayesian Model: Dirichlet-Multinomial Conjugate

### 1.1 Response Pattern Parameterisation

Because the two endpoints within each patient are **correlated**, the
bivariate binary outcome $(Y_{k1}, Y_{k2})$ is modelled jointly through
its four-cell probability vector

$$\mathbf{p}_k = (p_{k,00},\, p_{k,01},\, p_{k,10},\, p_{k,11})^\top,
\qquad \mathbf{p}_k \in \Delta^3,$$

where $\Delta^3$ denotes the 3-simplex (all entries non-negative and
summing to 1), and the subscripts refer to the values of
$(Y_{k1}, Y_{k2})$.

The observed count vector $\mathbf{x}_k = (x_{k,00}, x_{k,01}, x_{k,10},
x_{k,11})^\top$ with $\sum_{lm} x_{k,lm} = n_k$ follows:

$$\mathbf{x}_k \mid \mathbf{p}_k \;\sim\;
  \mathrm{Multinomial}(n_k,\, \mathbf{p}_k).$$

The marginal response rates and treatment effects are obtained by summing:

$$\pi_{k1} = p_{k,10} + p_{k,11}, \qquad
  \pi_{k2} = p_{k,01} + p_{k,11},$$

$$\theta_1 = \pi_{11} - \pi_{21}, \qquad
  \theta_2 = \pi_{12} - \pi_{22}.$$

### 1.2 Prior: Dirichlet Distribution

The conjugate prior for $\mathbf{p}_k$ is the Dirichlet distribution:

$$\mathbf{p}_k \;\sim\; \mathrm{Dir}(\alpha_{k,00},\, \alpha_{k,01},\,
  \alpha_{k,10},\, \alpha_{k,11}),$$

where all hyperparameters $\alpha_{k,lm} > 0$.  The symmetric choice
$\alpha_{k,lm} = 0.25$ for all $lm$ is the natural extension of the
Jeffreys prior to the four-cell multinomial and is used throughout
this vignette.

### 1.3 Posterior Distribution

By conjugacy of the Dirichlet-Multinomial model, the posterior is:

$$\mathbf{p}_k \mid \mathbf{x}_k \;\sim\;
  \mathrm{Dir}(\alpha_{k,00}^*,\, \alpha_{k,01}^*,\,
  \alpha_{k,10}^*,\, \alpha_{k,11}^*),$$

where the posterior parameters are simply

$$\alpha_{k,lm}^* = \alpha_{k,lm} + x_{k,lm}, \qquad
  lm \in \{00, 01, 10, 11\}.$$

The posterior mean for each cell is
$E[p_{k,lm} \mid \mathbf{x}_k] = \alpha_{k,lm}^* / A_k^*$,
where $A_k^* = \sum_{lm} \alpha_{k,lm}^*$.

### 1.4 Within-Arm Correlation

The within-arm Pearson correlation between Endpoint 1 and Endpoint 2 is

$$\rho_k = \frac{p_{k,11} - \pi_{k1}\,\pi_{k2}}
               {\sqrt{\pi_{k1}(1-\pi_{k1})\,\pi_{k2}(1-\pi_{k2})}}.$$

When specifying operating characteristics scenarios, the true parameter
vector $\mathbf{p}_k$ is specified via the reparameterisation
$(\pi_{k1}, \pi_{k2}, \rho_k)$, with the feasibility constraint:

$$\rho_{\min} = \frac{\max(0,\pi_{k1}+\pi_{k2}-1) - \pi_{k1}\pi_{k2}}
                     {\sqrt{\pi_{k1}(1-\pi_{k1})\pi_{k2}(1-\pi_{k2})}}
\;\le\; \rho_k \;\le\;
\frac{\min(\pi_{k1},\pi_{k2}) - \pi_{k1}\pi_{k2}}
     {\sqrt{\pi_{k1}(1-\pi_{k1})\pi_{k2}(1-\pi_{k2})}}
= \rho_{\max}.$$

The helper function `getjointbin()` converts $(\pi_{k1}, \pi_{k2}, \rho_k)$
to $(p_{k,00}, p_{k,01}, p_{k,10}, p_{k,11})$:

```{r getjointbin}
# Convert marginal rates + correlation to cell probabilities
getjointbin(pi1 = 0.30, pi2 = 0.35, rho = 0.20)
getjointbin(pi1 = 0.20, pi2 = 0.20, rho = 0.00)   # independence
```

### 1.5 Nine-Region Grid (Posterior Probability)

The bivariate threshold grid for $(\theta_1, \theta_2)$ creates the same
$3 \times 3$ partition as for the two continuous endpoint case:

```{r nine-region-bin, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Nine-region grid for two-endpoint posterior probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="3" style="border:1px solid #aaa; padding:6px;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &gt; &theta;<sub>TV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>TV1</sub> &ge; &theta;<sub>1</sub> &gt; &theta;<sub>MAV1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>MAV1</sub> &ge; &theta;<sub>1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="3" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg); font-weight:bold;">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>TV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
    <td style="border:1px solid #aaa; padding:6px;">R7</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>TV2</sub> &ge; &theta;<sub>2</sub> &gt; &theta;<sub>MAV2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R5</td>
    <td style="border:1px solid #aaa; padding:6px;">R8</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>MAV2</sub> &ge; &theta;<sub>2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
    <td style="border:1px solid #aaa; padding:6px;">R6</td>
    <td style="border:1px solid #aaa; padding:6px;">R9</td>
  </tr>
</tbody>
</table>
')
```

Typical choices: **Go if $P(R1) \ge \gamma_1$**,
**NoGo if $P(R9) \ge \gamma_2$**.

### 1.6 Four-Region Grid (Predictive Probability)

For the predictive probability, a $2 \times 2$ partition is used:

```{r four-region-bin, echo = FALSE, results = 'asis'}
cat('
<table style="border-collapse:collapse; text-align:center; font-size:0.9em;">
<caption>Four-region grid for two-endpoint predictive probability</caption>
<thead>
  <tr>
    <th colspan="2" rowspan="2" style="border:1px solid #aaa; background:
      linear-gradient(to top right, white 49.5%, #aaa 49.5%, #aaa 50.5%, white 50.5%);
      min-width:80px; min-height:60px; padding:4px;">
    </th>
    <th colspan="2" style="border:1px solid #aaa; padding:6px;">Endpoint 1</th>
  </tr>
  <tr>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &gt; &theta;<sub>NULL1</sub></th>
    <th style="border:1px solid #aaa; padding:6px;">&theta;<sub>1</sub> &le; &theta;<sub>NULL1</sub></th>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="2" style="border:1px solid #aaa; padding:6px; writing-mode:vertical-rl;
        transform:rotate(180deg); font-weight:bold;">Endpoint 2</td>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &gt; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R1</td>
    <td style="border:1px solid #aaa; padding:6px;">R3</td>
  </tr>
  <tr>
    <td style="border:1px solid #aaa; padding:6px; text-align:left;">
      &theta;<sub>2</sub> &le; &theta;<sub>NULL2</sub></td>
    <td style="border:1px solid #aaa; padding:6px;">R2</td>
    <td style="border:1px solid #aaa; padding:6px;">R4</td>
  </tr>
</tbody>
</table>
')
```

---

## 2. Posterior Predictive Distribution

### 2.1 Dirichlet-Multinomial Predictive Distribution

Let $\tilde{\mathbf{x}}_k \sim \mathrm{Multinomial}(m_k, \mathbf{p}_k)$
be the future count vector with $m_k$ future patients.  Integrating out
$\mathbf{p}_k$ over its Dirichlet posterior gives the Dirichlet-Multinomial
predictive distribution:

$$P(\tilde{\mathbf{x}}_k = \mathbf{c} \mid \mathbf{x}_k) =
  \binom{m_k}{\mathbf{c}}
  \frac{B(\boldsymbol{\alpha}_k^* + \mathbf{c})}{B(\boldsymbol{\alpha}_k^*)},
  \quad \sum_{lm} c_{lm} = m_k,$$

where $B(\cdot)$ is the multivariate Beta function and
$\binom{m_k}{\mathbf{c}} = m_k! / \prod_{lm} c_{lm}!$ is the multinomial
coefficient.

### 2.2 Monte Carlo Evaluation

Because the Dirichlet-Multinomial does not yield a tractable closed form
for the joint probability that $(\tilde\theta_1, \tilde\theta_2)$ falls
in a given region, all region probabilities are estimated by Monte Carlo:

$$\widehat{P}(R_j) = \frac{1}{n_\mathrm{MC}} \sum_{i=1}^{n_\mathrm{MC}}
  \mathbf{1}\!\left[(p_1^{(i)} - p_2^{(i)}) \in R_j\right],$$

where $p_k^{(i)} \sim \mathrm{Dir}(\boldsymbol{\alpha}_k^*)$ are draws
from the Dirichlet posterior.  For the predictive probability, future
proportion differences $\tilde\theta_\ell^{(i)} = \tilde{y}_{1\ell}^{(i)}/m_1
- \tilde{y}_{2\ell}^{(i)}/m_2$ are computed from corresponding
Binomial draws conditioned on the Dirichlet samples.

---

## 3. Study Designs

### 3.1 Controlled Design

Both arms are observed in the PoC trial.  The posterior parameters are:

$$\alpha_{k,lm}^* = \alpha_{k,lm} + x_{k,lm}.$$

```{r ctrl-post}
set.seed(42)
p_post_ctrl <- pbayespostpred2bin(
  prob = 'posterior', design = 'controlled',
  theta.TV1 = 0.20, theta.MAV1 = 0.10,
  theta.TV2 = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  x1_00 = 2L, x1_01 = 1L, x1_10 = 3L, x1_11 = 4L,
  x2_00 = 4L, x2_01 = 2L, x2_10 = 2L, x2_11 = 2L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 1000L
)
print(round(p_post_ctrl, 4))
cat(sprintf(
  "\nGo  region (R1): P = %.4f  >= gamma1 (0.80)? %s\n",
  p_post_ctrl["R1"], ifelse(p_post_ctrl["R1"] >= 0.80, "YES -> Go", "NO")
))
cat(sprintf(
  "NoGo region (R9): P = %.4f  >= gamma2 (0.80)? %s\n",
  p_post_ctrl["R9"], ifelse(p_post_ctrl["R9"] >= 0.80, "YES -> NoGo", "NO")
))
```

**Posterior predictive probability** (future Phase III: $m_1 = m_2 = 20$):

```{r ctrl-pred}
set.seed(42)
p_pred_ctrl <- pbayespostpred2bin(
  prob = 'predictive', design = 'controlled',
  theta.TV1 = NULL, theta.MAV1 = NULL,
  theta.TV2 = NULL, theta.MAV2 = NULL,
  theta.NULL1 = 0.15, theta.NULL2 = 0.15,
  x1_00 = 2L, x1_01 = 1L, x1_10 = 3L, x1_11 = 4L,
  x2_00 = 4L, x2_01 = 2L, x2_10 = 2L, x2_11 = 2L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = 15L, m2 = 15L,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 1000L
)
print(round(p_pred_ctrl, 4))
cat(sprintf(
  "\nGo region (R1): P = %.4f  >= gamma1 (0.80)? %s\n",
  p_pred_ctrl["R1"], ifelse(p_pred_ctrl["R1"] >= 0.80, "YES -> Go", "NO")
))
```

### 3.2 Uncontrolled Design

When no concurrent control arm is enrolled, the control distribution is
specified via a **hypothetical count vector**
$\mathbf{z} = (z_{00}, z_{01}, z_{10}, z_{11})$ combined with the
Dirichlet prior:

$$\mathbf{p}_2 \;\sim\;
  \mathrm{Dir}(\alpha_{2,00} + z_{00},\; \alpha_{2,01} + z_{01},\;
  \alpha_{2,10} + z_{10},\; \alpha_{2,11} + z_{11}).$$

This specification encodes prior belief about the bivariate control
response rates — including any assumed within-arm correlation — without
requiring observed control data.

```{r unctrl-post}
set.seed(1)
p_unctrl <- pbayespostpred2bin(
  prob = 'posterior', design = 'uncontrolled',
  theta.TV1 = 0.20, theta.MAV1 = 0.10,
  theta.TV2 = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  x1_00 = 2L, x1_01 = 1L, x1_10 = 3L, x1_11 = 4L,
  x2_00 = NULL, x2_01 = NULL, x2_10 = NULL, x2_11 = NULL,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  z00 = 2L, z01 = 1L, z10 = 2L, z11 = 1L,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 1000L
)
print(round(p_unctrl, 4))
```

### 3.3 External Control Design (Power Prior)

External data for group $k$ (count vector $\mathbf{x}_{ek}$) are
incorporated via a **Dirichlet power prior** with weight
$a_{0k} \in (0, 1]$.  The augmented prior parameters are:

$$\alpha_{k,lm}^{*} = \alpha_{k,lm} + a_{0k} \cdot x_{ek,lm}.$$

Setting $a_{0k} = 1$ treats the external data as equivalent to current
trial data; $a_{0k} \to 0$ recovers the original prior.  The current
PoC data then update these augmented parameters in the usual way:

$$\alpha_{k,lm}^{**} = \alpha_{k,lm}^{*} + x_{k,lm}
  = \alpha_{k,lm} + a_{0k}\, x_{ek,lm} + x_{k,lm}.$$

```{r ext-post}
set.seed(2)
p_ext <- pbayespostpred2bin(
  prob = 'posterior', design = 'external',
  theta.TV1 = 0.20, theta.MAV1 = 0.10,
  theta.TV2 = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  x1_00 = 2L, x1_01 = 1L, x1_10 = 3L, x1_11 = 4L,
  x2_00 = 4L, x2_01 = 2L, x2_10 = 2L, x2_11 = 2L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = 3L,   xe2_01 = 1L,   xe2_10 = 2L,   xe2_11 = 1L,
  ae1 = NULL, ae2 = 0.5,
  nMC = 1000L
)
print(round(p_ext, 4))
```

**Sensitivity to borrowing weight** $a_{0,2}$:

```{r ext-borrowing}
ae_seq <- c(0.01, seq(0.1, 1.0, by = 0.1))
p_ae <- sapply(ae_seq, function(ae) {
  set.seed(99)
  res <- pbayespostpred2bin(
    prob = 'posterior', design = 'external',
    theta.TV1 = 0.20, theta.MAV1 = 0.10,
    theta.TV2 = 0.20, theta.MAV2 = 0.10,
    theta.NULL1 = NULL, theta.NULL2 = NULL,
    x1_00 = 2L, x1_01 = 1L, x1_10 = 3L, x1_11 = 4L,
    x2_00 = 4L, x2_01 = 2L, x2_10 = 2L, x2_11 = 2L,
    a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
    a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
    m1 = NULL, m2 = NULL,
    z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
    xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
    xe2_00 = 3L,   xe2_01 = 1L,   xe2_10 = 2L,   xe2_11 = 1L,
    ae1 = NULL, ae2 = ae, nMC = 500L   # fast: single posterior call per ae
  )
  res["R1"]
})
data.frame(ae2 = ae_seq, P_R1 = round(p_ae, 4))
```

---

## 4. Operating Characteristics

### 4.1 Definition

For given true parameters $(\mathbf{p}_1, \mathbf{p}_2)$, the operating
characteristics are computed by **exact enumeration** over all possible
multinomial count combinations via a two-stage strategy.

**Stage 1 (precomputation)**: All count combinations
$\mathbf{x}_{t,i}$ ($K_t$ combinations) and $\mathbf{x}_{c,j}$ ($K_c$
combinations) are enumerated by `allmultinom()`.  For each pair $(i, j)$,
`pbayespostpred2bin()` computes the region probability vector once,
yielding $\hat{g}_{\mathrm{Go},ij}$ and $\hat{g}_{\mathrm{NoGo},ij}$
independent of any decision threshold.

**Stage 2 (weighting)**: For given thresholds $(\gamma_1, \gamma_2)$:

$$\Pr(\mathrm{Go}) = \sum_{i=1}^{K_t} \sum_{j=1}^{K_c}
  w_{ij}\, \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},ij} \ge \gamma_1
  \text{ and } \hat{g}_{\mathrm{NoGo},ij} < \gamma_2\right],$$

$$\Pr(\mathrm{NoGo}) = \sum_{i=1}^{K_t} \sum_{j=1}^{K_c}
  w_{ij}\, \mathbf{1}\!\left[\hat{g}_{\mathrm{NoGo},ij} \ge \gamma_2
  \text{ and } \hat{g}_{\mathrm{Go},ij} < \gamma_1\right],$$

where $w_{ij} = P_\mathrm{Mult}(\mathbf{x}_{t,i};\, n_1, \mathbf{p}_1)
\times P_\mathrm{Mult}(\mathbf{x}_{c,j};\, n_2, \mathbf{p}_2)$,
and

$$\hat{g}_{\mathrm{Go},ij}   = \sum_{r \in \text{GoRegions}}   \hat{P}(R_r \mid \mathbf{x}_{t,i}, \mathbf{x}_{c,j}),$$
$$\hat{g}_{\mathrm{NoGo},ij} = \sum_{r \in \text{NoGoRegions}} \hat{P}(R_r \mid \mathbf{x}_{t,i}, \mathbf{x}_{c,j}).$$

A Miss ($\hat{g}_{\mathrm{Go},ij} \ge \gamma_1$ AND
$\hat{g}_{\mathrm{NoGo},ij} \ge \gamma_2$) indicates an inconsistent
threshold configuration and triggers an error by default
(`error_if_Miss = TRUE`). Gray comprises all remaining outcomes.

For large $n_1$ or $n_2$, the `method = 'MC'` option replaces full
enumeration with $n_\mathrm{sim}$ multinomial draws, deduplicates them
into $K \ll n_\mathrm{sim}$ unique pairs, and evaluates
`pbayespostpred2bin()` only for those unique pairs.

> **Note on computation time.** The number of multinomial count
> combinations grows rapidly with $n$: for $n_k$ patients and 4 response
> categories the number of combinations is $\binom{n_k + 3}{3}$, which
> equals 286 for $n_k = 10$ and 1771 for $n_k = 20$.  For the Exact method
> with two arms this means up to $286^2 \approx 82{,}000$ unique pairs each
> requiring `nMC` Dirichlet draws.  When predictive probability is used
> ($m_k > 0$), an additional layer of Multinomial sampling is added inside
> each Dirichlet draw, further multiplying computation time.  Use small
> `nMC` (100--500) and `n1 = n2 = 10` for demonstration; switch to
> `method = 'MC'` with moderate `nsim` for larger sample sizes.

### 4.2 Controlled Design --- Multiple Scenarios

```{r oc-controlled}
oc_ctrl <- pbayesdecisionprob2bin(
  prob = 'posterior', design = 'controlled',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.80,
  pi_t1 = c(0.15, 0.30, 0.50),
  pi_t2 = c(0.15, 0.30, 0.50),
  rho_t = c(0.0,  0.0,  0.0),
  pi_c1 = rep(0.15, 3),
  pi_c2 = rep(0.15, 3),
  rho_c = rep(0.0,  3),
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 500L, method = 'Exact',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_ctrl)
```

### 4.3 Figure: Go/NoGo/Gray Probabilities vs. True Treatment Effect

```{r oc-plot, fig.cap = "Operating characteristics (Go/NoGo/Gray) as a function of the true treatment effect on Endpoint 1 (pi_t1 - pi_c1) with equal effects on both endpoints and independence (rho = 0). Controlled design, Jeffreys prior, Exact method. n1 = n2 = 10, gamma1 = gamma2 = 0.80."}
pi_t_seq <- seq(0.15, 0.65, by = 0.05)
n_scen   <- length(pi_t_seq)

oc_grid <- pbayesdecisionprob2bin(
  prob = 'posterior', design = 'controlled',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.80,
  pi_t1 = pi_t_seq, pi_t2 = pi_t_seq,
  rho_t = rep(0.0, n_scen),
  pi_c1 = rep(0.15, n_scen),
  pi_c2 = rep(0.15, n_scen),
  rho_c = rep(0.0,  n_scen),
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 200L, method = 'Exact',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)

delta1 <- oc_grid$pi_t1 - 0.15

old_par <- par(mar = c(4.5, 4.5, 3, 1))

plot(delta1, oc_grid$Go,
     type = "b", pch = 16, lwd = 2, col = "#2166AC",
     ylim = c(0, 1),
     xlab = expression(pi[t1] - pi[c1]),
     ylab = "Probability",
     main = "Operating Characteristics: Two Binary Endpoints",
     las  = 1)
lines(delta1, oc_grid$NoGo,
      type = "b", pch = 17, lwd = 2, col = "#D6604D")
lines(delta1, oc_grid$Gray,
      type = "b", pch = 15, lwd = 2, col = "#878787")

abline(h = c(0.20, 0.80), lty = 2, col = "gray50")
abline(v = c(0.10, 0.20), lty = 3, col = "gray70")

text(0.10 + 0.005, 0.04,
     expression(theta[MAV1]), col = "gray40", cex = 0.8, adj = 0)
text(0.20 + 0.005, 0.04,
     expression(theta[TV1]),  col = "gray40", cex = 0.8, adj = 0)

legend("right",
       legend = c("Pr(Go)", "Pr(NoGo)", "Pr(Gray)"),
       col    = c("#2166AC", "#D6604D", "#878787"),
       pch    = c(16, 17, 15), lwd = 2, bty = "n")

par(old_par)
```

### 4.4 Effect of Within-Arm Correlation

The within-arm correlation $\rho_t$ affects the operating characteristics
because it changes the joint cell probabilities and hence the multinomial
weight distribution.

```{r oc-correlation}
rho_seq <- c(-0.3, 0.0, 0.3)

oc_rho <- pbayesdecisionprob2bin(
  prob = 'posterior', design = 'controlled',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.80,
  pi_t1 = rep(0.40, 3), pi_t2 = rep(0.40, 3),
  rho_t = rho_seq,
  pi_c1 = rep(0.15, 3), pi_c2 = rep(0.15, 3),
  rho_c = rep(0.0,  3),
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 500L, method = 'Exact',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_rho)
```

### 4.5 Uncontrolled Design --- Operating Characteristics

```{r oc-uncontrolled}
oc_unctrl <- pbayesdecisionprob2bin(
  prob = 'posterior', design = 'uncontrolled',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.80,
  pi_t1 = c(0.15, 0.30, 0.50),
  pi_t2 = c(0.15, 0.30, 0.50),
  rho_t = c(0.0,  0.0,  0.0),
  pi_c1 = rep(0.15, 3),
  pi_c2 = rep(0.15, 3),
  rho_c = rep(0.0,  3),
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  z00 = 3L, z01 = 2L, z10 = 2L, z11 = 1L,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 500L, method = 'Exact',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_unctrl)
```

### 4.6 External Control Design --- Operating Characteristics

```{r oc-external}
oc_ext <- pbayesdecisionprob2bin(
  prob = 'posterior', design = 'external',
  GoRegions = 1L, NoGoRegions = 9L,
  gamma1 = 0.80, gamma2 = 0.80,
  pi_t1 = c(0.15, 0.30, 0.50),
  pi_t2 = c(0.15, 0.30, 0.50),
  rho_t = c(0.0,  0.0,  0.0),
  pi_c1 = rep(0.15, 3),
  pi_c2 = rep(0.15, 3),
  rho_c = rep(0.0,  3),
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  m1 = NULL, m2 = NULL,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = 3L,   xe2_01 = 2L,   xe2_10 = 3L,   xe2_11 = 2L,
  ae1 = NULL, ae2 = 0.5,
  nMC = 500L, method = 'Exact',
  error_if_Miss = TRUE, Gray_inc_Miss = FALSE
)
print(oc_ext)
```

---

## 5. Optimal Threshold Search

### 5.1 Objective and Algorithm

`getgamma2bin()` finds the optimal pair $(\gamma_1^*, \gamma_2^*)$ by
the same **two-stage precompute-then-sweep** strategy as
`pbayesdecisionprob2bin()`, but sweeps over the two-dimensional grid
$\Gamma_1 \times \Gamma_2$.

**Stage 1 (precomputation)**: `pbayespostpred2bin()` is called for every
unique multinomial outcome combination $(x_{t,i}, x_{c,j})$ enumerated
by `allmultinom()`.  The region probability vector is summed over
`GoRegions` and `NoGoRegions` to obtain $\hat{g}_{\mathrm{Go},ij}$ and
$\hat{g}_{\mathrm{NoGo},ij}$, independent of $(\gamma_1, \gamma_2)$.

**Stage 2 (gamma sweep)**: For every pair
$(\gamma_1, \gamma_2) \in \Gamma_1 \times \Gamma_2$:

$$\Pr(\mathrm{Go} \mid \gamma_1, \gamma_2) =
  \sum_{i,j} w_{ij}\,
  \mathbf{1}\!\left[\hat{g}_{\mathrm{Go},ij} \ge \gamma_1 \text{ and }
  \hat{g}_{\mathrm{NoGo},ij} < \gamma_2\right],$$

$$\Pr(\mathrm{NoGo} \mid \gamma_1, \gamma_2) =
  \sum_{i,j} w_{ij}\,
  \mathbf{1}\!\left[\hat{g}_{\mathrm{NoGo},ij} \ge \gamma_2 \text{ and }
  \hat{g}_{\mathrm{Go},ij} < \gamma_1\right].$$

Results are stored in $|\Gamma_1| \times |\Gamma_2|$ matrices
`PrGo_grid` and `PrNoGo_grid`.

**Stage 3 (optimal selection)**: For each candidate $\gamma_1$, the
worst-case $\Pr(\mathrm{Go})$ over all $\gamma_2$ is compared against
`target_go`; analogously for $\gamma_2$.

### 5.2 Example: Controlled Design, Posterior Probability

Null scenario: $\pi_{t1} = \pi_{c1} = 0.15$, $\pi_{t2} = \pi_{c2} = 0.15$,
$\rho_t = \rho_c = 0$ (no treatment effect, independence).

```{r getgamma-ctrl}
res_gamma <- getgamma2bin(
  prob = 'posterior', design = 'controlled',
  GoRegions = 1L, NoGoRegions = 9L,
  pi_t1 = 0.15, pi_t2 = 0.15, rho_t = 0.0,
  pi_c1 = 0.15, pi_c2 = 0.15, rho_c = 0.0,
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  m1 = NULL, m2 = NULL,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 200L, method = 'Exact',
  gamma1_grid = seq(0.05, 0.95, by = 0.05),
  gamma2_grid = seq(0.05, 0.95, by = 0.05)
)
cat(sprintf("Optimal gamma1* = %.2f  (Pr(Go) = %.4f)\n",
            res_gamma$gamma1, res_gamma$PrGo_at_gamma))
cat(sprintf("Optimal gamma2* = %.2f  (Pr(NoGo) = %.4f)\n",
            res_gamma$gamma2, res_gamma$PrNoGo_at_gamma))
```

### 5.3 Figure: Contour Plot of Pr(Go) over $(\gamma_1, \gamma_2)$ Grid

```{r gamma-contour-go, fig.cap = "Contour plot of Pr(Go) as a function of the candidate threshold pair (gamma1, gamma2) under the null scenario (pi_t = pi_c = (0.15, 0.15), rho = 0). The 0.05 target contour is highlighted in bold (blue). Dashed lines mark the optimal gamma1* (blue) and gamma2* (red)."}
g1   <- res_gamma$gamma1_grid
g2   <- res_gamma$gamma2_grid
Z_go <- res_gamma$PrGo_grid   # rows = gamma1, cols = gamma2

go_upper  <- ceiling(quantile(Z_go, 0.95, na.rm = TRUE) * 20) / 20
go_upper  <- max(go_upper, 0.10)
go_breaks <- seq(0, go_upper, length.out = 11)
go_cols   <- hcl.colors(10, palette = "Blues", rev = TRUE)

old_par2 <- par(mar = c(4.5, 4.5, 3.5, 5))

image(x = g1, y = g2, z = pmin(Z_go, go_upper),
      breaks = go_breaks, col = go_cols,
      xlab = expression(gamma[1]),
      ylab = expression(gamma[2]),
      main = expression("Pr(Go) over (" * gamma[1] * ", " * gamma[2] * ") grid \u2014 Null scenario"),
      las  = 1)
contour(x = g1, y = g2, z = Z_go,
        levels = 0.05, add = TRUE,
        lwd = 2.5, col = "#2166AC", labcex = 0.9)
if (!is.na(res_gamma$gamma1))
  abline(v = res_gamma$gamma1, lty = 2, lwd = 2, col = "#2166AC")
if (!is.na(res_gamma$gamma2))
  abline(h = res_gamma$gamma2, lty = 2, lwd = 2, col = "#D6604D")
legend("topright",
       legend = c(
         paste0("gamma1* = ", res_gamma$gamma1),
         paste0("gamma2* = ", res_gamma$gamma2),
         "Pr(Go) = 0.05 contour"),
       col = c("#2166AC", "#D6604D", "#2166AC"),
       lty = c(2, 2, 1), lwd = c(2, 2, 2.5),
       bty = "n", cex = 0.85)

# Color bar
par(new = TRUE, mar = c(4.5, 0, 3.5, 3))
barplot(rep(1, 10), col = go_cols, border = NA,
        horiz = FALSE, axes = FALSE, space = 0,
        width = 1, xlim = c(0, 10))
axis(4, at = seq(0, 10, length.out = 6),
     labels = round(seq(0, go_upper, length.out = 6), 2),
     las = 1, cex.axis = 0.8)

par(old_par2)
```

### 5.4 Figure: Contour Plot of Pr(NoGo) over $(\gamma_1, \gamma_2)$ Grid

```{r gamma-contour-nogo, fig.cap = "Contour plot of Pr(NoGo) as a function of (gamma1, gamma2) under the null scenario. The 0.20 target contour is highlighted in bold (red). Dashed lines mark the optimal thresholds."}
Z_nogo <- res_gamma$PrNoGo_grid

nogo_upper  <- ceiling(quantile(Z_nogo, 0.95, na.rm = TRUE) * 20) / 20
nogo_upper  <- max(nogo_upper, 0.30)
nogo_breaks <- seq(0, nogo_upper, length.out = 11)
nogo_cols   <- hcl.colors(10, palette = "Reds", rev = TRUE)

old_par3 <- par(mar = c(4.5, 4.5, 3.5, 5))

image(x = g1, y = g2, z = pmin(Z_nogo, nogo_upper),
      breaks = nogo_breaks, col = nogo_cols,
      xlab = expression(gamma[1]),
      ylab = expression(gamma[2]),
      main = expression("Pr(NoGo) over (" * gamma[1] * ", " * gamma[2] * ") grid \u2014 Null scenario"),
      las  = 1)
contour(x = g1, y = g2, z = Z_nogo,
        levels = 0.20, add = TRUE,
        lwd = 2.5, col = "#D6604D", labcex = 0.9)
if (!is.na(res_gamma$gamma1))
  abline(v = res_gamma$gamma1, lty = 2, lwd = 2, col = "#2166AC")
if (!is.na(res_gamma$gamma2))
  abline(h = res_gamma$gamma2, lty = 2, lwd = 2, col = "#D6604D")
legend("topright",
       legend = c(
         paste0("gamma1* = ", res_gamma$gamma1),
         paste0("gamma2* = ", res_gamma$gamma2),
         "Pr(NoGo) = 0.20 contour"),
       col = c("#2166AC", "#D6604D", "#D6604D"),
       lty = c(2, 2, 1), lwd = c(2, 2, 2.5),
       bty = "n", cex = 0.85)

# Color bar
par(new = TRUE, mar = c(4.5, 0, 3.5, 3))
barplot(rep(1, 10), col = nogo_cols, border = NA,
        horiz = FALSE, axes = FALSE, space = 0,
        width = 1, xlim = c(0, 10))
axis(4, at = seq(0, 10, length.out = 6),
     labels = round(seq(0, nogo_upper, length.out = 6), 2),
     las = 1, cex.axis = 0.8)

par(old_par3)
```

### 5.5 Example: Uncontrolled Design, Posterior Probability

```{r getgamma-unctrl}
res_gamma_unctrl <- getgamma2bin(
  prob = 'posterior', design = 'uncontrolled',
  GoRegions = 1L, NoGoRegions = 9L,
  pi_t1 = 0.15, pi_t2 = 0.15, rho_t = 0.0,
  pi_c1 = NULL, pi_c2 = NULL, rho_c = NULL,
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  theta.TV1  = 0.20, theta.MAV1 = 0.10,
  theta.TV2  = 0.20, theta.MAV2 = 0.10,
  theta.NULL1 = NULL, theta.NULL2 = NULL,
  m1 = NULL, m2 = NULL,
  z00 = 3L, z01 = 2L, z10 = 2L, z11 = 1L,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 200L, method = 'Exact',
  gamma1_grid = seq(0.05, 0.95, by = 0.05),
  gamma2_grid = seq(0.05, 0.95, by = 0.05)
)
cat(sprintf("Optimal gamma1* = %.2f  (Pr(Go) = %.4f)\n",
            res_gamma_unctrl$gamma1, res_gamma_unctrl$PrGo_at_gamma))
cat(sprintf("Optimal gamma2* = %.2f  (Pr(NoGo) = %.4f)\n",
            res_gamma_unctrl$gamma2, res_gamma_unctrl$PrNoGo_at_gamma))
```

### 5.6 Example: Controlled Design, Predictive Probability

> **Note.** When `prob = 'predictive'`, `getgamma2bin()` must evaluate
> `pbayespostpred2bin()` for every outcome combination $(x_{t,i}, x_{c,j})$
> using `nMC` Dirichlet draws and $m_k$ future observations each.  This
> substantially increases computation time relative to the posterior case.
> The `method = 'MC'` option with moderate `nsim` and small `nMC` is
> recommended for vignette-level demonstration.

```{r getgamma-pred}
res_gamma_pred <- getgamma2bin(
  prob = 'predictive', design = 'controlled',
  GoRegions = 1L, NoGoRegions = 4L,
  pi_t1 = 0.15, pi_t2 = 0.15, rho_t = 0.0,
  pi_c1 = 0.15, pi_c2 = 0.15, rho_c = 0.0,
  target_go = 0.05, target_nogo = 0.20,
  crit_go   = '<',  crit_nogo   = '<',
  sel_go    = 'smallest', sel_nogo = 'largest',
  n1 = 10L, n2 = 10L,
  a1_00 = 0.25, a1_01 = 0.25, a1_10 = 0.25, a1_11 = 0.25,
  a2_00 = 0.25, a2_01 = 0.25, a2_10 = 0.25, a2_11 = 0.25,
  theta.TV1  = NULL, theta.MAV1 = NULL,
  theta.TV2  = NULL, theta.MAV2 = NULL,
  theta.NULL1 = 0.15, theta.NULL2 = 0.15,
  m1 = 15L, m2 = 15L,
  z00 = NULL, z01 = NULL, z10 = NULL, z11 = NULL,
  xe1_00 = NULL, xe1_01 = NULL, xe1_10 = NULL, xe1_11 = NULL,
  xe2_00 = NULL, xe2_01 = NULL, xe2_10 = NULL, xe2_11 = NULL,
  ae1 = NULL, ae2 = NULL,
  nMC = 100L, method = 'MC', nsim = 1000L,
  gamma1_grid = seq(0.05, 0.95, by = 0.05),
  gamma2_grid = seq(0.05, 0.95, by = 0.05)
)
cat(sprintf("Optimal gamma1* = %.2f  (Pr(Go) = %.4f)\n",
            res_gamma_pred$gamma1, res_gamma_pred$PrGo_at_gamma))
cat(sprintf("Optimal gamma2* = %.2f  (Pr(NoGo) = %.4f)\n",
            res_gamma_pred$gamma2, res_gamma_pred$PrNoGo_at_gamma))
```

---

## 6. Summary

This vignette covered two binary endpoint analysis in BayesianQDM:

1. **Bayesian model**: Dirichlet-Multinomial conjugate, modelling the
   bivariate binary outcome through the four-cell probability vector
   $\mathbf{p}_k$.  Posterior update:
   $\alpha_{k,lm}^* = \alpha_{k,lm} + x_{k,lm}$.  The symmetric
   Jeffreys-type prior $\alpha_{k,lm} = 0.25$ is recommended.
2. **Within-arm correlation**: parameterised via $\rho_k$ using the
   moment-matching identity; feasible range enforced by `getjointbin()`.
3. **Nine-region grid** for posterior probability and **four-region grid**
   for predictive probability, identical in structure to the two continuous
   endpoint case.
4. **Monte Carlo evaluation** of region probabilities via Dirichlet draws;
   no closed-form expression for the joint probability exists in the
   bivariate binary case.
5. **Three study designs**: controlled (standard posterior update),
   uncontrolled (hypothetical count vector $\mathbf{z}$ fixes the control
   Dirichlet distribution), external (Dirichlet power prior with
   $\alpha_{k,lm}^* = \alpha_{k,lm} + a_{0k}\, x_{ek,lm}$).
6. **Exact operating characteristics**: two-stage precomputation over all
   multinomial count combinations via `allmultinom()`, with multinomial
   probability weighting (no outer Monte Carlo needed for small $n$).
7. **Optimal threshold search**: three-stage precompute-then-sweep in
   `getgamma2bin()`, producing $|\Gamma_1| \times |\Gamma_2|$ matrices
   `PrGo_grid` and `PrNoGo_grid`, visualised as contour plots.

See `vignette("overview")` for a comparison across all four endpoint types.
